### 2019-02-19 paper.tex
Add main template.

% please run ``latex paper'' and then ``makeglossaries paper'' from the command line to create the glossaries files
% add. options: [seceqn,secthm,crcready,onecolumn]
\documentclass[sw]{iosart2x}

%\usepackage{dcolumn}
%\usepackage{endnotes}

\usepackage{booktabs}
\usepackage{csquotes}
\usepackage{graphicx}
\usepackage{listings}
\lstset{language=SPARQL,breaklines=true}

\usepackage{cleveref}

%%%%%%%%%%% End of definitions

\pubyear{2019}
\volume{0}
\firstpage{1}
\lastpage{1}

\begin{document}

\begin{frontmatter}

%\pretitle{}
\title{The SNIK Ontology of Hospital Information Management}
\runningtitle{SNIK}
%\subtitle{}

% Two or more authors:
\author[A]{\inits {K.}\fnms{Konrad} \snm{Höffner}\ead[label=e1]{konrad.hoeffner@imise.uni-leipzig.de}%
\thanks{Corresponding author. \printead{e1}.}},
\author[A]{\fnms{Franziska} \snm{Jahn}\ead[label=e2]{konrad.hoeffner@imise.uni-leipzig.de}}
\author[A]{\fnms{Birgit} \snm{Schneider}\ead[label=e3]{birgit.schneider@imise.uni-leipzig.de}}
\author[A]{\fnms{Anna} \snm{Lörke}\ead[label=e4]{anna.loerke@imise.uni-leipzig.de}}
\author[A]{\fnms{Thomas} \snm{Pause}\ead[label=e5]{thomas.pause@imise.uni-leipzig.de}}
\author[A]{\fnms{Alfred} \snm{Winter}\ead[label=e6]{alfred.winter@imise.uni-leipzig.de}}
\runningauthor{}
\address[A]{Institute for Medical Informatics, Statistics and Epidemiology (IMISE),
\orgname{University of Leipzig}, \cny{Germany}\printead[presep={}}}]{e1,e2,e3,e4,e5,e6}}
%Medical Informatics, Management of Health Information Systemsi
%Härtelstraße 16--18, D-04107 Leipzig
E-mail: \{khoeffner,fjahn,alfred\}@imise.informatik.uni-leipzig.de}
\begin{abstract}
Abstract text.
\end{abstract}
\begin{keyword}
\kwd{Hospital Information Management}
\end{keyword}
\end{frontmatter}
%%%%%%%%%%% The article body starts:
\section{Introduction}\label{sec:introduction}
A W3C design issue~\cite{puttinggovernmentdataonline} motivates making government data available online as Linked Data for three reasons:
\enquote{1) Increasing citizen awareness of government functions to enable greater accountability; 2) Contributing valuable information about the world; and 3) Enabling the government, the country, and the world to function more efficiently.}
Increasing the transparency of government spending specifically is in high demand from the public. For instance, in the survey publication~\cite{localtransparency}, \enquote{Public access to records is crucial to the functioning government} was rated with a mean of 4.14 (1 = disagree completely, 5 =  agree completely).
Open spending data can reduce corruption by increasing accountability and strengthening democracy because voters can make better informed decisions.
Furthermore, an informed and trusting public also strengthens the government itself because it is more likely to commit to large projects (see~\cite{fiscaltransparency} for details).

%Both effects are proven in a study~\cite{fiscaltransparency}.
Several States and Unions are bound to financial transparency by law, such as the European Union\footnote{\enquote{2. The Commission shall make available, in an appropriate
and timely manner, information on recipients, as well as the nature and purpose of the measure financed from the budget[...]}~\cite{euregulations}} with its \emph{Financial Transparency System (FTS)}\footnote{\url{http://ec.europa.eu/budget/fts}}~\cite{martin-fts}.
Public spending services satisfy basic information needs, but in their current form they do not allow queries which go further than simple keyword search or which cannot be answered with data from one system alone.
Linked Data solves those problems by providing a unified format, a powerful query language and the possibility of integration with linked data sets from other services.

Our contribution is an RDF transformation of the OpenSpending\footnote{\url{http://openspending.org}} project which provides government spending financial transactions from all over the world and is thus suitable as a core knowledge base that can be enriched and integrated with other, more focused data sets.
Transforming OpenSpending to Linked Data and publishing it adds to and profits from the Semantic Web which offers benefits including a standardized interface, easier data integration and complex queries over multiple knowledge bases.

%Some advantages of an RDF version over a relational database are (1) querying over multiple data sets at once using interlinks and (2) an additional custom vocabulary for each of the datasets, which are very heterogeneous (see \autoref{tab:sparqlqueries} for SPARQL queries for common use cases).

 %linkage provides benefits blablabla ...
%\todo{more motivation why RDF is good users are able to get this and that, what is linked, e.g. financial transparency system, cordis -> big benefit at least it is possible to link to this and that list other data sets, such as ... e.g. the paper about greece}

%We present LinkedSpending, which provides approximately two million financial transactions from all over the world extracted from OpenSpending as \emph{Linked Open Data}.
%information needs such as browsing of data sets, visualization and simple queries but they suffer from

The structure of the paper is as follows.
\Cref{sec:motivation} motivates the work and presents use cases.
\Cref{sec:openspending} describes OpenSpending, which is the source of the data, and its statistical data model.
\Cref{sec:conversion} explains the target RDF Data Cube vocabulary and the transformation process to it.
\Cref{sec:publishing} describes, how and where the data set is published and in which way users can access the data.
%\Cref{sec:overview} gives an overview over the transformed linked dataset, describes the its schema and shows typical SPARQL queries.
\Cref{sec:overview} gives an overall view of the data sets, gives details about the licence used and describes the data sets it is interlinked to.
%This is followed by a description of usage scenarios in \Cref{sec:usagescenarios}.
\Cref{sec:relatedwork} presents related spending data sets as Linked Open Data (LOD).
%The data quality is analyzed in \cref{sec:dataquality} and
The last section discusses known shortcomings of the data sets and future work. The prefixes used throughout this publication are defined in \Cref{tab:namespaces}.
In order to save space, prefixes are used even when technically incorrect, such as in \resource{ls:berlin\_de/model}.

\begin{table}[hbtp]
\caption{Namespaces and prefixes used in the paper}
\label{tab:namespaces}
%\tiny
\scriptsize
%\footnotesize
\renewcommand{\tabcolsep}{5pt}
\begin{tabularx}{\columnwidth}{
   >{\hsize=.25\hsize}L 
   >{\hsize=1.75\hsize}L 
}
\toprule
\textbf{prefix}	&\textbf{URL}\\
\midrule
os			&\os\\
owl			&\url{http://www.w3.org/2002/07/owl\#}\\
ls			&\ls\\
lso			&\lso\\
qb			&\url{http://purl.org/linked-data/cube\#}\\
sdmxd			&\url{http://purl.org/linked-data/sdmx/2009/dimension\#}\\
dbpedia			&\url{http://dbpedia.org/resource/}\\
dbp			&\url{http://dbpedia.org/property/}\\
\bottomrule
\end{tabularx}
\vspace{-20pt}
\end{table}

%while \cref{sec:transformation} explains the Data Cube model and the transformation.
%-The purpose of LinkedSpending is explained in \cref{sec:purposeoflinkedspending}.

%While the OpenSpending API is enough to satisfy basic information needs such as browsing of data sets, visualization and simple queries, Linked Data provides many more benefits such as complex queries and the possible integration of other datasets such as the FTS and
%OpenSpending.org is a an open platform that provides public finance data from governments around the world.
%At the time of this writing,

%We converted this data to RDF Data Cubes using a Java program which can be called again should data be added or existing data updated.

\section{Motivation}\label{sec:motivation}
%\section{Usage Scenarios}\label{sec:usagescenarios}
In a time of globalization,
%in the finance sector
financial data becomes an international network.
RDF data with its linked nature supports a representation that takes this network nature into account.
As a machine interpretable format, it lowers the access barrier for application developers.
For instance, generic Linked Data tools such as OntoWiki, CubeViz and Facete provide end users with the means to explore the data and discover new insights.

%\textbf{Community Empowerment:}
%network effect 
%reduce corruptions
% corruption by increasing accountability and strengthening democracy because voters can make better informed decisions.
%Furthermore, an informed and trusting public also strengthens the government itself because it is more likely to commit to large projects.
%hub

%domain specific tool on dataset

%generic tool

%\paragraph{Internationalization Support}

%Apart from the information needs that can already be satisfied using the source data, there are usage scenarios that are easier or only possible with LinkedSpending.

%\subsection{Economic Analysis}
\paragraph{Economic Analysis}
LinkedSpending is represented in Linked Open Data, which facilitates data integration.
Currencies from DBpedia and countries from LinkedGeoData are already integrated.
Financial data offers further integration candidates, such as political or other statistical, policy-influencing data such as health care.
This allows queries such as query 7 in \autoref{tab:sparqlqueries}, which asks for data sets with currencies whose inflation rates are greater than \SI{10}{\percent}.
%\todo{Ev. noch ein weiteres Beispiel um das noch zu untermauern und etwas mehr Substanz in die Subsection zu bringen - auch im Vergleich zu den folgenden beiden Subsections.}.

LinkedSpending can also be used to compute economic indicators across several data sets.
A possible indicator is a country's spending on education per person where the population size can be taken from the LinkedGeoData countries linked from one or more budget data sets.
One such data set is \dataset{ugandabudget}, which contains the Uganda Budget and Aid to Uganda, 2003--2006.
LinkedSpending serves as a hub for the integration of those data sets and their provenance information.
More data sets can be integrated with similarity-based interlinking tools such as LIMES~\cite{NGON11}.
%\footnotetext{and its web interface SAIM\cite{Lyko2013}}%, available at \url{http://saim.aksw.org}}

%\subsection{Putting Spending Amounts into Perspective}
%\subsection{Finding and Comparing Relevant Datasets}
\paragraph{Finding and Comparing Relevant Data Sets}

Government spending amounts are often much higher than the sums ordinary people are used to dealing with but even for policy makers it is hard to understand whether a certain amount of money spent is too high or normal.
Comparing data sets and finding those which are similar to another one helps separating common values from outliers which should be further investigated.
For example, if another country has a similar budget structure but spends way less on health care with a similar health level,
it should be investigated whether that discrepancy is caused by inherent differences such as different minimum wages or a different climate or if it is due to preventable factors such as inefficiencies or corruption.
While OpenSpending provides several hundreds of data sets which can be searched and it allows browsing and visualization of any single one, it does not provide a comparison function between data sets.
Because of the mechanism to identify equivalent properties (see \autoref{sec:conversion}), SPARQL queries can compare different data sets, e.g.~between similar structures in different countries.
Query 9 in \autoref{tab:sparqlqueries} shows a simple query to detect data sets which are most similar to any particular data set.
This is done by calculating the number of common measures, attributes and dimensions.
\iffalse
\subsection{Increased E-Government Transparency}
While OpenSpending makes government spending data available, the data can be made more understandable by providing additional knowledge about the described entities and by facilitating access to the data by non-expert users.
\paragraph{Additional Knowledge}
Through integration with other knowledge bases, users can for example learn about the country and currency a data set refers to at the same time as they query the data. \todo{Scheint mir nicht ideal formuliert. Während des Queries etwas lernen, kann man höchstens wenn der Endpoint sehr langsam ist. Ev. "while they browse and explore the data", falls das gemeint ist? Es wäre auch gut, wenn 7.1 und 7.3 nicht beide identisch Integration von country und currency erwähnen, da es sonst so wirkt, als ob es nicht viele andere Möglichkeiten gibt und man versucht den gleichen Punkt mehrfach an die Reviewer zu verkaufen.}
This helps interpreting the values and constructing helpful indicators. For example, the budget countries spend on health care varies wildly but less so when it is divided by the population number.

\todo{ich habe das qa jetzt rausgenommen da mir kein besonders guter use case eingefallen ist und die aus den example sparql queries zu kompliziert wären für eine erfolgreiche beantwortung}
\fi
\iffalse
\paragraph{Question Answering}
The OpenSpending data is available through numerous interfaces such as a JSON API, a keyword search and graphical visualizations.
Linked Open Data however allows the additional usage of Question Answering applications in order to to query the data in a way which is both more intuitive for non-experts than SQL and more expressive than keyword queries, which suffer from the inability to express relations and context, or faceted search, which only allows certain types of search restrictions.
Question Answering approaches, such as TBSL~\cite{unger2012template}, can thus be adapted to the spending domain and then allow not only computer science experts, but journalists or just ordinary people to know more about it more intuitively using natural, full-text sentences.
\todo{Der Paragraph sollte noch etwas angepasst werden um zu zeigen, warum Question Answering gerade hier günstig ist, z.B. eine konkrete Abfrage oder man probiert tatsächlich TBSL mit den Daten aus. Aktuell liest sich der Paragraph eher wie eine Begründung warum Question Answering über RDF allgemein sinnvoll ist (was zweifelslos der Fall ist, aber die Relevanz für das Paper noch nicht begründet).}
\fi

%\section{Source for the Data and Purpose of the Transformation}\label{sec:sourceforthedataandtopiccoverage}
%\section{Purpose of LinkedSpending}\label{sec:purposeoflinkedspending}
% viellecht besser passende beispiele nehmen, also anstelle von Kontinenten Gruppen und Hauptgruppen nehmen
%For example the coded property \url{:continents} could have the range \url{:Afrika}, \url{:America}, \url{:Asia}, \url{:Australia} and the child property \url{:countries} ... (spezielle properties für jeden kontinent?)
%This parent-child relationship is different from rdfs:subClassOf because the values are disjunctive but instead the childrens values are parts of the parents values.

%grouped in slices data sets 
%structure defined by datastructure definitions

%At the time of this writing, OpenSpending contains 217 different data sets.
%While they share the general \emph{budget} domain, they still have slightly different contexts and are described in different vocabularies.

\iffalse
2-3 beispiele

statistik darüber bringen

manuelles proof über die qualität

am ende die datacubes nehmen

data quality dashboard vom data cube nehmen
\fi

\section{OpenSpending Source Data}\label{sec:openspending}
OpenSpending\footnote{\url{http://openspending.org/}} is a project which aims to track and analyze public spending worldwide and, at May 2014, contains more than 25 million financial entries in \numberofopenspendingdatasets{} data sets\footnote{As some of them contain errors, the number of LinkedSpending data sets is slightly smaller.}.
Data Sets can be submitted and modified by anyone but they have to pass a sanity check from the OpenSpending Data Team which also cleans the data before publishing.\footnote{\url{http://community.openspending.org/contribute/data/}}
OpenSpending hosts transactional as well as budgetary data with a focus on government finance.\footnote{\url{http://community.openspending.org/help/guide/en/financial-data-types/}}
It contains this data in structured form stored in database tables and provides searching and filtering as well as visualizations and a JSON REST interface.
The data sets differ in granularity and type of accompanying information, but they share the same meta model.
%, e.g. from the FTS or CORDIS\footnotemark.
\subsection{The Data Cube Model}%\label{sec:rdfdatacubemodel}
The domain model of OpenSpending is a \emph{data cube} (also \emph{OLAP cube}, \emph{hypercube}), which represents multi-dimensional statistical observations.
Each cell corresponds to an observation (an instance of spending or revenue) that contains measurements (e.g. the amount of money spent or received).
The context of the measurement is provided by the \emph{dimensions} like the purpose, department and time of a spending item and optionally by \emph{attributes}, which further describe the measured value, e.g., the unit of the measurement.
%Apart from the fixed data cube meta model, the structure of each dataset is completely up to the creator.
%The budget data sets for example have the dimensions \emph{time} and \emph{department}, the measure \emph{amount spent} and the attribute \emph{currency} with the value \emph{euro}.
\begin{figure}[h]
\footnotesize
\begin{lstlisting}[language=json]
"sub-programme": {
 "label": "Sub-programme",
 "type": "compound",
},
"amount": {
    "datatype": "float",
    "label": "Total",
    "type": "measure",
}
\end{lstlisting}
\caption{simplified excerpt of an OpenSpending \emph{model}} 
\label{fig:jsonmodel}
\end{figure}

\begin{figure}[h]
{\footnotesize
\begin{lstlisting}[language=json]
"sub-programme": {
 "label":"Security and safeguarding liberties",
 "html_url":"http://openspending.org/eu-budget/sub-programme/security-and-safeguarding-liberties",
 "name":"security-and-safeguarding-liberties"
},
"html_url": "http://openspending.org/eu-budget/entries/017dfcb58d05671ef9eb5a9f77fef39c8b14150c",
"amount": 41.2
\end{lstlisting}
}
\iffalse
"main-programme": {

    "description": null,
    "label": "Main programme",
    "facet": true,
    "key": true,
    "attributes": {
        "name": {
            "column": "StringCAT1",
            "datatype": "id",
            "default_value": ""
        },
        "label": {
            "column": "StringCAT1",
            "datatype": "string",
            "default_value": ""
        }
    },
    "type": "compound",
    "dimension": "main-programme"
},
"main-programme": {
 "label":"Citizenship, Freedom ,Security and Justice",
 "html_url":"http://openspending.org/eu-budget/main-programme/citizenship-freedom-security-and-justice",
 "name":"citizenship-freedom-security-and-justice"
},

\fi
\caption{simplified excerpt from an OpenSpending \emph{entry}}
\label{fig:jsonentry}
\end{figure}

\Cref{fig:jsonmodel} shows an excerpt from the model of the OpenSpending data set \emph{eu-budget} with the dimension \emph{sub-programme} and the measure \emph{amount}.
\Cref{fig:jsonentry} shows the an entry that contains the actual values for the dimension and the measure of the observation.

\subsection{Problems}
While the data is well-structured and thus suitable for conversion without data cleaning or extensive preprocessing, it still poses problems that need to be taken into account:
\begin{inparaenum}
\item New data sets are frequently added (approximately 50 per month) and, less often, existing data sets are modified.
\item Some data sets do not specify a value for all properties in all observations.
\item There are properties with the same name in different data sets where it is unknown if they specify the same property.
\item Data Cube is a meta model. The deep structure of the data sets is heterogeneous and
 described only shallowly.
\item The language of literals is varying between and even within data sets but the language used is not specified.
\end{inparaenum}
Points 1 to 3 are addressed in the next section while points 4 and 5 are discussed in \cref{sec:conclusions}.

\section{Conversion of OpenSpending to RDF}\label{sec:conversion}

%The conversion from OpenSpending to Linked Data consists of the extraction of the source data and its transformation to RDF.
%Because the source data adheres to the data cube model, a conversion of the data to RDF needs an appropriate RDF vocabulary.
%The RDF DataCube vocabulary~\cite{rdfdatacube} is an ideal fit for the transformed data.
%Because the source data uses the data cube model, the RDF DataCube vocabulary~\cite{rdfdatacube} is an ideal fit for the transformed data.

\begin{table*}[hbtp]
\caption{Conversion of OpenSpending to LinkedSpending classes and instances}
\label{tab:transformation}
\scriptsize
\renewcommand{\tabcolsep}{5pt}
\begin{tabularx}{\textwidth}{
% sum of hsizes needs to be 2
   >{\hsize=.1\hsize}L 
   >{\hsize=.4\hsize}L 
   >{\hsize=.8\hsize}L 
   >{\hsize=.6\hsize}L 
   >{\hsize=.3\hsize}L          
}
\toprule % a bit unelegant: with the replacement of \url by \urlX to circumvent hyperref the line break inside urlX does not work so there are 2 spaces
&\textbf{Source URL}  		&\textbf{JSON Path}				&\textbf{LinkedSpending class}		&\textbf{LS instance scheme}\\ 
\midrule
\rom{1}  		&\osdatasetjson 		&						&\class{qb:DataSet}			&\lsdataset\\
\rom{2}			&\osmodel 			&						&\class{qb:DataStructureDefinition}	&\dsd\\
\midrule
\multirow{3}*{\rom{3}} 	&\osmodel			&\texttt{\$.mapping.*}				&\class{os:\{Country,Time\}Component} \class{Specification}
													or \class{qb:ComponentSpecification}	&\urlX{lso:}\emph{propertyname} \urlX{-spec}\\
\midrule
\rom{4}			&\multirow{4}*{\osmodel}	&\texttt{\$.mapping.*[?(@.type="compound")]}	&\class{qb:DimensionProperty}		&\multirow{4}*{\urlX{lso:}\emph{propertyname}}\\
\rom{5}			&				&\texttt{\$.mapping.*[?(@.type="date")]}		&\class{qb:DimensionProperty}	&\\
\rom{6}			&				&\texttt{\$.mapping.*[?(@.type="measure")]}	&\class{qb:MeasureProperty}		&\\
\rom{7}			&				&\texttt{\$.mapping.*[?(@.type="attribute")]}	&\class{qb:AttributeProperty}		&\\
\midrule
\rom{8} 		&\osdataset\urlX{/entries.json} 		&\texttt{\$.results[*].dataset}	&\class{qb:Observation}				&\urlX{ls:}\urlX{observation-} \emph{dataset name}-\emph{hashvalue}\\
\bottomrule
\end{tabularx}
\end{table*}
%\iffalse
%\begin{sidewaystable}
%{\setlength{\tabcolsep}{0.5em}
%\scriptsize
%\begin{tabulary}{0.84\textwidth}{lllLp{3cm}p{3cm}}
%\toprule
%			&\textbf{Source URL}  		&\textbf{JSON Path}				&\textbf{LinkedSpending class}		&\textbf{OpenSpending URL}				&\textbf{LinkedSpending URI}\\ 
%\midrule
%\rom{1}  		&\osdatasetjson 		&						&\class{qb:DataSet}			&\osdataset						&\lsdataset\\
%\rom{2}			&\osmodel 			&						&\class{qb:DataStructureDefinition}	&\osmodel						&\dsd\\
%\midrule
%\multirow{3}*{\rom{3}} 	&\osmodel			&\texttt{\$.mapping.*}				&\class{os:{Country,Time}ComponentSpecification}
%													or \class{qb:ComponentSpecification}	&none							&\lsdataset\url{/}\emph{propertyname} \url{/component} \url{Specification}		\\
%\midrule
%\rom{4}			&\multirow{4}*{\osmodel}	&\texttt{\$.mapping.*[?(@.type="compound")]}	&\class{qb:DimensionProperty}		&\multirow{4}*{\osdataset\url{/}\emph{propertyname}}	&\multirow{4}*{\lsdataset\url{/}\emph{propertyname}}\\
%\rom{5}			&				&\texttt{\$.mapping.*[?(@.type="date")]}		&\class{qb:DimensionProperty}	&							&\\
%\rom{6}			&				&\texttt{\$.mapping.*[?(@.type="measure")]}	&\class{qb:MeasureProperty}		&							&\\
%\rom{7}			&				&\texttt{\$.mapping.*[?(@.type="attribute")]}	&\class{qb:AttributeProperty}		&							&\\
%\midrule
%\rom{8} 		&\osdataset\url{/entries.json} 		&\texttt{\$.results[*].dataset}	&\class{qb:Observation}			&\osmodel\url{/entries/} \emph{hashvalue}			&\lsdataset\url{/model/entries/}\emph{hashvalue}\\
%\bottomrule
%\end{tabulary}
%\caption{Transformation of OpenSpending entities to LinkedSpending classes}}
%\label{tab:transformation}
%\end{sidewaystable}
%\fi
%\todo{\autoref{tab:transformation} source uris take the ones where i get them from or where the source uri is that only describes it? should be both acutally}

%\paragraph{The RDF DataCube vocabulary}
\paragraph{}
The RDF DataCube vocabulary~\cite{rdfdatacube}, i.e.~an RDF variant of the previously explained data cube model, is an ideal fit for the transformed data. 
\begin{figure}[h]
\includegraphics[width=\columnwidth]{img/rdfdatacube_shrinked.pdf}
\caption{Used RDF DataCube concepts and their relationships\protect\footnotemark}
\label{fig:rdfdatacube}
\end{figure}
\footnotetext{Simplified version of the structure described in ~\cite{rdfdatacube}.}

First and foremost, this vocabulary provides the backbone structure for every LinkedSpending data set, see \cref{fig:rdfdatacube}.
Each data set is represented by an instance of \class{qb:DataSet} and an associated instance of \class{qb:DataStructureDefinition} which includes \emph{component specifications} (see \cref{fig:modellingexample} for an example).
Each component specification is associated to a \emph{component property} which can be either a \emph{dimension}, an \emph{attribute} or a \emph{measure}.
Commonly used concepts are specified in the model of the \emph{Statistical Data and Metadata eXchange (SDMX)} initiative\footnote{\url{http://sdmx.org}}.
%\Cref{fig:modellingexample} shows, how
%example, a reference area can make use of \resource{sdmxa:refArea}.
The RDF Data Cube vocabulary  is supported by the LOD2 Statistical Office Workbench\footnote{\url{http://demo.lod2.eu/lod2statworkbench}} which is part of the Linked Data Stack (an advanced version of the the LOD2 Stack~\cite{Auer+ISWC-2012}).
The workbench includes a DataCube validator, a split and merge component and a CKAN Publisher.
The OntoWiki~\cite{ontowiki_www}, which manages several parts of the the Linked Data Lifecycle~\cite{Auer+ISWC-2012}, such as Storage/Querying and Search/Browsing/Exploration offers a CSV import plugin for the format as well as a faceted RDF Data Cube browser, CubeViz.
Data cubes may contain slices, which are presets for certain dimension values, effectively selecting a subset of a cube. Users may create and visualize their own slices using the OntoWiki CubeViz plugin. Furthermore, the RDF DataCube vocabulary allows the persistence of slices which is used to represent preconfigured slices from OpenSpending.

% the Linked Data Stack~\cite{Auer+ISWC-2012}
% which others? ORE - Ontology Repair and Enrichment ?

\begin{figure}
\footnotesize
\begin{lstlisting}[language=ttl]
ls:berlin_de
 rdf:type	qb:DataSet;
 rdfs:label	"Berlin Budget";
 dc:source	os:berlin_de;
 qb:structure	ls:berlin_de/model;
 qb:slice	ls:berlin_de/views/nach-einzelplan.

ls:berlin_de/model
 rdf:type qb:DataStructureDefinition;
 qb:component
   lso:CountryComponentSpecification,
   lso:DateComponentSpecification,
   lso:Einzelplan-spec,
   lso:amount-spec.

lso:CountryComponentSpecification 
 rdf:type qb:ComponentSpecification;
 rdfs:label "country";
 qb:attribute sdmxa:refArea; 
 qb:componentRequired "false"^^xsd:boolean;
 qb:componentAttachment
   qb:DataSet,qb:Observation.
\end{lstlisting}
\caption{RDF DataCube vocabulary modelling excerpt of data set \resource{berlin\_de} (some properties and values omitted).}
\label{fig:modellingexample}
\end{figure}

\paragraph{Transformation}
All of the OpenSpending data sets describe observations referring to a specific point or period in time and thus undergo only minor changes.
New data sets however, are frequently added. Because of this, the huge number of data sets and their size, an automatic, repeatable transformation is required.
This is realized by a program\footnote{written in Java, available as open source at \url{https://github.com/AKSW/openspending2rdf}} which fetches a list of data sets on execution and only transforms the ones who are not transformed yet. Each data set is transformed separately.
\autoref{tab:transformation} shows for each class used by LinkedSpending, at which URL (abbreviated using the prefixes from \cref{tab:namespaces}) the information used to create the instances of those classes is found. % cref somehow doesnt work with this table
In case there are multiple instances described at one URL, a \emph{JSON path}\footnotemark{} expression is given, that locates the corresponding subnodes.
\footnotetext{\emph{JSON path} (\url{http://code.google.com/p/json-path/}) is a a query language for selecting nodes from a JSON documents, similar to XPath for XML}
Finally, the table contains the patterns that describe resulting LinkedSpending URLs.
For example, the OpenSpending URL \resource{os:berlin\_de/model} contains the node \texttt{\$.mapping.amount} which has a type value of \enquote{attribute} and is, thus, transformed to the OpenSpending instance \resource{lso:amount} of the class \resource{qb:AttributeProperty}.

Equivalent component properties (dimensions, attributes and measures) are identified as follows:
A configuration file optionally specifies the mapping of data set and property name to an entity in the LinkedSpending ontology.
By default, the property URI is derived from the property name.
Properties with the same name in different data sets not having a mapping entry that states otherwise are assumed to represent the same concept and thus given the same URL.\footnotemark
\footnotetext{Although that has the possibility of mismatches, such a mismatch has not been spotted yet. Still, evaluating and, if necessary, improving the automatic matching is part of future work.}

%\subsection{Use of Established Vocabularies}
\paragraph{Use of Established Vocabularies}
%\textbf{Use of Established Vocabularies:} 
In addition to the standard vocabularies, RDF, RDFS, OWL and XSD, the DCMI vocabulary is used for source and generation time metadata.
The data sets are modelled, first and foremost, according to the RDF Data Cube vocabulary, % (see \cref{sec:openspending}),
which specifies the structure of a data cube.
LinkedSpending follows the RDF Data Cube recommendation to make heavy use of the SDMX model for measures, attributes and dimensions.%, among others, sponsored by the United Nations Statistics Division, the International Monetary Fund and the World Bank.
The data sets are very heterogeneous but there are some properties which are commonly specified and thus modelled with established vocabularies.
The year and date, a data set and an observation refers to, respectively, is expressed by \resource{sdmx-dimension:refPeriod} and XSD.
%other than the usage of time and dates, which are expressed  currencies and countries.

Currencies are taken from DBpedia~\cite{dbpedia-swj} and countries are represented using the vocabulary of LinkedGeoData~\cite{SLHA11}, a hub for spatial linked data.
Some amount of data is imported from LinkedGeoData countries and DBpedia currencies. Because of the limited number of countries and currencies, and properties values imported per country and currency, the amount of data is too small to consider federated querying.
As most countries and currencies are stable in the medium term, this data needs to be updated only infrequently.
\paragraph{Interlinking}
%\textbf{Interlinking:}
There are two possibilities to align entities to another vocabulary: 1) to use the entities directly and 2) to create an own RDF resource with interlinks, like \resource{owl:sameAs}, to that vocabulary.
We generally preferred the first approach because a higher amount of reuse provides easier integration, better understandability and tool support.
%We generally preferred to use entities of a vocabularies directly, such as the currencies and countries, because this is a more elegant and less convoluted approach as long as the used vocabularies are stable and exactly correspond to the required meaning.
%, as long as the entities are always in the object, or property position.
% wir brauchen nicht extra begründen warum man subjects verlinken muss um dann zu sagen, dass es sowieso keine observations gibt, die man verknüpfen kann (da fragt sich der leser warum wir das schreiben)
%Using an entity of another vocabulary in the subject position is problematic, however, because, amongst other things, those additional triples are not reachable by standard Linked Data URI dereferencing.
%Because of this, the interlinks between observations of the data sets to observations in other data cubes need to be aligned by interlinking and not by replacement.
% Observations are only equal to others however, when they describe exactly the same transaction and overlapping data sets from other sources have not been found.
% While there are usually no interlinking targets for \emph{sameAs} links to other observations, there are many possibilities for interlinks between data sets or dimension values (which are always URIs, mostly with labels) and concepts they refer to.
While we did not find \emph{sameAs} link targets on observation level, i.e.~exactly the same statistical observations described in other data sets, there are many possibilities for interlinks between data sets or dimension values and concepts they refer to.
Using the labels of those data sets and dimension values, it is possible, for example, to link values of the dimension \enquote{region} of a federal budget, and thus indirectly also the observations which use those values, to the cities in DBpedia or LinkedGeoData whose labels are contained in the label of the region value URI.

%\todo{Wenn zwei Datensätze sich z.B. auf Städte beziehen, könnte man vermutlich schon mehr Sachen verknüpfen. Das Problem bei dem Absatz ist, dass Du hier sagst, dass es quasi keine Linkmöglichkeiten gibt, aber unter 7.1 betonst, dass man mit mehr Links noch mehr Analysen durchführen kann.}

%, such as \ls\url{/berlin\_de}, 
%, can browse single URLs as HTML through the Pubby\footnote{\url{http://wifo5-03.informatik.uni-mannheim.de/pubby/}} interface by dereferencing the URIs, such as \ls\url{/berlin\_de}, in a browser.

%The user can browse single URLs as HTML through the Pubby\footnote{\url{http://wifo5-03.informatik.uni-mannheim.de/pubby/}} interface by dereferencing the URIs, such as \ls\url{/berlin\_de}, in a browser.

%Because mass processing such as interlinking requires computationally expensive queries against the SPARQL endpoint, an RDF dump\footnote{\ntriples} is also available. \todo{Satz würde ich eher weglassen bzw. verschieben. Es ist recht klar, dass es einen Dump gibt und es passt nicht in die Section.}
%\Cref{tab:technicaldetails}

\paragraph{Error Handling}
The OpenSpending API lists \numberofopenspendingdatasets{} data sets with \numberofdatasets{} of them having a LinkedSpending equivalent.
The discrepancy is caused by loss in several stages.
To prevent timeouts and to reduce the impact of disrupted connections, the source data set is downloaded in several parts with a maximum number of entries.
These parts are then merged so that each file corresponds to exactly one data set.
Data sets without observations are removed and the remaining data sets are transformed, noting the missing values for all component properties.
If the first 1000 values are all missing, the transformation is aborted, otherwise a \resource{lso:completeness} value $c = \frac{|\textnormal{existing values}|}{|\textnormal{observations}||\textnormal{component properties}|}$ is attached to the data set.
Besides empty or non-existing data sets, there were no other types of error observed.
%here are however several cases of component properties with the same name which raises the problem of determining equivalent component properties. The chosen approach is to regard as equal all properties with exactly the same name.
%A%ll data sets, with the exception of ones where the first 1000 expected values are all missing, are now fully converted and each dataset now has a property “” which has an assigned float value measuring 1-%(missing number of values)/(expected number of values).

%In order to detect faulty data sets and to guarantee high quality data, there is a threshold on the number of errors.
%If the error count of the transformation of a dataset reaches both at least 30 and \SI{10}{\%} of all observation-property pairs yet processed, the transformation is aborted, which removes 5 additional data sets, leaving \numberofdatasets{}.

\paragraph{Sustainability}
The data conversion process is controlled by a web application\footnotemark{}, which constantly checks for added and modified data sets from OpenSpending, which are automatically queued for conversion but can also be manually managed.
%\footnotetext{Available at \url{http://linkedspending.aksw.org/api}, developed at \github{}}
\footnotetext{\url{http://linkedspending.aksw.org/api}}
Updates don't interrupt the accessibility of the SPARQL endpoint and the services building on it.
On average, about 50 new data sets became available on each month between September 2013 and March 2014.
A service monitor constantly checks the state of the application and reports errors.

%thus the OntoWiki and CubeViz instances. % würde ich weglassen, da OntoWiki+CubeViz erst im nächsten Kapitel richtig erklärt werden

%(\todo{warten auf publikation von michas paper} \remark{JL: Ich kenne den Status nicht, aber auf andere Publikationen warten, ist selten eine gute Idee.})
%This facilitates the consumation of the data sets.

%Observations can also aggregate measurements and thus allow views of different detail level over the same base data.
%The dimensions are either data type properties, like \enquote{xsd:dateTime} or coded properties which have \emph{code lists} detailing their possible values.
%The values can be organized in hierarchical code lists, where each parent property (and code list) aggregates the values of its children.

\paragraph{Performance}
The transformation of a data set takes less than an hour on average on a 2 GHz virtual machine, using less then 2 GB of RAM.

\section{Publishing}\label{sec:publishing}
The data is published using OntoWiki~\cite{ontowiki_www}.
The interface for human and machine consumption is available at \url{http://linkedspending.aksw.org}.
Depending on the actor and needs, OntoWiki provides various abilities to gather the published RDF data.
\begin{figure}[b!]
\includegraphics[width=\columnwidth]{img/ontowiki2.png}
\caption{View of the data set \dataset{berlin\_de} in the OntoWiki}
\label{fig:ontowiki}
\end{figure}
%The data can be explored and (with the appropriate access rights) authored collaboratively \todo{Die Daten kann man zwar editieren, aber sie werden beim nächsten Durchlauf überschrieben, oder?} (see \autoref{fig:ontowiki}).
It can be explored by viewing the properties of a resource, its values and by following links to other resources (see \autoref{fig:ontowiki}).
Using the SPARQL endpoint\footnote{\sparql} provided by the underlying \emph{Virtuoso Triple Store}\footnote{\url{http://virtuoso.openlinksw.com}}, actors are able to satisfy complex information needs.
%\footnote{According to benchmarks~\cite{Bizer2009, MOR+11, MOR+12} Virtuoso  }
\begin{figure}[t!]
\includegraphics[width=\columnwidth]{img/cubeviz_circles.pdf}
\caption{Faceted browsing in CubeViz by restricting values of dimensions}
\label{fig:cubeviz_circles}
\end{figure}

Faceted search offers a selection of values for certain properties and thus slice and dice of the data set according to the interests on the fly.
For example, depicted in \autoref{fig:cubeviz_circles} is all Greek police spending in a certain region.
Visualization supports discovery of underlying patterns and gain of new insights about the data, for example about the relative proportions of a budget (see \autoref{fig:cubeviz_piechart}).
%As can be seen from both depictions, exploration of statistical data using a vocabulary-adopted tool is more efficient and comfortable. \todo{"more efficient" im Vergleich zu was? Außerdem glaube ich kaum, dass man so etwas anhand von einem Screenshot beurteilen kann.}
We set up the RDF DataCube Browser CubeViz~\cite{salas-ijsc-2013} as part of the human consumption interface.
%Both exploration paradigms are already available on OpenSpending, however CubeViz offers data discovery by generating new instead of predefined visualization and takes advantage of the openness of the data by publishing additional information from interlinked data sets.

\begin{table}[hbtp]
\caption{Technical details of the LinkedSpending data set}
\label{tab:technicaldetails}
\tiny
\renewcommand{\tabcolsep}{5pt}
\begin{tabularx}{\columnwidth}{
   >{\hsize=.7\hsize}L 
   >{\hsize=1.3\hsize}L 
}
\toprule
URL				&\lswebsite\\
Version date and number		&2013-8-14, 0.1\\
                      		&2014-4-11, 2014-3\\
License				&PDDL 1.0\footnotemark\\
SPARQL endpoint			&\sparql\\
%VoiD description		&\url{ls:void}\\
%Schema				&\url{ls:ontology/}\\
Compressed N-Triples Dump 	&\ntriples\\
datahub entry			&\url{http://datahub.io/dataset/linkedspending}\\
\bottomrule
\end{tabularx}
\vspace{-10pt}
\end{table}
\footnotetext{\url{http://opendatacommons.org/licenses/pddl/1.0/}}

\paragraph{Licensing}
All published data is openly licensed under the PDDL 1.0. in accordance with the open definition\footnote{\url{http://opendefinition.org/}}.
%which is summarized by: \todo{Ist es wirklich notwendig die Summary hier zu bringen. Der Link reicht doch.} \blockquote{A piece of data or content is open if anyone is free to use, reuse, and redistribute it — subject only, at most, to the requirement to attribute and/or share-alike}.\footnote{\url{http://opendefinition.org/}}
%The license details are described in a \emph{VoID}~\iffalse ~\footnote{\url{http://www.w3.org/TR/void/}}\fi description

\section{Overview over the Data Sets}\label{sec:overview}
LinkedSpending consists of \numberofdatasets{} data sets (continually growing) with \numberofobservationsapproximate{} observations total.
The amount of observations of the individual data sets varies considerably between two (spendings in Prague of about \SI{5000}{CZK} for an unknown purpose) and \SI{242209} (\enquote{Spending from ministries under the Danish government}).
\Cref{tab:amountofdata} details the average and total amount of data in bytes, triples, and observations as well as the number of links to external data sets, which, for the presented version of 2014-3, amounts to more than 9 million links to LinkedGeoData countries and 1.5 million links to DBpedia currencies.\footnotemark
\footnotetext{The links are inflated as they originate in observations instead of data sets, which allows better querying and tool support.}
\Cref{fig:histogram} shows the distribution of the numbers of measures, attributes and dimensions of the data sets.\footnote{This analysis relates to version 0.1, which contains less data sets.}
Measures represent the quantity that an observation describes.
All data sets have at least one measure which is the amount of money spent or received.
For most of them (217) that is the only one but there are data sets with up to 7 measures.
%\footnote{\dataset{dk-corporate-tax-list} contains measures for different kinds of profits as well as losses.}.
Attributes give further context to the measurement.
The number of attributes is more varied, ranging from 2 to 26, with all data sets having at least a currency and a country, and most of them additionally the time the observations refer to.
While the number of dimensions ranges from none\footnote{There is only one data set with no dimensions which a test data set on OpenSpending, as a data cube with no dimensions is not useful.} to 32, almost all of the data sets have between 1 and 6 dimensions, 
the most common ones being the year and the time the data set and the observations refers to, respectively.
Technical details about the data sets are described in \cref{tab:technicaldetails}.

\begin{figure}[h]
\includegraphics[width=0.9\columnwidth]{img/histogram.pdf}
\caption{Histogram of measures, attributes and dimensions (version 0.1). \num{217} data sets have exactly one measure (clipped bar).}
\label{fig:histogram}
\end{figure}

\begin{table}[hbtp]
\caption{Amount of data for version 2014-3. All values are rounded to the nearest integer.}
\label{tab:amountofdata}
\tiny
\renewcommand{\tabcolsep}{5pt}
\begin{tabularx}{\columnwidth}{
   >{\hsize=\hsize}L 
   >{\hsize=\hsize}L 
   >{\hsize=\hsize}L 
}
\toprule	
	&\textbf{Total}    		&\textbf{Average}\\ 
\midrule
\textbf{number of data sets}    		&\numberofdatasets{}\\
\textbf{file size (RDF/N-Triples)}	&\SI{24585}{MB}		&\SI{39}{MB}\\
\textbf{triples}     		    &\num{113640534}&	\num{181245}\\
\textbf{observations} 			& \num{5026393}		&\num{8017}\\
\textbf{links to external data sets} 			&\num{10696614}		&\num{17060}\\
% OLD VERSION
%\textbf{filesize (RDF/N-Triples)}	&\SI{10012}{MB}		&\SI{41}{MB}\\
%\textbf{triples}     			&\num{49619572}&	\num{200889}\\
%\textbf{observations} 			&\num{2404313}		&\num{9734}\\
%\textbf{links to external data sets} 			&\num{}		&\num{}\\
\bottomrule
\end{tabularx}
\end{table}
%\subsection{Example Queries}
\paragraph{Example Queries} 
%\textbf{Example Queries:} 
\Cref{tab:sparqlqueries} contains queries for common use cases:
Queries 1--6 are basic queries.
Query 7 uses the interlinking to DBpedia currencies by querying over two different graphs.\footnotemark{}
\footnotetext{Parts of DBpedia and LinkedGeoData describing countries and currencies have been integrated in the SPARQL endpoint. With federated querying however, nearly the whole LOD cloud can be queried.}
Query 8 uses the custom vocabulary\footnotemark{} which is available for each data set.
\footnotetext{In this case, the \enquote{Hauptfunktion} and \enquote{Oberfunktion} are unique to the \dataset{berlin\_de} dataset.}

\begin{table*}[t]
\lstset{aboveskip=-8pt,belowskip=-\baselineskip}

\scriptsize
\caption{Examplary SPARQL queries for typical use cases.}
\label{tab:sparqlqueries}
\begin{tabular}{@{\makebox[1em][r]{\rownumber\space}}p{0.2\textwidth}p{0.8\textwidth}}
\toprule
\textbf{information need}	&\textbf{SPARQL Query}\\
\midrule
list of all data sets&\begin{lstlisting}
select * {?d a qb:DataSet}
\end{lstlisting}\\
all measures of the dataset \dataset{berlin\_de}
&\begin{lstlisting}
select ?m { ls:berlin_de qb:structure ?s. ?s qb:component ?c. ?c qb:measure ?m.}
\end{lstlisting}\\
all years which have observations in the de-bund dataset from \num{2020} onwards&
\begin{lstlisting}
select distinct ?year {?o a qb:Observation. ?o qb:dataSet ls:de-bund. ?o lso:refYear ?year.FILTER (xsd:date(?year) >= "2020-1-1"^^xsd:date) }
\end{lstlisting}\\
spendings of more than \num{100} billion \euro&
\begin{lstlisting}
select * {?o lso:amount ?a. ?o dbo:currency dbpedia:Euro. FILTER(xsd:integer(?a)>"1E11"^^xsd:integer) }
\end{lstlisting}\\
data sets with multiple years&
\begin{lstlisting}
select ?d count(?y) as ?count { ?d a qb:DataSet. ?d lso:refYear ?y. } group by ?d having (count(?y)>1)
\end{lstlisting}\\
sums of amounts for each reference year of \dataset{berlin\_de}&
\begin{lstlisting}
select ?y (sum(xsd:integer(?amount)) as ?sum) 
{?o qb:dataSet ls:berlin_de. ?o lso:refYear ?y. ?o lso:amount ?amount.} group by ?y
\end{lstlisting}\\
data sets with currencies whose inflation rate is greater than \SI{10}{\%}&
\begin{lstlisting}
select distinct ?d ?c ?r {?o qb:dataSet ?d. ?o  dbo:currency ?c. ?c dbp:inflationRate ?r. filter(?r > 10)} 
\end{lstlisting}\\
Berlin city subsectors of research and education that have had their budget reduced from 2012 to 2013 (data set version 0.1)&
\begin{lstlisting}
select ?l (sum(xsd:integer(?amount12)) as ?sum12) (sum(xsd:integer(?amount13)) as ?sum13)
 {?o qb:dataSet ls:berlin_de.
  ?o lso:Hauptfunktion <http://openspending.org/berlin_de/Hauptfunktion/1>.
  ?o lso:Oberfunktion ?of. ?of rdfs:label ?l.
  {?o lso:refYear "2012"^^xsd:gYear. ?o lso:amount ?amount12.} UNION
  {?o lso:refYear "2013"^^xsd:gYear. ?o lso:amount ?amount13.}}
group by ?l
having (sum(xsd:integer(?amount12)) > sum(xsd:integer(?amount13)))
\end{lstlisting}\\
data sets ordered by their number of properties in common with \dataset{2012\_tax} (having at least one such common property)&
\begin{lstlisting}
select ?d (count(?c) as ?count)
{ls:2012_tax qb:structure ?s. ?s qb:component  ?c.
 ?d qb:structure ?s2.         ?s2 qb:component ?c.
 FILTER(?d!=ls:2012_tax)}
group by ?d
order by desc(?count)
\end{lstlisting}\\
\bottomrule
\end{tabular}
\end{table*}

%mashups mehrwertgenerierung durch transformation

%bei openspending ist alles da und super aaaber:
\iffalse
\section{OpenSpending}
\begin{table}
\begin{tabulary}{\columnwidth}{lS}
\toprule
\textbf{category}    	&\textbf{number of data sets}\\ 
\midrule
$\sum$			&\\
\bottomrule
\end{tabulary}
\caption{Categories of data sets from OpenSpending}
\end{table}
\fi

\iffalse
The data sets are divided into the categories \emph{budget}, \emph{spending} and \emph{other} and are further organized by country and language.
Each data set has a model that contains its core data set metadata (e.g. name, label, description, currency) and the dimensions, measures and views.
Each of the models can be transformed to an RDF Data Cube \verb|qb:DataStructureDefinition| and each data set to a \verb|qb:DataSet|.
Because of the large number of data sets, the models are not transformed manually but with \todo{insert the method here}.
Model dimensions are mapped to Data Cube dimension, where each of the json object names is manually mapped to an RDF property.

data cube: dimensions, attributes, measures
\fi

%\end{thebibliography}

\iffalse
\subsection{Intrinsic dimensions}

\paragraph{Accuracy}
Because the data sets were created by transformation and the values itself were not changed, their accuracy is equal to that of the source data of each data set.
A request for information about existing studies into the accuracy on the OpenSpending mailing list was not answered.
To measure the accuracy, more elaborate research is needed that checks the correctness of a random sample of the values by comparing them with a reference.

\paragraph{Objectivity}
\todo{das klingt alles viel zu schwammig, aber es ist ja auch nicht klar messbar. wie kann man das trotzdem wissenschaftlich schreiben?}
As for the accuracy, the objectivity depends on the individual source data provided originally by government agencies and then converted and transfered to OpenSpending by some third party.
So, assuming that the values were not changed on the way from the agencies to OpenSpending, the amount of thrust that the information is unbiased depends on the thrustworthyness of the governements and agencies themselves.
Finally, objectivity is a subjective dimension and depends on the use case.
Using the data to get an overview of the financial situation of a country is surely unproblematic but to allocate funding based on the exact number of the reports is thought to be questionable by the author of the paper.
%- es gibt auf jeden fall anreize, finanzdaten zu fälschen ()
\paragraph{Validity-of-documents}
The validity of the RDF syntax has been successfully verified by the Raptor RDF parsing and serializing utility.
Valid use of vocabularies is verified by \todo{...}

\paragraph{Consistency}
The way the program is written guarantees that for each observation there is only value for each measure.\todo{doesnt sound good, how to improve?}

\paragraph{Conciseness}
In RDF Data Cubes, attributes can be attached to a data set, a slice, a measure property or to an observation.

While attributes which take different values for each observations necessarily have to be attached at the observation level, there are conflicting arguments for attributes which have the same value throughout.
Attaching them at the observations ensures that all relevant information concerning an observation is in one place.
Thus users can satisfy their information needs with less complex SPARQL queries or by resolving a single URL.
Attaching them at the data set however removes redundancy and thus improves the conciseness.
Furthermore, queries which select certain data sets, are actually easier to formulate with the latter approach.
However, it is easier to create mash-ups and to use RDF DataCube visualization tools such as CubeViz~\cite{cubeviz}, when all attributes are attached at the observation level, so that we attach all attributes directly to the observations.
RDF however is flexible and permits attaching the countries additionally at the data sets. Thus datasets of a certain country are simpler and the overhead in data size is negligible.

%The countries are however also at the data sets because queries for data sets of a certain country are anticipated to be more frequent and cross-country spendings are less comparable as intra-country ones.
%Because of this, our modelling approach is varied:
%All attributes from the source data which are represented at the observation level, which are different for each data set, are kept so in the target data.
%It is possible to check if all values of a certain attribute are the same and to move those attributes to the data set level but this is confusing for the users and inconsistent.
%OpenSpending data sets on the other hand, have two fixed attributes: the countries whose governments do the financial transactions and the currency of the transactions.

%The currency of a transaction is however directly tied to the amount spent and search queries which separate data sets by currency are less anticipated.
%Thus the overhead of specifying the currency at each observation is deemed to weigh less than the .

\subsection{Trust Dimensions}

- depends on the source data, for each dataset
most governements - high believability and reputation

\subsection{Dataset Dynamicity Dimensions}
Because spending data is historical, currency, volatility and timeliness are not applicable.
\fi
\subsection{Contextual Dimensions}
\paragraph{Completeness}
Because all of the attributes, measures, dimensions and observations of the original data sets are carried over to their Linked Data versions, the all transformed data sets are fully complete both on the schema and the data (instance) level. \todo{Nur so ``complete'' wie die Ausgangsdaten, oder? Ist auch nicht so interessant die Aussage, da offensichtlich.}
However of the 321 data sets listed by OpenSpending, only \num{252} were retrievable and \numberofdatasets{} of those convertible.

%\paragraph{Amount-of-data}
%The amount of observations of the data sets varies between two (spendings in Prague of about \SI{5000}{CZK} for an unknown purpose) and 242209 (\enquote{Spending from ministries under the Danish government}).
%See \autoref{fig:histogram} for the numbers of measures, attributes and dimensions.
%See \autoref{fig:amountofdata} for the distribution of the numbers of triples and observations and \autoref{fig:histogram} for the numbers of measures, attributes and dimensions.
\iffalse
\begin{figure}[h]
\includegraphics[width=\columnwidth]{img/pointcloud.pdf}
\caption{Numbers of triples and observations of the data sets (logarithmic scales)}
\label{fig:amountofdata}
\end{figure}
\fi
\iffalse
old data:
\fi

%\paragraph{Relevancy} This dimension is dependent on the task at hand.

\subsection{Representational Dimensions}

%\todo{- average length of names of measures, attributes,dimensions necessary?}

%The URIs are kept short and human readable by reusing the existing OpenSpending URLs which are constructed based on the names of the resources and their parent resources.
%For example, the dimension \emph{recipient} of the dataset \emph{nigerianbudg13} has the URL \url{os:nigerianbudg13/recipient}.
%They were also reused to facilitate future integration with the source dataset using content negotiation to determine whether the user or application needs an RDF representation.
%For resources which do not have a single source URL the spirit of the OpenSpending naming scheme was still followed. For example, the view \emph{} of the dataset \emph{berlin_de} \todo{example}.

%\paragraph{Representational Consistency}
\iffalse moved to extraction workflow
First and foremost, the RDF Data Cube vocabulary provides the backbone structure for all of the data sets.
This vocabulary uses the model of the Statistical Data and Metadata eXchange (SDMX) initiative\footnote{\url{http://sdmx.org}} which is widely used and, among others, sponsored by the United Nations Statistics Division, the International Monetary Fund and the World Bank.
RDF Data Cube is supported by tools such as the CubeViz~\cite{Auer+ISWC-2012} plugin, that enables the management of RDF Data Cubes in the OntoWiki (\todo{warten auf publikation von michas paper} \remark{JL: Ich kenne den Status nicht, aber auf andere Publikationen warten, ist selten eine gute Idee.}), which manages several parts of the the Linked Data Lifecycle, such as publication, authoring and visualization.
This facilitates the consumation of the data sets.
\fi
%\todo{Micha: sind die Zitierungen richtig?}
%The SDMX vocabularies\footnote{\url{http://purl.org/linked-data/sdmx/2009{/dimension\#,measure\#,attribute\#,concept\#}}} are used for observation values, currencies and time periods.
%\todo{move lifecycle part to a more fitting place?}
%\paragraph{Representational Conciseness}
\paragraph{Representational Conciseness and Understandability}
Because the source data contains labels for all data sets, attributes, dimensions and measures, those labels are also provided for the transformed data sets.
The resource URLs\footnote{except the observations whose last part is a hash value} are concise and  easily understandable, because they are generated from the names of the entities they describe.% (see \emph{Representational conciseness}).
This facilitates writing SPARQL queries (see \autoref{tab:sparqlqueries}) and interpreting the results.

%\begin{tabular}{ll}
%x&\begin{Verbatim}dfdf\end{Verbatim}
%\end{tabular}

\paragraph{Interpretability}
All objects are represented by globally unique URIs, many of them with labels (see the paragraph above). \todo{Leider ebenfalls keine so interessante Aussage.}

\paragraph{Versatility}
The data sets are available as resolved URIs, on a SPARQL endpoint\footnote{\sparql} and as \iffalse Turtle\footnote{\turtle} and \fi N-Triples\footnote{\ntriples} dumps. \todo{Wurde ebenfalls schon vorher gesagt.}

%\section{Examples and Critical Discussion of Typical Knowledge Modeling Patterns Used}
\iffalse
\section{Use Cases}

- CubeViz visualisierung
- 2 länder mit ähnlichen data sets -> zusammen analysieren
4 use cases sind am besten mit sparql queries
\fi
\end{comment}

\section{Related Work}\label{sec:relatedwork}
The TWC Data-Gov Corpus~\cite{datagov1,datagov2} consists of linked government data from the Data-gov project. However, it only contains transactions made in the US and does not overlap with OpenSpending.
The publicspending.gr project generates and publishes~\cite{spendinggreece} public spending data from Greece based on the UK payment ontology and without using statistical data cubes.
The UK government expenditure dataset COINS\footnote{\url{http://data.gov.uk/dataset/coins}} is available as Linked Data\footnote{\url{http://openuplabs.tso.co.uk/sparql/gov-coins}, in a beta version}.
\emph{LOD Around-The-Clock (LATC)}\footnoteurl{http://latc-project.eu} is a project, which was funded by the European Union (EU) and converted European open government data into RDF.
One of its outcomes is the FTS\footnoteurl{http://ec.europa.eu/budget/fts}~\cite{martin-fts} project, which transforms and publishes financial transparency data on EU spending.
In comparison with LinkedSpending, those projects also contribute linked government data but with a different or more limited scope.

Furthermore, there is the Digital Agenda Scoreboard~\cite{martin-scoreboard} is an EU project which keeps track of the transformation of statistical data to RDF.

\section{Conclusions, Shortcomings, Future Work}\label{sec:conclusions}
As shown in \Cref{sec:conversion}, we converted several hundreds of financial data sets to RDF and, as shown in \Cref{sec:publishing}, we published them as Linked Open Data in several ways.
%Thus, our contribution is that we have made available as Linked Open Data a large number of data sets.
However, we recognise a few shortcomings and our goal is to enrich the meta data with the help of domain experts and to refine the structure of the individual data sets.
Furthermore, we plan to improve the automatic configuration of CubeViz.
%improve the user interface \todo{``improve the user interface'' wirkt aus dem Zusammenhang gerissen}

\paragraph{Multilinguality}
RDF itself provides support for multilingualism, which is one of its key advantages to other representation formats.
The source data does not contain language tags, however, and the the languages used do not always match the country, the data refers to.
Automatic language detection on single labels did not yield a satisfying precision and it is not possible to increase the precision of the language detection by combining the estimates about several labels of an observation as their language is not always identical.
We plan statistical examinations of the relations between labels of different entities and more complex schemes based on those examinations, which can achieve language detection with a higher precision.
Additionally, we plan to automatically translate all literals to several languages.

% \section{Known Shortcomings and Future Work}
\paragraph{Individual Modelling}
Because the source data is already structured, the transformation of all the data sets without the need of text extraction and in an automatic way was feasible.
On a deep level however, there is much unmodelled structure that is unique to each data set or at most shared between several of them, for instance the categorization of spending into several specific \enquote{plans} in German budgets.
Because of the amount of data sets, modelling all details, and thus also improving the internal and external connectivity, requires either a large-scale cooperation or a crowd-driven approach, which we did not perform yet.
%\todo{Ich denke idealerweise sollte ein SPARQL-Query genau so einen Fall aufzeigen, bei dem wir das modelliert haben und damit über mehrere data sets Abfragen bauen konnten.}

\paragraph{Drilldowns}
Because of the hierarchical organization of the different coded properties \enquote{groups} and \enquote{functions}, the visualizations on \url{openspending.org} permit \enquote{zooming} (drilldown) in and out of the different levels of the data.
The RDF Data Cube vocabulary specifies the use of \url{skos:ConceptScheme} or \url{qb:HierarchicalCodeList} but neither variant is fully implemented and it is not clear, which of those modelling possibilities will become standard.
%\todo{Ist hier ``not implemented in CubeViz'' gemeint?}  -> das kann auch sein aber es ist auch noch nicht in den Daten selbst drin
%As soon as hierarchical code lists are better specified, the linear hierarchy of the budget groups and function should be modelled.

\subsection{Future Work}

% interlinking
\paragraph{Interlinking}
Extensive interlinking of referenced entities to the all-purpose knowledge base of DBpedia provides additional context.
Coded property values, such as the budget areas healthcare and public transportation, can be interlinked with their respective DBpedia concepts.
This enables the usage of type hierarchies and thus new ways of structuring the data and provides more meaningful aggregations and new insights.

\paragraph{Question Answering}
We plan to develop a question answering system that allows accessing statistical Linked Data in the form of RDF Data Cubes using natural language questions~\cite{statisticalqashort}.
LinkedSpending is used both as the first knowledge base and for performance evaluation.

%\paragraph{more elaborate updating with different modules with a diff component a scheduler ...}

\iffalse
\paragraph{Reuse}
Because of the large amount of data sets, the conversion process is necessarily automatic.
While for some of the often used component properties, like date and time, resources from existing vocabularies such as SDMX are used, in the general case, new resources are created for all values.
A thorough manual inspection might yield in an increased usage of existing vocabulary.
Also some of the dimensions may be stable over different data sets and thus might be shared, which decreases the amount of redundancy and increases the benefit of those resources, should they be referenced from other resources.

\paragraph{Hosting integration with OpenSpending}
According to the Linked Data principles, the RDF resources we created should be resolvable.
As openspending.org already uses URLs with different extensions to deliver different content types like JSON and HTML, adding content negotiation and hosting the resources at there would be easy to implement.
To speed up the integration of new data, a web service that watches for changes in the input data and runs the transformation to keep the RDF data up to date would be useful. 
%\paragraph{Webservice}
\fi

\iffalse
\begin{table}
\begin{tabulary}{\columnwidth}{LSS}
\textbf{Country}	&\textbf{Number of data sets}	&\textbf{Number of observations}\\
\end{tabulary}
\caption{spending data per country}
\label{tab:countries}
\end{table}
\fi
%\todo{write somewhere: county codes modelled with SDMX Codelist CL_AREA following ISO 3166 alpha-2(for example \enquote{DE} stands for Germany) -> better: link with linkedgeodata}
\iffalse
\begin{table}
\begin{tabulary}{\columnwidth}{LS}:
\textbf{}	&	&\\
\end{tabulary}
\caption{spending data per country}
\label{tab:countries}
\end{table}
\paragraph{URL renaming}
The URLs of most of the transformed resources are equal (except for the prefixes) to those of OpenSpending in order to facilitate a common hosting of both RDF and the original resources with content negotiation.
However the original naming schemes lead less understandable to URLs like the dimension \url{ls:nigeriabudg13/recipient} which cannot be prefixed at the core schema level because RDF local names cannot contain the slash character \enquote{/}.
If this kind of hosting is not wanted by OpenSpending, other naming schemes are better, for example \ls\url{/dimensions/recipient}.
\fi
% \section*{Acknowledgements}
\medskip\textbf{Acknowledgements:}
Special thanks goes to the people behind the OpenSpending project, including Friedrich Lindenberg for suggesting the conversion.

% \cite{hoeffner-2013-kesw}
\bibliographystyle{abbrv}
\bibliography{paper,../../bib/aksw.swj}
%\bibliography{paper,../../bib/aksw.twocolumn.abbrv,../../bib/aksw.pending}

\end{document}

### 2019-03-01 paper.tex
Clean main template and correct todonotes display.

@@
-% please run ``latex paper'' and then ``makeglossaries paper'' from the command line to create the glossaries files
 % add. options: [seceqn,secthm,crcready,onecolumn]
@@
 %\usepackage{endnotes}
-
 \usepackage{booktabs}
 \usepackage{csquotes}
-\usepackage{graphicx}
-\usepackage{listings}
-\lstset{language=SPARQL,breaklines=true}
-
+%\usepackage{graphicx}
+\usepackage{todonotes}
+\setlength{\marginparwidth}{2.1cm}% enough space for todonotes
+%\usepackage{listings}
+%\lstset{language=SPARQL,breaklines=true}
 \usepackage{cleveref}
@@
 \thanks{Corresponding author. \printead{e1}.}},
-\author[A]{\fnms{Franziska} \snm{Jahn}\ead[label=e2]{konrad.hoeffner@imise.uni-leipzig.de}}
+\author[A]{\fnms{Franziska} \snm{Jahn}\ead[label=e2]{franziska.jahn@imise.uni-leipzig.de}}
 \author[A]{\fnms{Birgit} \snm{Schneider}\ead[label=e3]{birgit.schneider@imise.uni-leipzig.de}}
@@
 \address[A]{Institute for Medical Informatics, Statistics and Epidemiology (IMISE),
-\orgname{University of Leipzig}, \cny{Germany}\printead[presep={}}}]{e1,e2,e3,e4,e5,e6}}
+\orgname{University of Leipzig}, \cny{Germany}\printead[presep={\\}]{e1,e2,e3,e4,e5,e6}}
 %Medical Informatics, Management of Health Information Systemsi
 %Härtelstraße 16--18, D-04107 Leipzig
-E-mail: \{khoeffner,fjahn,alfred\}@imise.informatik.uni-leipzig.de}
+
 \begin{abstract}
@@
 \end{abstract}
+
 \begin{keyword}
-\kwd{Hospital Information Management}
+\kwd{information management, information systems, hospital information management}
 \end{keyword}
+
 \end{frontmatter}
-%%%%%%%%%%% The article body starts:
-\section{Introduction}\label{sec:introduction}
-A W3C design issue~\cite{puttinggovernmentdataonline} motivates making government data available online as Linked Data for three reasons:
-\enquote{1) Increasing citizen awareness of government functions to enable greater accountability; 2) Contributing valuable information about the world; and 3) Enabling the government, the country, and the world to function more efficiently.}
-Increasing the transparency of government spending specifically is in high demand from the public.
-For instance, in the survey publication~\cite{localtransparency}, \enquote{Public access to records is crucial to the functioning government} was rated with a mean of 4.14 (1 = disagree completely, 5 =  agree completely).
-Open spending data can reduce corruption by increasing accountability and strengthening democracy because voters can make better informed decisions.
-Furthermore, an informed and trusting public also strengthens the government itself because it is more likely to commit to large projects (see~\cite{fiscaltransparency} for details).
 
-%Both effects are proven in a study~\cite{fiscaltransparency}.
-Several States and Unions are bound to financial transparency by law, such as the European Union\footnote{\enquote{2. The Commission shall make available, in an appropriate
-and timely manner, information on recipients, as well as the nature and purpose of the measure financed from the budget[...]}~\cite{euregulations}} with its \emph{Financial Transparency System (FTS)}\footnote{\url{http://ec.europa.eu/budget/fts}}~\cite{martin-fts}.
-Public spending services satisfy basic information needs, but in their current form they do not allow queries which go further than simple keyword search or which cannot be answered with data from one system alone.
-Linked Data solves those problems by providing a unified format, a powerful query language and the possibility of integration with linked data sets from other services.
-
-Our contribution is an RDF transformation of the OpenSpending\footnote{\url{http://openspending.org}} project which provides government spending financial transactions from all over the world and is thus suitable as a core knowledge base that can be enriched and integrated with other, more focused data sets.
-Transforming OpenSpending to Linked Data and publishing it adds to and profits from the Semantic Web which offers benefits including a standardized interface, easier data integration and complex queries over multiple knowledge bases.
-
-%Some advantages of an RDF version over a relational database are (1) querying over multiple data sets at once using interlinks and (2) an additional custom vocabulary for each of the datasets, which are very heterogeneous (see \autoref{tab:sparqlqueries} for SPARQL queries for common use cases).
-
- %linkage provides benefits blablabla ...
-%\todo{more motivation why RDF is good users are able to get this and that, what is linked, e.g.
-financial transparency system, cordis -> big benefit at least it is possible to link to this and that list other data sets, such as ...
-e.g.
-the paper about greece}
-
-%We present LinkedSpending, which provides approximately two million financial transactions from all over the world extracted from OpenSpending as \emph{Linked Open Data}.
-%information needs such as browsing of data sets, visualization and simple queries but they suffer from
-
-The structure of the paper is as follows.
-\Cref{sec:motivation} motivates the work and presents use cases.
-\Cref{sec:openspending} describes OpenSpending, which is the source of the data, and its statistical data model.
-\Cref{sec:conversion} explains the target RDF Data Cube vocabulary and the transformation process to it.
-\Cref{sec:publishing} describes, how and where the data set is published and in which way users can access the data.
-%\Cref{sec:overview} gives an overview over the transformed linked dataset, describes the its schema and shows typical SPARQL queries.
-\Cref{sec:overview} gives an overall view of the data sets, gives details about the licence used and describes the data sets it is interlinked to.
-%This is followed by a description of usage scenarios in \Cref{sec:usagescenarios}.
-\Cref{sec:relatedwork} presents related spending data sets as Linked Open Data (LOD).
-%The data quality is analyzed in \cref{sec:dataquality} and
-The last section discusses known shortcomings of the data sets and future work. The prefixes used throughout this publication are defined in \Cref{tab:namespaces}.
-In order to save space, prefixes are used even when technically incorrect, such as in \resource{ls:berlin\_de/model}.
-
-\begin{table}[hbtp]
-\caption{Namespaces and prefixes used in the paper}
+\begin{table}
+\caption{}
 \label{tab:namespaces}
-%\tiny
-\scriptsize
-%\footnotesize
-\renewcommand{\tabcolsep}{5pt}
-\begin{tabularx}{\columnwidth}{
-   >{\hsize=.25\hsize}L 
-   >{\hsize=1.75\hsize}L 
-}
+\begin{tabular}{ll}
 \toprule
-\textbf{prefix}	&\textbf{URL}\\
+\textbf{x}	&\textbf{y}\\
 \midrule
-os			&\os\\
-owl			&\url{http://www.w3.org/2002/07/owl\#}\\
-ls			&\ls\\
-lso			&\lso\\
-qb			&\url{http://purl.org/linked-data/cube\#}\\
-sdmxd			&\url{http://purl.org/linked-data/sdmx/2009/dimension\#}\\
-dbpedia			&\url{http://dbpedia.org/resource/}\\
-dbp			&\url{http://dbpedia.org/property/}\\
 \bottomrule
-\end{tabularx}
-\vspace{-20pt}
+\end{tabular}
 \end{table}
 
-%while \cref{sec:transformation} explains the Data Cube model and the transformation.
-%-The purpose of LinkedSpending is explained in \cref{sec:purposeoflinkedspending}.
+\section{Background}
+\todo{not Konrad: write background of meta model}
 
-%While the OpenSpending API is enough to satisfy basic information needs such as browsing of data sets, visualization and simple queries, Linked Data provides many more benefits such as complex queries and the possible integration of other datasets such as the FTS and
-%OpenSpending.org is a an open platform that provides public finance data from governments around the world.
-%At the time of this writing,
+\section{Architecture}
+The \textbf{Semantic Network of Information Management in Hospitals} (\enquote{Krankenhaus} in German) is an OWL 2 DL ontology that consists of a central meta model and modular subontologies based on the meta model.
 
-%We converted this data to RDF Data Cubes using a Java program which can be called again should data be added or existing data updated.
+\section{Data Set Description}
 
-\section{Motivation}\label{sec:motivation}
-%\section{Usage Scenarios}\label{sec:usagescenarios}
-In a time of globalization,
-%in the finance sector
-financial data becomes an international network.
-RDF data with its linked nature supports a representation that takes this network nature into account.
-As a machine interpretable format, it lowers the access barrier for application developers.
-For instance, generic Linked Data tools such as OntoWiki, CubeViz and Facete provide end users with the means to explore the data and discover new insights.
+\section{Application}
 
-%\textbf{Community Empowerment:}
-%network effect 
-%reduce corruptions
-% corruption by increasing accountability and strengthening democracy because voters can make better informed decisions.
-%Furthermore, an informed and trusting public also strengthens the government itself because it is more likely to commit to large projects.
-%hub
-
-%domain specific tool on dataset
-
-%generic tool
-
-%\paragraph{Internationalization Support}
-
-%Apart from the information needs that can already be satisfied using the source data, there are usage scenarios that are easier or only possible with LinkedSpending.
-
-%\subsection{Economic Analysis}
-\paragraph{Economic Analysis}
-LinkedSpending is represented in Linked Open Data, which facilitates data integration.
-Currencies from DBpedia and countries from LinkedGeoData are already integrated.
-Financial data offers further integration candidates, such as political or other statistical, policy-influencing data such as health care.
-This allows queries such as query 7 in \autoref{tab:sparqlqueries}, which asks for data sets with currencies whose inflation rates are greater than \SI{10}{\percent}.
-%\todo{Ev. noch ein weiteres Beispiel um das noch zu untermauern und etwas mehr Substanz in die Subsection zu bringen - auch im Vergleich zu den folgenden beiden Subsections.}.
-
-LinkedSpending can also be used to compute economic indicators across several data sets.
-A possible indicator is a country's spending on education per person where the population size can be taken from the LinkedGeoData countries linked from one or more budget data sets.
-One such data set is \dataset{ugandabudget}, which contains the Uganda Budget and Aid to Uganda, 2003--2006.
-LinkedSpending serves as a hub for the integration of those data sets and their provenance information.
-More data sets can be integrated with similarity-based interlinking tools such as LIMES~\cite{NGON11}.
-%\footnotetext{and its web interface SAIM\cite{Lyko2013}}%, available at \url{http://saim.aksw.org}}
-
-%\subsection{Putting Spending Amounts into Perspective}
-%\subsection{Finding and Comparing Relevant Datasets}
-\paragraph{Finding and Comparing Relevant Data Sets}
-
-Government spending amounts are often much higher than the sums ordinary people are used to dealing with but even for policy makers it is hard to understand whether a certain amount of money spent is too high or normal.
-Comparing data sets and finding those which are similar to another one helps separating common values from outliers which should be further investigated.
-For example, if another country has a similar budget structure but spends way less on health care with a similar health level,
-it should be investigated whether that discrepancy is caused by inherent differences such as different minimum wages or a different climate or if it is due to preventable factors such as inefficiencies or corruption.
-While OpenSpending provides several hundreds of data sets which can be searched and it allows browsing and visualization of any single one, it does not provide a comparison function between data sets.
-Because of the mechanism to identify equivalent properties (see \autoref{sec:conversion}), SPARQL queries can compare different data sets, e.g.~between similar structures in different countries.
-Query 9 in \autoref{tab:sparqlqueries} shows a simple query to detect data sets which are most similar to any particular data set.
-This is done by calculating the number of common measures, attributes and dimensions.
-\iffalse
-\subsection{Increased E-Government Transparency}
-While OpenSpending makes government spending data available, the data can be made more understandable by providing additional knowledge about the described entities and by facilitating access to the data by non-expert users.
-\paragraph{Additional Knowledge}
-Through integration with other knowledge bases, users can for example learn about the country and currency a data set refers to at the same time as they query the data.
-\todo{Scheint mir nicht ideal formuliert.
-Während des Queries etwas lernen, kann man höchstens wenn der Endpoint sehr langsam ist.
-Ev.
-"while they browse and explore the data", falls das gemeint ist?
-Es wäre auch gut, wenn 7.1 und 7.3 nicht beide identisch Integration von country und currency erwähnen, da es sonst so wirkt, als ob es nicht viele andere Möglichkeiten gibt und man versucht den gleichen Punkt mehrfach an die Reviewer zu verkaufen.}
-This helps interpreting the values and constructing helpful indicators. For example, the budget countries spend on health care varies wildly but less so when it is divided by the population number.
-
-\todo{ich habe das qa jetzt rausgenommen da mir kein besonders guter use case eingefallen ist und die aus den example sparql queries zu kompliziert wären für eine erfolgreiche beantwortung}
-\fi
-\iffalse
-\paragraph{Question Answering}
-The OpenSpending data is available through numerous interfaces such as a JSON API, a keyword search and graphical visualizations.
-Linked Open Data however allows the additional usage of Question Answering applications in order to to query the data in a way which is both more intuitive for non-experts than SQL and more expressive than keyword queries, which suffer from the inability to express relations and context, or faceted search, which only allows certain types of search restrictions.
-Question Answering approaches, such as TBSL~\cite{unger2012template}, can thus be adapted to the spending domain and then allow not only computer science experts, but journalists or just ordinary people to know more about it more intuitively using natural, full-text sentences.
-\todo{Der Paragraph sollte noch etwas angepasst werden um zu zeigen, warum Question Answering gerade hier günstig ist, z.B.
-eine konkrete Abfrage oder man probiert tatsächlich TBSL mit den Daten aus.
-Aktuell liest sich der Paragraph eher wie eine Begründung warum Question Answering über RDF allgemein sinnvoll ist (was zweifelslos der Fall ist, aber die Relevanz für das Paper noch nicht begründet).}
-\fi
-
-%\section{Source for the Data and Purpose of the Transformation}\label{sec:sourceforthedataandtopiccoverage}
-%\section{Purpose of LinkedSpending}\label{sec:purposeoflinkedspending}
-% viellecht besser passende beispiele nehmen, also anstelle von Kontinenten Gruppen und Hauptgruppen nehmen
-%For example the coded property \url{:continents} could have the range \url{:Afrika}, \url{:America}, \url{:Asia}, \url{:Australia} and the child property \url{:countries} ...
-(spezielle properties für jeden kontinent?)
-%This parent-child relationship is different from rdfs:subClassOf because the values are disjunctive but instead the childrens values are parts of the parents values.
-
-%grouped in slices data sets 
-%structure defined by datastructure definitions
-
-%At the time of this writing, OpenSpending contains 217 different data sets.
-%While they share the general \emph{budget} domain, they still have slightly different contexts and are described in different vocabularies.
-
-\iffalse
-2-3 beispiele
-
-statistik darüber bringen
-
-manuelles proof über die qualität
-
-am ende die datacubes nehmen
-
-data quality dashboard vom data cube nehmen
-\fi
-
-\section{OpenSpending Source Data}\label{sec:openspending}
-OpenSpending\footnote{\url{http://openspending.org/}} is a project which aims to track and analyze public spending worldwide and, at May 2014, contains more than 25 million financial entries in \numberofopenspendingdatasets{} data sets\footnote{As some of them contain errors, the number of LinkedSpending data sets is slightly smaller.}.
-Data Sets can be submitted and modified by anyone but they have to pass a sanity check from the OpenSpending Data Team which also cleans the data before publishing.\footnote{\url{http://community.openspending.org/contribute/data/}}
-OpenSpending hosts transactional as well as budgetary data with a focus on government finance.\footnote{\url{http://community.openspending.org/help/guide/en/financial-data-types/}}
-It contains this data in structured form stored in database tables and provides searching and filtering as well as visualizations and a JSON REST interface.
-The data sets differ in granularity and type of accompanying information, but they share the same meta model.
-%, e.g. from the FTS or CORDIS\footnotemark.
-\subsection{The Data Cube Model}%\label{sec:rdfdatacubemodel}
-The domain model of OpenSpending is a \emph{data cube} (also \emph{OLAP cube}, \emph{hypercube}), which represents multi-dimensional statistical observations.
-Each cell corresponds to an observation (an instance of spending or revenue) that contains measurements (e.g. the amount of money spent or received).
-The context of the measurement is provided by the \emph{dimensions} like the purpose, department and time of a spending item and optionally by \emph{attributes}, which further describe the measured value, e.g., the unit of the measurement.
-%Apart from the fixed data cube meta model, the structure of each dataset is completely up to the creator.
-%The budget data sets for example have the dimensions \emph{time} and \emph{department}, the measure \emph{amount spent} and the attribute \emph{currency} with the value \emph{euro}.
-\begin{figure}[h]
-\footnotesize
-\begin{lstlisting}[language=json]
-"sub-programme": {
- "label": "Sub-programme",
- "type": "compound",
-},
-"amount": {
-    "datatype": "float",
-    "label": "Total",
-    "type": "measure",
-}
-\end{lstlisting}
-\caption{simplified excerpt of an OpenSpending \emph{model}} 
-\label{fig:jsonmodel}
-\end{figure}
-
-\begin{figure}[h]
-{\footnotesize
-\begin{lstlisting}[language=json]
-"sub-programme": {
- "label":"Security and safeguarding liberties",
- "html_url":"http://openspending.org/eu-budget/sub-programme/security-and-safeguarding-liberties",
- "name":"security-and-safeguarding-liberties"
-},
-"html_url": "http://openspending.org/eu-budget/entries/017dfcb58d05671ef9eb5a9f77fef39c8b14150c",
-"amount": 41.2
-\end{lstlisting}
-}
-\iffalse
-"main-programme": {
-
-    "description": null,
-    "label": "Main programme",
-    "facet": true,
-    "key": true,
-    "attributes": {
-        "name": {
-            "column": "StringCAT1",
-            "datatype": "id",
-            "default_value": ""
-        },
-        "label": {
-            "column": "StringCAT1",
-            "datatype": "string",
-            "default_value": ""
-        }
-    },
-    "type": "compound",
-    "dimension": "main-programme"
-},
-"main-programme": {
- "label":"Citizenship, Freedom ,Security and Justice",
- "html_url":"http://openspending.org/eu-budget/main-programme/citizenship-freedom-security-and-justice",
- "name":"citizenship-freedom-security-and-justice"
-},
-
-\fi
-\caption{simplified excerpt from an OpenSpending \emph{entry}}
-\label{fig:jsonentry}
-\end{figure}
-
-\Cref{fig:jsonmodel} shows an excerpt from the model of the OpenSpending data set \emph{eu-budget} with the dimension \emph{sub-programme} and the measure \emph{amount}.
-\Cref{fig:jsonentry} shows the an entry that contains the actual values for the dimension and the measure of the observation.
-
-\subsection{Problems}
-While the data is well-structured and thus suitable for conversion without data cleaning or extensive preprocessing, it still poses problems that need to be taken into account:
-\begin{inparaenum}
-\item New data sets are frequently added (approximately 50 per month) and, less often, existing data sets are modified.
-\item Some data sets do not specify a value for all properties in all observations.
-\item There are properties with the same name in different data sets where it is unknown if they specify the same property.
-\item Data Cube is a meta model. The deep structure of the data sets is heterogeneous and
- described only shallowly.
-\item The language of literals is varying between and even within data sets but the language used is not specified.
-\end{inparaenum}
-Points 1 to 3 are addressed in the next section while points 4 and 5 are discussed in \cref{sec:conclusions}.
-
-\section{Conversion of OpenSpending to RDF}\label{sec:conversion}
-
-%The conversion from OpenSpending to Linked Data consists of the extraction of the source data and its transformation to RDF.
-%Because the source data adheres to the data cube model, a conversion of the data to RDF needs an appropriate RDF vocabulary.
-%The RDF DataCube vocabulary~\cite{rdfdatacube} is an ideal fit for the transformed data.
-%Because the source data uses the data cube model, the RDF DataCube vocabulary~\cite{rdfdatacube} is an ideal fit for the transformed data.
-
-\begin{table*}[hbtp]
-\caption{Conversion of OpenSpending to LinkedSpending classes and instances}
-\label{tab:transformation}
-\scriptsize
-\renewcommand{\tabcolsep}{5pt}
-\begin{tabularx}{\textwidth}{
-% sum of hsizes needs to be 2
-   >{\hsize=.1\hsize}L 
-   >{\hsize=.4\hsize}L 
-   >{\hsize=.8\hsize}L 
-   >{\hsize=.6\hsize}L 
-   >{\hsize=.3\hsize}L          
-}
-\toprule % a bit unelegant: with the replacement of \url by \urlX to circumvent hyperref the line break inside urlX does not work so there are 2 spaces
-&\textbf{Source URL}  		&\textbf{JSON Path}				&\textbf{LinkedSpending class}		&\textbf{LS instance scheme}\\ 
-\midrule
-\rom{1}  		&\osdatasetjson 		&						&\class{qb:DataSet}			&\lsdataset\\
-\rom{2}			&\osmodel 			&						&\class{qb:DataStructureDefinition}	&\dsd\\
-\midrule
-\multirow{3}*{\rom{3}} 	&\osmodel			&\texttt{\$.mapping.*}				&\class{os:\{Country,Time\}Component} \class{Specification}
-													or \class{qb:ComponentSpecification}	&\urlX{lso:}\emph{propertyname} \urlX{-spec}\\
-\midrule
-\rom{4}			&\multirow{4}*{\osmodel}	&\texttt{\$.mapping.*[?(@.type="compound")]}	&\class{qb:DimensionProperty}		&\multirow{4}*{\urlX{lso:}\emph{propertyname}}\\
-\rom{5}			&				&\texttt{\$.mapping.*[?(@.type="date")]}		&\class{qb:DimensionProperty}	&\\
-\rom{6}			&				&\texttt{\$.mapping.*[?(@.type="measure")]}	&\class{qb:MeasureProperty}		&\\
-\rom{7}			&				&\texttt{\$.mapping.*[?(@.type="attribute")]}	&\class{qb:AttributeProperty}		&\\
-\midrule
-\rom{8} 		&\osdataset\urlX{/entries.json} 		&\texttt{\$.results[*].dataset}	&\class{qb:Observation}				&\urlX{ls:}\urlX{observation-} \emph{dataset name}-\emph{hashvalue}\\
-\bottomrule
-\end{tabularx}
-\end{table*}
-%\iffalse
-%\begin{sidewaystable}
-%{\setlength{\tabcolsep}{0.5em}
-%\scriptsize
-%\begin{tabulary}{0.84\textwidth}{lllLp{3cm}p{3cm}}
-%\toprule
-%			&\textbf{Source URL}  		&\textbf{JSON Path}				&\textbf{LinkedSpending class}		&\textbf{OpenSpending URL}				&\textbf{LinkedSpending URI}\\ 
-%\midrule
-%\rom{1}  		&\osdatasetjson 		&						&\class{qb:DataSet}			&\osdataset						&\lsdataset\\
-%\rom{2}			&\osmodel 			&						&\class{qb:DataStructureDefinition}	&\osmodel						&\dsd\\
-%\midrule
-%\multirow{3}*{\rom{3}} 	&\osmodel			&\texttt{\$.mapping.*}				&\class{os:{Country,Time}ComponentSpecification}
-%													or \class{qb:ComponentSpecification}	&none							&\lsdataset\url{/}\emph{propertyname} \url{/component} \url{Specification}		\\
-%\midrule
-%\rom{4}			&\multirow{4}*{\osmodel}	&\texttt{\$.mapping.*[?(@.type="compound")]}	&\class{qb:DimensionProperty}		&\multirow{4}*{\osdataset\url{/}\emph{propertyname}}	&\multirow{4}*{\lsdataset\url{/}\emph{propertyname}}\\
-%\rom{5}			&				&\texttt{\$.mapping.*[?(@.type="date")]}		&\class{qb:DimensionProperty}	&							&\\
-%\rom{6}			&				&\texttt{\$.mapping.*[?(@.type="measure")]}	&\class{qb:MeasureProperty}		&							&\\
-%\rom{7}			&				&\texttt{\$.mapping.*[?(@.type="attribute")]}	&\class{qb:AttributeProperty}		&							&\\
-%\midrule
-%\rom{8} 		&\osdataset\url{/entries.json} 		&\texttt{\$.results[*].dataset}	&\class{qb:Observation}			&\osmodel\url{/entries/} \emph{hashvalue}			&\lsdataset\url{/model/entries/}\emph{hashvalue}\\
-%\bottomrule
-%\end{tabulary}
-%\caption{Transformation of OpenSpending entities to LinkedSpending classes}}
-%\label{tab:transformation}
-%\end{sidewaystable}
-%\fi
-%\todo{\autoref{tab:transformation} source uris take the ones where i get them from or where the source uri is that only describes it? should be both acutally}
-
-%\paragraph{The RDF DataCube vocabulary}
-\paragraph{}
-The RDF DataCube vocabulary~\cite{rdfdatacube}, i.e.~an RDF variant of the previously explained data cube model, is an ideal fit for the transformed data. 
-\begin{figure}[h]
-\includegraphics[width=\columnwidth]{img/rdfdatacube_shrinked.pdf}
-\caption{Used RDF DataCube concepts and their relationships\protect\footnotemark}
-\label{fig:rdfdatacube}
-\end{figure}
-\footnotetext{Simplified version of the structure described in ~\cite{rdfdatacube}.}
-
-First and foremost, this vocabulary provides the backbone structure for every LinkedSpending data set, see \cref{fig:rdfdatacube}.
-Each data set is represented by an instance of \class{qb:DataSet} and an associated instance of \class{qb:DataStructureDefinition} which includes \emph{component specifications} (see \cref{fig:modellingexample} for an example).
-Each component specification is associated to a \emph{component property} which can be either a \emph{dimension}, an \emph{attribute} or a \emph{measure}.
-Commonly used concepts are specified in the model of the \emph{Statistical Data and Metadata eXchange (SDMX)} initiative\footnote{\url{http://sdmx.org}}.
-%\Cref{fig:modellingexample} shows, how
-%example, a reference area can make use of \resource{sdmxa:refArea}.
-The RDF Data Cube vocabulary  is supported by the LOD2 Statistical Office Workbench\footnote{\url{http://demo.lod2.eu/lod2statworkbench}} which is part of the Linked Data Stack (an advanced version of the the LOD2 Stack~\cite{Auer+ISWC-2012}).
-The workbench includes a DataCube validator, a split and merge component and a CKAN Publisher.
-The OntoWiki~\cite{ontowiki_www}, which manages several parts of the the Linked Data Lifecycle~\cite{Auer+ISWC-2012}, such as Storage/Querying and Search/Browsing/Exploration offers a CSV import plugin for the format as well as a faceted RDF Data Cube browser, CubeViz.
-Data cubes may contain slices, which are presets for certain dimension values, effectively selecting a subset of a cube.
-Users may create and visualize their own slices using the OntoWiki CubeViz plugin.
-Furthermore, the RDF DataCube vocabulary allows the persistence of slices which is used to represent preconfigured slices from OpenSpending.
-
-% the Linked Data Stack~\cite{Auer+ISWC-2012}
-% which others? ORE - Ontology Repair and Enrichment ?
-
-\begin{figure}
-\footnotesize
-\begin{lstlisting}[language=ttl]
-ls:berlin_de
- rdf:type	qb:DataSet;
- rdfs:label	"Berlin Budget";
- dc:source	os:berlin_de;
- qb:structure	ls:berlin_de/model;
- qb:slice	ls:berlin_de/views/nach-einzelplan.
-
-ls:berlin_de/model
- rdf:type qb:DataStructureDefinition;
- qb:component
-   lso:CountryComponentSpecification,
-   lso:DateComponentSpecification,
-   lso:Einzelplan-spec,
-   lso:amount-spec.
-
-lso:CountryComponentSpecification 
- rdf:type qb:ComponentSpecification;
- rdfs:label "country";
- qb:attribute sdmxa:refArea; 
- qb:componentRequired "false"^^xsd:boolean;
- qb:componentAttachment
-   qb:DataSet,qb:Observation.
-\end{lstlisting}
-\caption{RDF DataCube vocabulary modelling excerpt of data set \resource{berlin\_de} (some properties and values omitted).}
-\label{fig:modellingexample}
-\end{figure}
-
-\paragraph{Transformation}
-All of the OpenSpending data sets describe observations referring to a specific point or period in time and thus undergo only minor changes.
-New data sets however, are frequently added. Because of this, the huge number of data sets and their size, an automatic, repeatable transformation is required.
-This is realized by a program\footnote{written in Java, available as open source at \url{https://github.com/AKSW/openspending2rdf}} which fetches a list of data sets on execution and only transforms the ones who are not transformed yet.
-Each data set is transformed separately.
-\autoref{tab:transformation} shows for each class used by LinkedSpending, at which URL (abbreviated using the prefixes from \cref{tab:namespaces}) the information used to create the instances of those classes is found.
-% cref somehow doesnt work with this table
-In case there are multiple instances described at one URL, a \emph{JSON path}\footnotemark{} expression is given, that locates the corresponding subnodes.
-\footnotetext{\emph{JSON path} (\url{http://code.google.com/p/json-path/}) is a a query language for selecting nodes from a JSON documents, similar to XPath for XML}
-Finally, the table contains the patterns that describe resulting LinkedSpending URLs.
-For example, the OpenSpending URL \resource{os:berlin\_de/model} contains the node \texttt{\$.mapping.amount} which has a type value of \enquote{attribute} and is, thus, transformed to the OpenSpending instance \resource{lso:amount} of the class \resource{qb:AttributeProperty}.
-
-Equivalent component properties (dimensions, attributes and measures) are identified as follows:
-A configuration file optionally specifies the mapping of data set and property name to an entity in the LinkedSpending ontology.
-By default, the property URI is derived from the property name.
-Properties with the same name in different data sets not having a mapping entry that states otherwise are assumed to represent the same concept and thus given the same URL.\footnotemark
-\footnotetext{Although that has the possibility of mismatches, such a mismatch has not been spotted yet. Still, evaluating and, if necessary, improving the automatic matching is part of future work.}
-
-%\subsection{Use of Established Vocabularies}
-\paragraph{Use of Established Vocabularies}
-%\textbf{Use of Established Vocabularies:} 
-In addition to the standard vocabularies, RDF, RDFS, OWL and XSD, the DCMI vocabulary is used for source and generation time metadata.
-The data sets are modelled, first and foremost, according to the RDF Data Cube vocabulary, % (see \cref{sec:openspending}),
-which specifies the structure of a data cube.
-LinkedSpending follows the RDF Data Cube recommendation to make heavy use of the SDMX model for measures, attributes and dimensions.%, among others, sponsored by the United Nations Statistics Division, the International Monetary Fund and the World Bank.
-The data sets are very heterogeneous but there are some properties which are commonly specified and thus modelled with established vocabularies.
-The year and date, a data set and an observation refers to, respectively, is expressed by \resource{sdmx-dimension:refPeriod} and XSD.
-%other than the usage of time and dates, which are expressed  currencies and countries.
-
-Currencies are taken from DBpedia~\cite{dbpedia-swj} and countries are represented using the vocabulary of LinkedGeoData~\cite{SLHA11}, a hub for spatial linked data.
-Some amount of data is imported from LinkedGeoData countries and DBpedia currencies.
-Because of the limited number of countries and currencies, and properties values imported per country and currency, the amount of data is too small to consider federated querying.
-As most countries and currencies are stable in the medium term, this data needs to be updated only infrequently.
-\paragraph{Interlinking}
-%\textbf{Interlinking:}
-There are two possibilities to align entities to another vocabulary: 1) to use the entities directly and 2) to create an own RDF resource with interlinks, like \resource{owl:sameAs}, to that vocabulary.
-We generally preferred the first approach because a higher amount of reuse provides easier integration, better understandability and tool support.
-%We generally preferred to use entities of a vocabularies directly, such as the currencies and countries, because this is a more elegant and less convoluted approach as long as the used vocabularies are stable and exactly correspond to the required meaning.
-%, as long as the entities are always in the object, or property position.
-% wir brauchen nicht extra begründen warum man subjects verlinken muss um dann zu sagen, dass es sowieso keine observations gibt, die man verknüpfen kann (da fragt sich der leser warum wir das schreiben)
-%Using an entity of another vocabulary in the subject position is problematic, however, because, amongst other things, those additional triples are not reachable by standard Linked Data URI dereferencing.
-%Because of this, the interlinks between observations of the data sets to observations in other data cubes need to be aligned by interlinking and not by replacement.
-% Observations are only equal to others however, when they describe exactly the same transaction and overlapping data sets from other sources have not been found.
-% While there are usually no interlinking targets for \emph{sameAs} links to other observations, there are many possibilities for interlinks between data sets or dimension values (which are always URIs, mostly with labels) and concepts they refer to.
-While we did not find \emph{sameAs} link targets on observation level, i.e.~exactly the same statistical observations described in other data sets, there are many possibilities for interlinks between data sets or dimension values and concepts they refer to.
-Using the labels of those data sets and dimension values, it is possible, for example, to link values of the dimension \enquote{region} of a federal budget, and thus indirectly also the observations which use those values, to the cities in DBpedia or LinkedGeoData whose labels are contained in the label of the region value URI.
-
-%\todo{Wenn zwei Datensätze sich z.B.
-auf Städte beziehen, könnte man vermutlich schon mehr Sachen verknüpfen.
-Das Problem bei dem Absatz ist, dass Du hier sagst, dass es quasi keine Linkmöglichkeiten gibt, aber unter 7.1 betonst, dass man mit mehr Links noch mehr Analysen durchführen kann.}
-
-%, such as \ls\url{/berlin\_de}, 
-%, can browse single URLs as HTML through the Pubby\footnote{\url{http://wifo5-03.informatik.uni-mannheim.de/pubby/}} interface by dereferencing the URIs, such as \ls\url{/berlin\_de}, in a browser.
-
-%The user can browse single URLs as HTML through the Pubby\footnote{\url{http://wifo5-03.informatik.uni-mannheim.de/pubby/}} interface by dereferencing the URIs, such as \ls\url{/berlin\_de}, in a browser.
-
-%Because mass processing such as interlinking requires computationally expensive queries against the SPARQL endpoint, an RDF dump\footnote{\ntriples} is also available.
-\todo{Satz würde ich eher weglassen bzw.
-verschieben.
-Es ist recht klar, dass es einen Dump gibt und es passt nicht in die Section.}
-%\Cref{tab:technicaldetails}
-
-\paragraph{Error Handling}
-The OpenSpending API lists \numberofopenspendingdatasets{} data sets with \numberofdatasets{} of them having a LinkedSpending equivalent.
-The discrepancy is caused by loss in several stages.
-To prevent timeouts and to reduce the impact of disrupted connections, the source data set is downloaded in several parts with a maximum number of entries.
-These parts are then merged so that each file corresponds to exactly one data set.
-Data sets without observations are removed and the remaining data sets are transformed, noting the missing values for all component properties.
-If the first 1000 values are all missing, the transformation is aborted, otherwise a \resource{lso:completeness} value $c = \frac{|\textnormal{existing values}|}{|\textnormal{observations}||\textnormal{component properties}|}$ is attached to the data set.
-Besides empty or non-existing data sets, there were no other types of error observed.
-%here are however several cases of component properties with the same name which raises the problem of determining equivalent component properties.
-The chosen approach is to regard as equal all properties with exactly the same name.
-%A%ll data sets, with the exception of ones where the first 1000 expected values are all missing, are now fully converted and each dataset now has a property “” which has an assigned float value measuring 1-%(missing number of values)/(expected number of values).
-
-%In order to detect faulty data sets and to guarantee high quality data, there is a threshold on the number of errors.
-%If the error count of the transformation of a dataset reaches both at least 30 and \SI{10}{\%} of all observation-property pairs yet processed, the transformation is aborted, which removes 5 additional data sets, leaving \numberofdatasets{}.
-
-\paragraph{Sustainability}
-The data conversion process is controlled by a web application\footnotemark{}, which constantly checks for added and modified data sets from OpenSpending, which are automatically queued for conversion but can also be manually managed.
-%\footnotetext{Available at \url{http://linkedspending.aksw.org/api}, developed at \github{}}
-\footnotetext{\url{http://linkedspending.aksw.org/api}}
-Updates don't interrupt the accessibility of the SPARQL endpoint and the services building on it.
-On average, about 50 new data sets became available on each month between September 2013 and March 2014.
-A service monitor constantly checks the state of the application and reports errors.
-
-%thus the OntoWiki and CubeViz instances. % würde ich weglassen, da OntoWiki+CubeViz erst im nächsten Kapitel richtig erklärt werden
-
-%(\todo{warten auf publikation von michas paper} \remark{JL: Ich kenne den Status nicht, aber auf andere Publikationen warten, ist selten eine gute Idee.})
-%This facilitates the consumation of the data sets.
-
-%Observations can also aggregate measurements and thus allow views of different detail level over the same base data.
-%The dimensions are either data type properties, like \enquote{xsd:dateTime} or coded properties which have \emph{code lists} detailing their possible values.
-%The values can be organized in hierarchical code lists, where each parent property (and code list) aggregates the values of its children.
-
-\paragraph{Performance}
-The transformation of a data set takes less than an hour on average on a 2 GHz virtual machine, using less then 2 GB of RAM.
-
-\section{Publishing}\label{sec:publishing}
-The data is published using OntoWiki~\cite{ontowiki_www}.
-The interface for human and machine consumption is available at \url{http://linkedspending.aksw.org}.
-Depending on the actor and needs, OntoWiki provides various abilities to gather the published RDF data.
-\begin{figure}[b!]
-\includegraphics[width=\columnwidth]{img/ontowiki2.png}
-\caption{View of the data set \dataset{berlin\_de} in the OntoWiki}
-\label{fig:ontowiki}
-\end{figure}
-%The data can be explored and (with the appropriate access rights) authored collaboratively \todo{Die Daten kann man zwar editieren, aber sie werden beim nächsten Durchlauf überschrieben, oder?} (see \autoref{fig:ontowiki}).
-It can be explored by viewing the properties of a resource, its values and by following links to other resources (see \autoref{fig:ontowiki}).
-Using the SPARQL endpoint\footnote{\sparql} provided by the underlying \emph{Virtuoso Triple Store}\footnote{\url{http://virtuoso.openlinksw.com}}, actors are able to satisfy complex information needs.
-%\footnote{According to benchmarks~\cite{Bizer2009, MOR+11, MOR+12} Virtuoso  }
-\begin{figure}[t!]
-\includegraphics[width=\columnwidth]{img/cubeviz_circles.pdf}
-\caption{Faceted browsing in CubeViz by restricting values of dimensions}
-\label{fig:cubeviz_circles}
-\end{figure}
-
-Faceted search offers a selection of values for certain properties and thus slice and dice of the data set according to the interests on the fly.
-For example, depicted in \autoref{fig:cubeviz_circles} is all Greek police spending in a certain region.
-Visualization supports discovery of underlying patterns and gain of new insights about the data, for example about the relative proportions of a budget (see \autoref{fig:cubeviz_piechart}).
-%As can be seen from both depictions, exploration of statistical data using a vocabulary-adopted tool is more efficient and comfortable.
-\todo{"more efficient" im Vergleich zu was?
-Außerdem glaube ich kaum, dass man so etwas anhand von einem Screenshot beurteilen kann.}
-We set up the RDF DataCube Browser CubeViz~\cite{salas-ijsc-2013} as part of the human consumption interface.
-%Both exploration paradigms are already available on OpenSpending, however CubeViz offers data discovery by generating new instead of predefined visualization and takes advantage of the openness of the data by publishing additional information from interlinked data sets.
-
-\begin{table}[hbtp]
-\caption{Technical details of the LinkedSpending data set}
-\label{tab:technicaldetails}
-\tiny
-\renewcommand{\tabcolsep}{5pt}
-\begin{tabularx}{\columnwidth}{
-   >{\hsize=.7\hsize}L 
-   >{\hsize=1.3\hsize}L 
-}
-\toprule
-URL				&\lswebsite\\
-Version date and number		&2013-8-14, 0.1\\
-                      		&2014-4-11, 2014-3\\
-License				&PDDL 1.0\footnotemark\\
-SPARQL endpoint			&\sparql\\
-%VoiD description		&\url{ls:void}\\
-%Schema				&\url{ls:ontology/}\\
-Compressed N-Triples Dump 	&\ntriples\\
-datahub entry			&\url{http://datahub.io/dataset/linkedspending}\\
-\bottomrule
-\end{tabularx}
-\vspace{-10pt}
-\end{table}
-\footnotetext{\url{http://opendatacommons.org/licenses/pddl/1.0/}}
-
-\paragraph{Licensing}
-All published data is openly licensed under the PDDL 1.0. in accordance with the open definition\footnote{\url{http://opendefinition.org/}}.
-%which is summarized by: \todo{Ist es wirklich notwendig die Summary hier zu bringen.
-Der Link reicht doch.} \blockquote{A piece of data or content is open if anyone is free to use, reuse, and redistribute it — subject only, at most, to the requirement to attribute and/or share-alike}.\footnote{\url{http://opendefinition.org/}}
-%The license details are described in a \emph{VoID}~\iffalse ~\footnote{\url{http://www.w3.org/TR/void/}}\fi description
-
-\section{Overview over the Data Sets}\label{sec:overview}
-LinkedSpending consists of \numberofdatasets{} data sets (continually growing) with \numberofobservationsapproximate{} observations total.
-The amount of observations of the individual data sets varies considerably between two (spendings in Prague of about \SI{5000}{CZK} for an unknown purpose) and \SI{242209} (\enquote{Spending from ministries under the Danish government}).
-\Cref{tab:amountofdata} details the average and total amount of data in bytes, triples, and observations as well as the number of links to external data sets, which, for the presented version of 2014-3, amounts to more than 9 million links to LinkedGeoData countries and 1.5 million links to DBpedia currencies.\footnotemark
-\footnotetext{The links are inflated as they originate in observations instead of data sets, which allows better querying and tool support.}
-\Cref{fig:histogram} shows the distribution of the numbers of measures, attributes and dimensions of the data sets.\footnote{This analysis relates to version 0.1, which contains less data sets.}
-Measures represent the quantity that an observation describes.
-All data sets have at least one measure which is the amount of money spent or received.
-For most of them (217) that is the only one but there are data sets with up to 7 measures.
-%\footnote{\dataset{dk-corporate-tax-list} contains measures for different kinds of profits as well as losses.}.
-Attributes give further context to the measurement.
-The number of attributes is more varied, ranging from 2 to 26, with all data sets having at least a currency and a country, and most of them additionally the time the observations refer to.
-While the number of dimensions ranges from none\footnote{There is only one data set with no dimensions which a test data set on OpenSpending, as a data cube with no dimensions is not useful.} to 32, almost all of the data sets have between 1 and 6 dimensions,
-the most common ones being the year and the time the data set and the observations refers to, respectively.
-Technical details about the data sets are described in \cref{tab:technicaldetails}.
-
-\begin{figure}[h]
-\includegraphics[width=0.9\columnwidth]{img/histogram.pdf}
-\caption{Histogram of measures, attributes and dimensions (version 0.1). \num{217} data sets have exactly one measure (clipped bar).}
-\label{fig:histogram}
-\end{figure}
-
-\begin{table}[hbtp]
-\caption{Amount of data for version 2014-3. All values are rounded to the nearest integer.}
-\label{tab:amountofdata}
-\tiny
-\renewcommand{\tabcolsep}{5pt}
-\begin{tabularx}{\columnwidth}{
-   >{\hsize=\hsize}L 
-   >{\hsize=\hsize}L 
-   >{\hsize=\hsize}L 
-}
-\toprule	
-	&\textbf{Total}    		&\textbf{Average}\\ 
-\midrule
-\textbf{number of data sets}    		&\numberofdatasets{}\\
-\textbf{file size (RDF/N-Triples)}	&\SI{24585}{MB}		&\SI{39}{MB}\\
-\textbf{triples}     		    &\num{113640534}&	\num{181245}\\
-\textbf{observations} 			& \num{5026393}		&\num{8017}\\
-\textbf{links to external data sets} 			&\num{10696614}		&\num{17060}\\
-% OLD VERSION
-%\textbf{filesize (RDF/N-Triples)}	&\SI{10012}{MB}		&\SI{41}{MB}\\
-%\textbf{triples}     			&\num{49619572}&	\num{200889}\\
-%\textbf{observations} 			&\num{2404313}		&\num{9734}\\
-%\textbf{links to external data sets} 			&\num{}		&\num{}\\
-\bottomrule
-\end{tabularx}
-\end{table}
-%\subsection{Example Queries}
-\paragraph{Example Queries} 
-%\textbf{Example Queries:} 
-\Cref{tab:sparqlqueries} contains queries for common use cases:
-Queries 1--6 are basic queries.
-Query 7 uses the interlinking to DBpedia currencies by querying over two different graphs.\footnotemark{}
-\footnotetext{Parts of DBpedia and LinkedGeoData describing countries and currencies have been integrated in the SPARQL endpoint.
-With federated querying however, nearly the whole LOD cloud can be queried.}
-Query 8 uses the custom vocabulary\footnotemark{} which is available for each data set.
-\footnotetext{In this case, the \enquote{Hauptfunktion} and \enquote{Oberfunktion} are unique to the \dataset{berlin\_de} dataset.}
-
-\begin{table*}[t]
-\lstset{aboveskip=-8pt,belowskip=-\baselineskip}
-
-\scriptsize
-\caption{Examplary SPARQL queries for typical use cases.}
-\label{tab:sparqlqueries}
-\begin{tabular}{@{\makebox[1em][r]{\rownumber\space}}p{0.2\textwidth}p{0.8\textwidth}}
-\toprule
-\textbf{information need}	&\textbf{SPARQL Query}\\
-\midrule
-list of all data sets&\begin{lstlisting}
-select * {?d a qb:DataSet}
-\end{lstlisting}\\
-all measures of the dataset \dataset{berlin\_de}
-&\begin{lstlisting}
-select ?m { ls:berlin_de qb:structure ?s. ?s qb:component ?c. ?c qb:measure ?m.}
-\end{lstlisting}\\
-all years which have observations in the de-bund dataset from \num{2020} onwards&
-\begin{lstlisting}
-select distinct ?year {?o a qb:Observation. ?o qb:dataSet ls:de-bund. ?o lso:refYear ?year.FILTER (xsd:date(?year) >= "2020-1-1"^^xsd:date) }
-\end{lstlisting}\\
-spendings of more than \num{100} billion \euro&
-\begin{lstlisting}
-select * {?o lso:amount ?a. ?o dbo:currency dbpedia:Euro. FILTER(xsd:integer(?a)>"1E11"^^xsd:integer) }
-\end{lstlisting}\\
-data sets with multiple years&
-\begin{lstlisting}
-select ?d count(?y) as ?count { ?d a qb:DataSet. ?d lso:refYear ?y. } group by ?d having (count(?y)>1)
-\end{lstlisting}\\
-sums of amounts for each reference year of \dataset{berlin\_de}&
-\begin{lstlisting}
-select ?y (sum(xsd:integer(?amount)) as ?sum) 
-{?o qb:dataSet ls:berlin_de. ?o lso:refYear ?y. ?o lso:amount ?amount.} group by ?y
-\end{lstlisting}\\
-data sets with currencies whose inflation rate is greater than \SI{10}{\%}&
-\begin{lstlisting}
-select distinct ?d ?c ?r {?o qb:dataSet ?d. ?o  dbo:currency ?c. ?c dbp:inflationRate ?r. filter(?r > 10)} 
-\end{lstlisting}\\
-Berlin city subsectors of research and education that have had their budget reduced from 2012 to 2013 (data set version 0.1)&
-\begin{lstlisting}
-select ?l (sum(xsd:integer(?amount12)) as ?sum12) (sum(xsd:integer(?amount13)) as ?sum13)
- {?o qb:dataSet ls:berlin_de.
-  ?o lso:Hauptfunktion <http://openspending.org/berlin_de/Hauptfunktion/1>.
-  ?o lso:Oberfunktion ?of. ?of rdfs:label ?l.
-  {?o lso:refYear "2012"^^xsd:gYear. ?o lso:amount ?amount12.} UNION
-  {?o lso:refYear "2013"^^xsd:gYear. ?o lso:amount ?amount13.}}
-group by ?l
-having (sum(xsd:integer(?amount12)) > sum(xsd:integer(?amount13)))
-\end{lstlisting}\\
-data sets ordered by their number of properties in common with \dataset{2012\_tax} (having at least one such common property)&
-\begin{lstlisting}
-select ?d (count(?c) as ?count)
-{ls:2012_tax qb:structure ?s. ?s qb:component  ?c.
- ?d qb:structure ?s2.         ?s2 qb:component ?c.
- FILTER(?d!=ls:2012_tax)}
-group by ?d
-order by desc(?count)
-\end{lstlisting}\\
-\bottomrule
-\end{tabular}
-\end{table*}
-
-%mashups mehrwertgenerierung durch transformation
-
-%bei openspending ist alles da und super aaaber:
-\iffalse
-\section{OpenSpending}
-\begin{table}
-\begin{tabulary}{\columnwidth}{lS}
-\toprule
-\textbf{category}    	&\textbf{number of data sets}\\ 
-\midrule
-$\sum$			&\\
-\bottomrule
-\end{tabulary}
-\caption{Categories of data sets from OpenSpending}
-\end{table}
-\fi
-
-\iffalse
-The data sets are divided into the categories \emph{budget}, \emph{spending} and \emph{other} and are further organized by country and language.
-Each data set has a model that contains its core data set metadata (e.g. name, label, description, currency) and the dimensions, measures and views.
-Each of the models can be transformed to an RDF Data Cube \verb|qb:DataStructureDefinition| and each data set to a \verb|qb:DataSet|.
-Because of the large number of data sets, the models are not transformed manually but with \todo{insert the method here}.
-Model dimensions are mapped to Data Cube dimension, where each of the json object names is manually mapped to an RDF property.
-
-data cube: dimensions, attributes, measures
-\fi
-
-%\end{thebibliography}
-
-\iffalse
-\subsection{Intrinsic dimensions}
-
-\paragraph{Accuracy}
-Because the data sets were created by transformation and the values itself were not changed, their accuracy is equal to that of the source data of each data set.
-A request for information about existing studies into the accuracy on the OpenSpending mailing list was not answered.
-To measure the accuracy, more elaborate research is needed that checks the correctness of a random sample of the values by comparing them with a reference.
-
-\paragraph{Objectivity}
-\todo{das klingt alles viel zu schwammig, aber es ist ja auch nicht klar messbar. wie kann man das trotzdem wissenschaftlich schreiben?}
-As for the accuracy, the objectivity depends on the individual source data provided originally by government agencies and then converted and transfered to OpenSpending by some third party.
-So, assuming that the values were not changed on the way from the agencies to OpenSpending, the amount of thrust that the information is unbiased depends on the thrustworthyness of the governements and agencies themselves.
-Finally, objectivity is a subjective dimension and depends on the use case.
-Using the data to get an overview of the financial situation of a country is surely unproblematic but to allocate funding based on the exact number of the reports is thought to be questionable by the author of the paper.
-%- es gibt auf jeden fall anreize, finanzdaten zu fälschen ()
-\paragraph{Validity-of-documents}
-The validity of the RDF syntax has been successfully verified by the Raptor RDF parsing and serializing utility.
-Valid use of vocabularies is verified by \todo{...}
-
-\paragraph{Consistency}
-The way the program is written guarantees that for each observation there is only value for each measure.\todo{doesnt sound good, how to improve?}
-
-\paragraph{Conciseness}
-In RDF Data Cubes, attributes can be attached to a data set, a slice, a measure property or to an observation.
-
-While attributes which take different values for each observations necessarily have to be attached at the observation level, there are conflicting arguments for attributes which have the same value throughout.
-Attaching them at the observations ensures that all relevant information concerning an observation is in one place.
-Thus users can satisfy their information needs with less complex SPARQL queries or by resolving a single URL.
-Attaching them at the data set however removes redundancy and thus improves the conciseness.
-Furthermore, queries which select certain data sets, are actually easier to formulate with the latter approach.
-However, it is easier to create mash-ups and to use RDF DataCube visualization tools such as CubeViz~\cite{cubeviz}, when all attributes are attached at the observation level, so that we attach all attributes directly to the observations.
-RDF however is flexible and permits attaching the countries additionally at the data sets. Thus datasets of a certain country are simpler and the overhead in data size is negligible.
-
-%The countries are however also at the data sets because queries for data sets of a certain country are anticipated to be more frequent and cross-country spendings are less comparable as intra-country ones.
-%Because of this, our modelling approach is varied:
-%All attributes from the source data which are represented at the observation level, which are different for each data set, are kept so in the target data.
-%It is possible to check if all values of a certain attribute are the same and to move those attributes to the data set level but this is confusing for the users and inconsistent.
-%OpenSpending data sets on the other hand, have two fixed attributes: the countries whose governments do the financial transactions and the currency of the transactions.
-
-%The currency of a transaction is however directly tied to the amount spent and search queries which separate data sets by currency are less anticipated.
-%Thus the overhead of specifying the currency at each observation is deemed to weigh less than the .
-
-\subsection{Trust Dimensions}
-
-- depends on the source data, for each dataset
-most governements - high believability and reputation
-
-\subsection{Dataset Dynamicity Dimensions}
-Because spending data is historical, currency, volatility and timeliness are not applicable.
-\fi
-\subsection{Contextual Dimensions}
-\paragraph{Completeness}
-Because all of the attributes, measures, dimensions and observations of the original data sets are carried over to their Linked Data versions, the all transformed data sets are fully complete both on the schema and the data (instance) level.
-\todo{Nur so ``complete'' wie die Ausgangsdaten, oder?
-Ist auch nicht so interessant die Aussage, da offensichtlich.}
-However of the 321 data sets listed by OpenSpending, only \num{252} were retrievable and \numberofdatasets{} of those convertible.
-
-%\paragraph{Amount-of-data}
-%The amount of observations of the data sets varies between two (spendings in Prague of about \SI{5000}{CZK} for an unknown purpose) and 242209 (\enquote{Spending from ministries under the Danish government}).
-%See \autoref{fig:histogram} for the numbers of measures, attributes and dimensions.
-%See \autoref{fig:amountofdata} for the distribution of the numbers of triples and observations and \autoref{fig:histogram} for the numbers of measures, attributes and dimensions.
-\iffalse
-\begin{figure}[h]
-\includegraphics[width=\columnwidth]{img/pointcloud.pdf}
-\caption{Numbers of triples and observations of the data sets (logarithmic scales)}
-\label{fig:amountofdata}
-\end{figure}
-\fi
-\iffalse
-old data:
-\fi
-
-%\paragraph{Relevancy} This dimension is dependent on the task at hand.
-
-\subsection{Representational Dimensions}
-
-%\todo{- average length of names of measures, attributes,dimensions necessary?}
-
-%The URIs are kept short and human readable by reusing the existing OpenSpending URLs which are constructed based on the names of the resources and their parent resources.
-%For example, the dimension \emph{recipient} of the dataset \emph{nigerianbudg13} has the URL \url{os:nigerianbudg13/recipient}.
-%They were also reused to facilitate future integration with the source dataset using content negotiation to determine whether the user or application needs an RDF representation.
-%For resources which do not have a single source URL the spirit of the OpenSpending naming scheme was still followed. For example, the view \emph{} of the dataset \emph{berlin_de} \todo{example}.
-
-%\paragraph{Representational Consistency}
-\iffalse moved to extraction workflow
-First and foremost, the RDF Data Cube vocabulary provides the backbone structure for all of the data sets.
-This vocabulary uses the model of the Statistical Data and Metadata eXchange (SDMX) initiative\footnote{\url{http://sdmx.org}} which is widely used and, among others, sponsored by the United Nations Statistics Division, the International Monetary Fund and the World Bank.
-RDF Data Cube is supported by tools such as the CubeViz~\cite{Auer+ISWC-2012} plugin, that enables the management of RDF Data Cubes in the OntoWiki (\todo{warten auf publikation von michas paper} \remark{JL: Ich kenne den Status nicht, aber auf andere Publikationen warten, ist selten eine gute Idee.}), which manages several parts of the the Linked Data Lifecycle, such as publication, authoring and visualization.
-This facilitates the consumation of the data sets.
-\fi
-%\todo{Micha: sind die Zitierungen richtig?}
-%The SDMX vocabularies\footnote{\url{http://purl.org/linked-data/sdmx/2009{/dimension\#,measure\#,attribute\#,concept\#}}} are used for observation values, currencies and time periods.
-%\todo{move lifecycle part to a more fitting place?}
-%\paragraph{Representational Conciseness}
-\paragraph{Representational Conciseness and Understandability}
-Because the source data contains labels for all data sets, attributes, dimensions and measures, those labels are also provided for the transformed data sets.
-The resource URLs\footnote{except the observations whose last part is a hash value} are concise and  easily understandable, because they are generated from the names of the entities they describe.% (see \emph{Representational conciseness}).
-This facilitates writing SPARQL queries (see \autoref{tab:sparqlqueries}) and interpreting the results.
-
-%\begin{tabular}{ll}
-%x&\begin{Verbatim}dfdf\end{Verbatim}
-%\end{tabular}
-
-\paragraph{Interpretability}
-All objects are represented by globally unique URIs, many of them with labels (see the paragraph above). \todo{Leider ebenfalls keine so interessante Aussage.}
-
-\paragraph{Versatility}
-The data sets are available as resolved URIs, on a SPARQL endpoint\footnote{\sparql} and as \iffalse Turtle\footnote{\turtle} and \fi N-Triples\footnote{\ntriples} dumps.
-\todo{Wurde ebenfalls schon vorher gesagt.}
-
-%\section{Examples and Critical Discussion of Typical Knowledge Modeling Patterns Used}
-\iffalse
-\section{Use Cases}
-
-- CubeViz visualisierung
-- 2 länder mit ähnlichen data sets -> zusammen analysieren
-4 use cases sind am besten mit sparql queries
-\fi
-\end{comment}
-
-\section{Related Work}\label{sec:relatedwork}
-The TWC Data-Gov Corpus~\cite{datagov1,datagov2} consists of linked government data from the Data-gov project.
-However, it only contains transactions made in the US and does not overlap with OpenSpending.
-The publicspending.gr project generates and publishes~\cite{spendinggreece} public spending data from Greece based on the UK payment ontology and without using statistical data cubes.
-The UK government expenditure dataset COINS\footnote{\url{http://data.gov.uk/dataset/coins}} is available as Linked Data\footnote{\url{http://openuplabs.tso.co.uk/sparql/gov-coins}, in a beta version}.
-\emph{LOD Around-The-Clock (LATC)}\footnoteurl{http://latc-project.eu} is a project, which was funded by the European Union (EU) and converted European open government data into RDF.
-One of its outcomes is the FTS\footnoteurl{http://ec.europa.eu/budget/fts}~\cite{martin-fts} project, which transforms and publishes financial transparency data on EU spending.
-In comparison with LinkedSpending, those projects also contribute linked government data but with a different or more limited scope.
-
-Furthermore, there is the Digital Agenda Scoreboard~\cite{martin-scoreboard} is an EU project which keeps track of the transformation of statistical data to RDF.
-
-\section{Conclusions, Shortcomings, Future Work}\label{sec:conclusions}
-As shown in \Cref{sec:conversion}, we converted several hundreds of financial data sets to RDF and, as shown in \Cref{sec:publishing}, we published them as Linked Open Data in several ways.
-%Thus, our contribution is that we have made available as Linked Open Data a large number of data sets.
-However, we recognise a few shortcomings and our goal is to enrich the meta data with the help of domain experts and to refine the structure of the individual data sets.
-Furthermore, we plan to improve the automatic configuration of CubeViz.
-%improve the user interface \todo{``improve the user interface'' wirkt aus dem Zusammenhang gerissen}
-
-\paragraph{Multilinguality}
-RDF itself provides support for multilingualism, which is one of its key advantages to other representation formats.
-The source data does not contain language tags, however, and the the languages used do not always match the country, the data refers to.
-Automatic language detection on single labels did not yield a satisfying precision and it is not possible to increase the precision of the language detection by combining the estimates about several labels of an observation as their language is not always identical.
-We plan statistical examinations of the relations between labels of different entities and more complex schemes based on those examinations, which can achieve language detection with a higher precision.
-Additionally, we plan to automatically translate all literals to several languages.
-
-% \section{Known Shortcomings and Future Work}
-\paragraph{Individual Modelling}
-Because the source data is already structured, the transformation of all the data sets without the need of text extraction and in an automatic way was feasible.
-On a deep level however, there is much unmodelled structure that is unique to each data set or at most shared between several of them, for instance the categorization of spending into several specific \enquote{plans} in German budgets.
-Because of the amount of data sets, modelling all details, and thus also improving the internal and external connectivity, requires either a large-scale cooperation or a crowd-driven approach, which we did not perform yet.
-%\todo{Ich denke idealerweise sollte ein SPARQL-Query genau so einen Fall aufzeigen, bei dem wir das modelliert haben und damit über mehrere data sets Abfragen bauen konnten.}
-
-\paragraph{Drilldowns}
-Because of the hierarchical organization of the different coded properties \enquote{groups} and \enquote{functions}, the visualizations on \url{openspending.org} permit \enquote{zooming} (drilldown) in and out of the different levels of the data.
-The RDF Data Cube vocabulary specifies the use of \url{skos:ConceptScheme} or \url{qb:HierarchicalCodeList} but neither variant is fully implemented and it is not clear, which of those modelling possibilities will become standard.
-%\todo{Ist hier ``not implemented in CubeViz'' gemeint?}  -> das kann auch sein aber es ist auch noch nicht in den Daten selbst drin
-%As soon as hierarchical code lists are better specified, the linear hierarchy of the budget groups and function should be modelled.
-
-\subsection{Future Work}
-
-% interlinking
-\paragraph{Interlinking}
-Extensive interlinking of referenced entities to the all-purpose knowledge base of DBpedia provides additional context.
-Coded property values, such as the budget areas healthcare and public transportation, can be interlinked with their respective DBpedia concepts.
-This enables the usage of type hierarchies and thus new ways of structuring the data and provides more meaningful aggregations and new insights.
-
-\paragraph{Question Answering}
-We plan to develop a question answering system that allows accessing statistical Linked Data in the form of RDF Data Cubes using natural language questions~\cite{statisticalqashort}.
-LinkedSpending is used both as the first knowledge base and for performance evaluation.
-
-%\paragraph{more elaborate updating with different modules with a diff component a scheduler ...}
-
-\iffalse
-\paragraph{Reuse}
-Because of the large amount of data sets, the conversion process is necessarily automatic.
-While for some of the often used component properties, like date and time, resources from existing vocabularies such as SDMX are used, in the general case, new resources are created for all values.
-A thorough manual inspection might yield in an increased usage of existing vocabulary.
-Also some of the dimensions may be stable over different data sets and thus might be shared, which decreases the amount of redundancy and increases the benefit of those resources, should they be referenced from other resources.
-
-\paragraph{Hosting integration with OpenSpending}
-According to the Linked Data principles, the RDF resources we created should be resolvable.
-As openspending.org already uses URLs with different extensions to deliver different content types like JSON and HTML, adding content negotiation and hosting the resources at there would be easy to implement.
-To speed up the integration of new data, a web service that watches for changes in the input data and runs the transformation to keep the RDF data up to date would be useful. 
-%\paragraph{Webservice}
-\fi
-
-\iffalse
-\begin{table}
-\begin{tabulary}{\columnwidth}{LSS}
-\textbf{Country}	&\textbf{Number of data sets}	&\textbf{Number of observations}\\
-\end{tabulary}
-\caption{spending data per country}
-\label{tab:countries}
-\end{table}
-\fi
-%\todo{write somewhere: county codes modelled with SDMX Codelist CL_AREA following ISO 3166 alpha-2(for example \enquote{DE} stands for Germany) -> better: link with linkedgeodata}
-\iffalse
-\begin{table}
-\begin{tabulary}{\columnwidth}{LS}:
-\textbf{}	&	&\\
-\end{tabulary}
-\caption{spending data per country}
-\label{tab:countries}
-\end{table}
-\paragraph{URL renaming}
-The URLs of most of the transformed resources are equal (except for the prefixes) to those of OpenSpending in order to facilitate a common hosting of both RDF and the original resources with content negotiation.
-However the original naming schemes lead less understandable to URLs like the dimension \url{ls:nigeriabudg13/recipient} which cannot be prefixed at the core schema level because RDF local names cannot contain the slash character \enquote{/}.
-If this kind of hosting is not wanted by OpenSpending, other naming schemes are better, for example \ls\url{/dimensions/recipient}.
-\fi
-% \section*{Acknowledgements}
-\medskip\textbf{Acknowledgements:}
-Special thanks goes to the people behind the OpenSpending project, including Friedrich Lindenberg for suggesting the conversion.
-
-% \cite{hoeffner-2013-kesw}
-\bibliographystyle{abbrv}
-\bibliography{paper,../../bib/aksw.swj}
-%\bibliography{paper,../../bib/aksw.twocolumn.abbrv,../../bib/aksw.pending}
+\section{Acknowledgments}
+We thank the publishers ... and ... for granting us permission to publish the subontologies based on the textbooks [][][] under an open license.
+The SNIK project is supported by the DFG (German Research Foundation) under the grant numbers 1605/7-1 and 1387/8-1.
+\nocite{*} 
+\bibliographystyle{ios1}
+\bibliography{paper}
 

### 2019-03-01 paper.tex
Add references and extend structure.

@@
 \begin{abstract}
-Abstract text.
+This paper describes the ontology of the \textbf{Semantic Network of Information Management in Hospitals} (SNIK).  
 \end{abstract}
@@
 
-\section{Background}
+\section{Background}\label{sec:background}
 \todo{not Konrad: write background of meta model}
+\citet{domaene} contains an initial structure of the meta model.
 
-\section{Architecture}
+\section{Architecture}\label{sec:architecture}
+\todo{Describe the meta model here}
 The \textbf{Semantic Network of Information Management in Hospitals} (\enquote{Krankenhaus} in German) is an OWL 2 DL ontology that consists of a central meta model and modular subontologies based on the meta model.
 
-\section{Data Set Description}
+\section{Data Set Description}\label{sec:dsd}
+The SNIK ontologies are available at \url{http://www.snik.eu/ontology/} under the Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International license.
+The SNIK ontologies contain knowledge about information management in hospitals from three textbooks, an interview and a standard.
+- Blue Book~\citep{bb}
+- Orange Book~\citep{ob}
+- CIOX
+- Heinrich~\citep{he}
+- IT4IT
 
-\section{Application}
+\section{Application}\label{sec:application}
+The SNIK ontologies were initially intended as the basis for a decision support system for hospital CIOs~\citep{domaene}.
+While such a system was designed and a prototype was implemented, it was ultimately not integrated into the informational infrastructure of the Unilklinikum Leipzig.
+
+- verweis auf existierende papers
 

### 2019-03-04 paper.tex
Add todos.

@@
 \todo{Describe the meta model here}
-The \textbf{Semantic Network of Information Management in Hospitals} (\enquote{Krankenhaus} in German) is an OWL 2 DL ontology that consists of a central meta model and modular subontologies based on the meta model.
+The \textbf{Semantic Network of Information Management in Hospitals} (\enquote{Krankenhaus} in German) is an OWL 2 DL ontology that consists of a central meta model and modular subontologies. 
+The SNIK ontologies contain knowledge about information management in hospitals from different sources:% three textbooks, an interview and a standard.
+The textbooks \citet{bb} \citet{ob} and \citet{he} form the ontologies BB, OB and HE, respectively. 
+CIOX is based on an interview with the CIO of the Universitätsklinikum Leipzig.
+IT4IT is based on the IT4IT-Standard.\todo{Sebastian: Was über Standards und IT4IT schreiben}
 
@@
 The SNIK ontologies are available at \url{http://www.snik.eu/ontology/} under the Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International license.
-The SNIK ontologies contain knowledge about information management in hospitals from three textbooks, an interview and a standard.
-- Blue Book~\citep{bb}
-- Orange Book~\citep{ob}
-- CIOX
-- Heinrich~\citep{he}
-- IT4IT
 
@@
 The SNIK ontologies were initially intended as the basis for a decision support system for hospital CIOs~\citep{domaene}.
-While such a system was designed and a prototype was implemented, it was ultimately not integrated into the informational infrastructure of the Unilklinikum Leipzig.
+While such a system was designed and a prototype was implemented\todo{welches paper ist das?}, it was ultimately not integrated into the informational infrastructure of the Unilklinikum Leipzig.
 
-- verweis auf existierende papers
+\todo{verweis auf existierende papers}
+\todo{benutzung durch braunschweig}
+\todo{benutzung durch amsterdam}
 

### 2019-03-07 paper.tex
Extend paper.

@@
 \usepackage{todonotes}
+%\usepackage{natbib} % does not seem to work with ios 1
+\renewcommand{\citet}{\cite}% citet is not defined without natbib
+\renewcommand{\citep}{\cite}% citep is not defined without natbib
 \setlength{\marginparwidth}{2.1cm}% enough space for todonotes
@@
 %\lstset{language=SPARQL,breaklines=true}
-\usepackage{cleveref}
-
+%\usepackage{cleveref}
 %%%%%%%%%%% End of definitions
@@
 
+\section{Call Excerpt}
+Source: \url{http://www.semantic-web-journal.net/authors}\\
+Short papers describing ontology modeling and creation efforts.
+The descriptions should be brief and pointed, indicating the design principles, methodologies applied at creation, comparison with other ontologies on the same topic, and pointers to existing applications or use-case experiments.
+It is strongly encouraged, that the described ontologies are free, open, and accessible on the Web.
+These submissions will be reviewed along the following dimensions:
+\begin{enumerate}
+\item Quality and relevance of the described ontology (convincing evidence must be provided).
+\item Illustration, clarity and readability of the describing paper, which shall convey to the reader the key aspects of the described ontology.
+\end{enumerate}
+
 \section{Background}\label{sec:background}
@@
 \todo{Describe the meta model here}
-The \textbf{Semantic Network of Information Management in Hospitals} (\enquote{Krankenhaus} in German) is an OWL 2 DL ontology that consists of a central meta model and modular subontologies. 
-The SNIK ontologies contain knowledge about information management in hospitals from different sources:% three textbooks, an interview and a standard.
-The textbooks \citet{bb} \citet{ob} and \citet{he} form the ontologies BB, OB and HE, respectively. 
+The \textbf{Semantic Network of Information Management in Hospitals} (SNIK\footnote{Hospital means \enquote{Krankenhaus} in German.}) is a modular OWL 2 DL ontology.
+The \enquote{Meta Model} defines three basic disjunctive classes and their possible relations: Roles (who), Function (does what) and Entity Types (and which information is therefore needed).
+A set of modular subontologies define subclasses of those three classes and their relations as described by a certain knowledge source about information management in hospitals:
+%from different sources:% three textbooks, an interview and a standard.
+The textbooks \citet{bb}, \citet{ob} and \citet{he} form the ontologies BB, OB and HE, respectively. 
 CIOX is based on an interview with the CIO of the Universitätsklinikum Leipzig.
-IT4IT is based on the IT4IT-Standard.\todo{Sebastian: Was über Standards und IT4IT schreiben}
+IT4IT is based on the IT4IT standard.\todo{Sebastian: Was über Standards und IT4IT schreiben}
 
@@
 \section{Application}\label{sec:application}
-The SNIK ontologies were initially intended as the basis for a decision support system for hospital CIOs~\citep{domaene}.
-While such a system was designed and a prototype was implemented\todo{welches paper ist das?}, it was ultimately not integrated into the informational infrastructure of the Unilklinikum Leipzig.
+The SNIK ontologies were initially intended~\citep{domaene} as the data source for \textbf{software to support hospital CIOs}:
+(1) The requirements engineering decision support system TOREOnto~\citep{toreonto} and (2) the knowledge exploration and navigation visualization CIONo~\citep{ciono}.
+While those applications were designed and prototypes were implemented, they were ultimately not integrated into the informational infrastructure of the Unilklinikum Leipzig.
 
+Recent efforts focus on ontology assisted \textbf{teaching}. 
+SNIK Graph~\citep{snikgraph}
 \todo{verweis auf existierende papers}
@@
 \section{Acknowledgments}
-We thank the publishers ... and ... for granting us permission to publish the subontologies based on the textbooks [][][] under an open license.
+We thank the publishers ... and ... for granting us permission to publish the subontologies based on the textbooks \citet{bb,ob,he} under an open license.
 The SNIK project is supported by the DFG (German Research Foundation) under the grant numbers 1605/7-1 and 1387/8-1.
-\nocite{*} 
+%\nocite{*} 
 \bibliographystyle{ios1}

### 2019-04-09 paper.tex
Add missing commas.

@@
 \thanks{Corresponding author. \printead{e1}.}},
-\author[A]{\fnms{Franziska} \snm{Jahn}\ead[label=e2]{franziska.jahn@imise.uni-leipzig.de}}
-\author[A]{\fnms{Birgit} \snm{Schneider}\ead[label=e3]{birgit.schneider@imise.uni-leipzig.de}}
-\author[A]{\fnms{Anna} \snm{Lörke}\ead[label=e4]{anna.loerke@imise.uni-leipzig.de}}
-\author[A]{\fnms{Thomas} \snm{Pause}\ead[label=e5]{thomas.pause@imise.uni-leipzig.de}}
+\author[A]{\fnms{Franziska} \snm{Jahn}\ead[label=e2]{franziska.jahn@imise.uni-leipzig.de}},
+\author[A]{\fnms{Birgit} \snm{Schneider}\ead[label=e3]{birgit.schneider@imise.uni-leipzig.de}},
+\author[A]{\fnms{Anna} \snm{Lörke}\ead[label=e4]{anna.loerke@imise.uni-leipzig.de}},
+\author[A]{\fnms{Thomas} \snm{Pause}\ead[label=e5]{thomas.pause@imise.uni-leipzig.de}},
 \author[A]{\fnms{Alfred} \snm{Winter}\ead[label=e6]{alfred.winter@imise.uni-leipzig.de}}
@@
 SNIK Graph~\citep{snikgraph}
+
 \todo{verweis auf existierende papers}

### 2019-09-10 paper.tex
Change focus to dataset description paper.

@@
 \usepackage{booktabs}
+\usepackage{tabulary}
 \usepackage{csquotes}
+\usepackage{siunitx}
 %\usepackage{graphicx}
@@
 \setlength{\marginparwidth}{2.1cm}% enough space for todonotes
-%\usepackage{listings}
-%\lstset{language=SPARQL,breaklines=true}
+\usepackage{listings}
+\lstset{language=SPARQL,breaklines=true}
 %\usepackage{cleveref}
@@
 \title{The SNIK Ontology of Hospital Information Management}
-\runningtitle{SNIK}
+\runningtitle{The SNIK Ontology of Hospital Information Management}
 %\subtitle{}
@@
 \author[A]{\fnms{Alfred} \snm{Winter}\ead[label=e6]{alfred.winter@imise.uni-leipzig.de}}
-\runningauthor{}
+\runningauthor{K. Höffner et al.}
 \address[A]{Institute for Medical Informatics, Statistics and Epidemiology (IMISE),
@@
 
-\section{Call Excerpt}
-Source: \url{http://www.semantic-web-journal.net/authors}\\
-Short papers describing ontology modeling and creation efforts.
-The descriptions should be brief and pointed, indicating the design principles, methodologies applied at creation, comparison with other ontologies on the same topic, and pointers to existing applications or use-case experiments.
-It is strongly encouraged, that the described ontologies are free, open, and accessible on the Web.
-These submissions will be reviewed along the following dimensions:
-\begin{enumerate}
-\item Quality and relevance of the described ontology (convincing evidence must be provided).
-\item Illustration, clarity and readability of the describing paper, which shall convey to the reader the key aspects of the described ontology.
-\end{enumerate}
+\section{Introduction}
+Publishing textbook knowledge as Linked Data enables different ways of teaching.
+The data is the result of 
+The extraction effort includes 
+The data is continously revised: users constantly report wrong or missing data in the visualization, which is then corrected, respectively revised, by the researchers of the project.
 
-\section{Background}\label{sec:background}
-\todo{not Konrad: write background of meta model}
+This paper describes the current state of digitalized Hospital Information Management textbook knowledge.
+
+Using a data model of Information Management 
+Depending on the learning style and information need of the student, several tools are available to cosnume that knowledge.
+
+\section{Unsorted}
+
+\begin{tabular*}{\columnwidth}{l}
+Which components of a healthcare network are not healthcare institutions?\\
+\begin{lstlisting}
+SELECT ?x {bb:HealthCareNetwork meta:entityTypeComponent ?x.
+  MINUS {?x rdfs:subClassOf* bb:HealthCareInstitution.}}
+\end{lstlisting}\\
+%bb:GovernmentalAuthority\\
+
+How many functions is the CIO responsible for?\\
+\begin{lstlisting}
+SELECT COUNT(?f) {bb:ChiefInformationOfficer meta:isResponsibleForFunction ?f.}
+\end{lstlisting}\\
+%15\\
+
+Do transinstitutional health information systems support telemicroscopy?\\
+\begin{lstlisting}
+ASK {bb:TransinstitutionalHealthInformationSystem meta:supports bb:Telemicroscopy.}
+\end{lstlisting}
+\end{tabular*}
+
+\begin{tabulary}{\columnwidth}{lL}
+\toprule
+URL		&\url{http://www.snik.eu/ontology}\\
+Version		&2018-10-10, 0.4.1\\
+License		&CC BY-NC-SA 4.0\\
+SPARQL Endpoint	&\url{http://www.snik.eu/sparql}\\
+Visualization	&\url{http://www.snik.eu/graph}\\
+RDF Browser	&\url{http://www.snik.eu/ontology}\\
+Download	&\url{https://github.com/IMISE/snik-ontology/releases/download/0.8.0/snik-0.8.zip}\\
+\# Triples	&\num{112747}\\
+\# Classes	&\num{4729}\\
+\# Properties	&\num{329}\\
+\# Interlinks	&\num{713}\\
+%Subontologies	&\url{http://www.snik.eu/ontology/bb}\\
+%		&\url{http://www.snik.eu/ontology/ciox,}\\
+%		&\url{http://www.snik.eu/ontology/ob}\\
+%		&\url{http://www.snik.eu/ontology/he}\\
+%		&\url{http://www.snik.eu/ontology/it4it}\\
+\bottomrule
+\end{tabulary}%The source code for the services is available at \url{https://github.com/imise}
+
+\includegraphics[width=\columnwidth]{img/metamodel9s.pdf}
+\includegraphics[width=\columnwidth]{img/hierarchy.pdf}
+\begin{center}
+\begin{tabular*}{0.96\columnwidth}{lll}
+\toprule
+\textbf{Ontology}				&\textbf{Prefix}&\textbf{Source}\\
+\midrule
+\url{http://www.snik.eu/ontology/meta}		&meta		&All Sources\\
+\url{http://www.snik.eu/ontology/bb}		&bb		&Textbook~\cite{bb}\\
+\url{http://www.snik.eu/ontology/ob}		&ob		&Textbook~\cite{ob}\\
+\url{http://www.snik.eu/ontology/he}		&he		&Textbook~\cite{he}\\
+\url{http://www.snik.eu/ontology/ciox}		&ciox		&CIO Interview\\
+\url{http://www.snik.eu/ontology/it4it}		&it4it		&Standard~\cite{it4it}\\
+\bottomrule
+\end{tabular*}
+\end{center}
+
+\section{Sources}\label{sec:sources}
+Three textbooks provide different views on the domain of Hospital Information Management:
+\citet{bb} presents a broad view on \enquote{typical architectures of health information systems and their systematic strategic management}.
+\citep{ob} concentrates on the \emph{tactical} management of information systems in general and on healthcare in particular.
+The focus of tactical management lies on the planning and operation of projects.  
+\citet{he} explains information management beyond the scope of healthcare. 
+CIOX is based on an interview with the CIO of the Universitätsklinikum Leipzig.
+IT4IT is based on the IT4IT standard.\todo{Sebastian: Was über Standards und IT4IT schreiben}
+
+\section{Data Model}\label{sec:architecture}
 \citet{domaene} contains an initial structure of the meta model.
-
-\section{Architecture}\label{sec:architecture}
 \todo{Describe the meta model here}
@@
 %from different sources:% three textbooks, an interview and a standard.
-The textbooks \citet{bb}, \citet{ob} and \citet{he} form the ontologies BB, OB and HE, respectively. 
-CIOX is based on an interview with the CIO of the Universitätsklinikum Leipzig.
-IT4IT is based on the IT4IT standard.\todo{Sebastian: Was über Standards und IT4IT schreiben}
 
@@
 (1) The requirements engineering decision support system TOREOnto~\citep{toreonto} and (2) the knowledge exploration and navigation visualization CIONo~\citep{ciono}.
-While those applications were designed and prototypes were implemented, they were ultimately not integrated into the informational infrastructure of the Unilklinikum Leipzig.
+While those applications were designed and prototypes were implemented, they were ultimately not integrated into the informational infrastructure of the Uniklinikum Leipzig.
 

### 2019-09-11 paper.tex
Add content from rejected medinfo paper. Start concessions to practicality.

@@
 \usepackage{listings}
+\usepackage{aurl}
 \lstset{language=SPARQL,breaklines=true}
@@
 %\pretitle{}
-\title{The SNIK Ontology of Hospital Information Management}
-\runningtitle{The SNIK Ontology of Hospital Information Management}
+\title{Textbook Knowledge on Hospital Information Management as Linked Data}
+%\runningtitle{The SNIK Ontology of Hospital Information Management}
 %\subtitle{}
@@
 \author[A]{\fnms{Birgit} \snm{Schneider}\ead[label=e3]{birgit.schneider@imise.uni-leipzig.de}},
-\author[A]{\fnms{Anna} \snm{Lörke}\ead[label=e4]{anna.loerke@imise.uni-leipzig.de}},
+%\author[A]{\fnms{Anna} \snm{Lörke}\ead[label=e4]{anna.loerke@imise.uni-leipzig.de}},
 \author[A]{\fnms{Thomas} \snm{Pause}\ead[label=e5]{thomas.pause@imise.uni-leipzig.de}},
@@
 \address[A]{Institute for Medical Informatics, Statistics and Epidemiology (IMISE),
-\orgname{University of Leipzig}, \cny{Germany}\printead[presep={\\}]{e1,e2,e3,e4,e5,e6}}
+\orgname{University of Leipzig}, \cny{Germany}\printead[presep={\\}]{e1,e2,e3,e5,e6}}
 %Medical Informatics, Management of Health Information Systemsi
@@
 \begin{abstract}
-This paper describes the ontology of the \textbf{Semantic Network of Information Management in Hospitals} (SNIK).  
+Textbooks contain abstract knowledge about a domain.
+In the case of Hospital Information Management, much of this knowledge consists of systems, their parts and their relations, making it a natural fit for RDF, RDFS and OWL.
+As the domain is large and highly relevant to students of Medical Informatics, modelling the knowledge is not only possible but also very useful.
+We present a common data model for the domain and the result of applying the data model to three textbooks.
+We publish the result over several interfaces that are useful for researchers, practitioners and students, depending on their objectives and their Semantic Web skills.
 \end{abstract}
@@
 \section{Introduction}
+Given a care delivery organization, its health information system (HIS) can be defined as the part of the organization that processes and stores data, information, and knowledge.
+It usually consists of a large number of different application systems, computers, and network components~\citep{bb}.
+Managing an HIS comprises planning, monitoring and directing activities.
+Due to the complexity and the unique conditions in health care, HIS management is an exceptionally challenging task.
+There is an enormous amount of frameworks, textbooks and articles describing the scope of HIS management from the perspective of medical informatics.
+%However, the disciplines of business informatics and information systems (IS) provide an even broader view on information systems and their management.
+A structured representation of the different perspectives leads to a holistic view on HIS management and helps help researchers and students connect their existing knowledge with further knowledge from other sources during research and learning.
+In order to integrate different knowledge sources and to provide the knowledge in a structured, machine-readable data format, we extracted knowledge about HIS management from three textbooks~\citep{bb,ob,he} and other sources.
+The combination of this knowledge~\citep{einsemantischesnetz,domaene,approachtosupport} results in SNIK, the Semantic Network of Information Management in Hospitals (“Krankenhaus” in German).
+
+In order to encourage and enable other researchers, students and health informatics professionals to use available knowledge of HIS management, we answer the following questions in this paper:
+
+We discuss advantages and disadvantages of the interfaces and give recommendations on which ones present the best compromise for different use cases and target audiences.
+We conclude with plans for future work on interlinking and visualization.
+
+As the domain is large and highly relevant to students of Medical Informatics, modelling the knowledge is not only possible but also very useful.
+
+We present a common data model for the domain and the result of applying the data model to three textbooks.
+We publish the result over several interfaces that are useful for researchers, practitioners and students, depending on their objectives and their Semantic Web skills.
+
 Publishing textbook knowledge as Linked Data enables different ways of teaching.
@@
 
+\section{Data Model}
+In order to specify, which information should be extracted from the books and to facilitate comparisons, we use a common data model.
+Because processed textbooks contain abstract knowledge instead of information about any specific hospital, all concepts are modelled as classes.
+We thus call our data model the \enquote{meta model}, as it is an ontology.
+
+The Meta Model (see Figure 1) provides a common vocabulary for the domain of HIS management and thus defines, which superclasses and relations can be used.
+SNIK version 0.4.1 comprises five subontologies that are built upon the Meta Model, see Table 1.
+At the head of the class hierarchy is the “Top” class, which has exactly three disjunctive subclasses.
+Following the Meta Model, each class has to be a subclass of exactly one of them.
+The correct superclass of a new concept can be found by answering the question: Who (“Role”) does what (“Function”) and which information (“EntityType”) is needed?
+If a concept is neither of them, it cannot be modeled using the Meta Model.
+As the subclass relation is transitive, a new class can be placed further down the hierarchy and it can still be inferred, whether it is a Function, Role or EntityType (see Figure 3).
+Besides the subclass relationship, two classes can be connected with relations provided by the Meta Model.
+The generic “is associated with” relationship is possible between all classes but carries little information.
+Which more specific relations between two classes are available depends on their superclasses.
+For example, a role and a function can be connected as “is involved in” “is responsible for” and “approves”.
+Relations that are neither of them can either be modeled by using the generic “is associated with” relation or by creating and using a new sub relation of “is associated with”
+
+\section{Concessions to Practicality}
+Our main focus is that SNIK provides a net benefit to its users, many of which do not have Semantic Web expertise.
+This led to several concessions to usability and actual usage by sacrificing full conformance to the W3C OWL recommendation in the following cases:
+
+\paragraph{OWL Restrictions}
+The meta model offers several properties between the different subtopclasses.
+The definitions of \aurl{rdfs}{domain} and \aurl{rdfs}{range} state that the subjects, respectively objects of all triples with that property are instances of that class.
+Because SNIK contains classes, which are subclasses, not instances of the subtopclasses, using those properties directly in SNIK is technically wrong and OWL property restrictions should be used.
+This presented the following problems:
+\begin{itemize}
+\item Tools such as LodLive don't understand those OWL constructs and only show the RDF form with blank nodes and helper constructs, which is confusing to the users
+\item Communicating OWL restrictions to extractors and non-Semantic Web expert project members is difficult.
+\item Extractors only enter the name of the property, not whether it is a "someValuesFrom" or "allValuesFrom" restriction.
+For example, it is unclear whether each CEO is involved in at least one project review, or whether each CEO is only involved in project reviews.
+It may even be the case that both of those statements are not what the author intended.
+\item Deleting or adding a connection between classes in the OntoWiki is no longer an atomic operation and may result in unreferenced blank nodes or partially restored undo results.
+\end{itemize}
+
+To avoid those problems, we use direct triples between the classes of SNIK as if they were instances, not classes.
+
+\paragraph{}
+
+\section{Data Quality}
+- extractors not always same people who wrote book/part of book
+- mistake when extracting
+- reasoning does not work (test)
+\begin{itemize}
+\item Multiple Subtops
+\item Inconsistent Subtop with Subclass
+\item SKOS Link to Different Subtop
+\item Subclass Cycles
+\item Missing superclass
+\item Undefined Objects
+\item Domain Violation
+\item Range Violation
+\item Class URL Naming Convention Violations
+\item Property URL Naming Convention Violations
+\item Missing Definition
+\item Literals with Semicolons
+\item Classes with too many subclasses
+\item No Restriction
+\item Non-HTTP URI
+\item OWL 2 DL: Undefined Properties
+\item OWL 2 DL: Multiply Defined Properties
+\item OWL 2 DL: Class Typing Constraints
+\end{itemize}
+-> use SHAPE/SHACL ?
+
 \section{Unsorted}
-
-\begin{tabular*}{\columnwidth}{l}
-Which components of a healthcare network are not healthcare institutions?\\
-\begin{lstlisting}
-SELECT ?x {bb:HealthCareNetwork meta:entityTypeComponent ?x.
-  MINUS {?x rdfs:subClassOf* bb:HealthCareInstitution.}}
-\end{lstlisting}\\
-%bb:GovernmentalAuthority\\
-
-How many functions is the CIO responsible for?\\
-\begin{lstlisting}
-SELECT COUNT(?f) {bb:ChiefInformationOfficer meta:isResponsibleForFunction ?f.}
-\end{lstlisting}\\
-%15\\
-
-Do transinstitutional health information systems support telemicroscopy?\\
-\begin{lstlisting}
-ASK {bb:TransinstitutionalHealthInformationSystem meta:supports bb:Telemicroscopy.}
-\end{lstlisting}
-\end{tabular*}
 
@@
 
+%The technical environment of SNIK is characterized in previous work [11].
+\section{Discussion}
+SNIK is based on the Meta Model from Figure \todo{reference}.
+Thus, SNIK can be considered as an archetype for ontologies describing for a given domain (here “HIS management”), what functions a certain role has to carry out and what information a person of this role needs and what information she/he provides while carrying out the functions.
+This may be used for domains like the medical department of a hospital as well.
+This would be valuable knowledge, for instance as basis for systems analysis projects [1].
+Different user groups can use the interfaces of SNIK that are most suitable to them.
+For example, the IT management of a hospital can use SNIK as a vocabulary to integrate data from different formats that result from application components from different vendors.
+Using the SPARQL1 query language, experts can also integrate applications with the SNIK ontology.
+ Experts can also use the SPARQL query editor to answer specific questions, see Table 2.
+Students and teachers, on the other hand, benefit more from SNIK Graph, which intuitively presents knowledge without the need for technical skills.
+
+Figure 4– A “Spider Worm” between the classes “Change Management” and “HIS Quality” in the SNIK Graph visualization web application.
+The content of SNIK is taken from textbooks which are protected by property rights of publishing houses or taken from an interview with a CIO holding property rights as well.
+Making the SNIK ontology openly available may therefore conflict with these property rights.
+In case of SNIK, we appreciate the approvals of the publishing houses Springer, de Gruyter and Schattauer and of the CIO.
+In general, missing approvals may prevent the creation of such open ontologies and thus the provision of textbook content as open knowledge.
+The approvals for SNIK were possible because for each class and relation being part of SNIK there is a clear indication of its source (see code “bb” and chapter information in Table 1).
+In future work we will use these indications for establishing hyperlinks into the sources such that users of SNIK may use SNIK as a smart index for retrieving more detailed knowledge and supporting illustrations from the original textbooks – if digitally available.
+We are convinced that this way an ontology like SNIK is not a competitor of textbooks but a smart guide for using a textbook.
+Thus, it might be attractive for publishing houses.
+
 \todo{verweis auf existierende papers}
@@
 \todo{benutzung durch amsterdam}
+
+\section{Conclusions}
+We showed that knowledge on the management of information systems in medicine and health care is made publicly available using open standards over several interfaces with different compromises between expressivity and accessibility.
+It can be combined with other knowledge in biomedical and health informatics and in other disciplines.
+
+\section{Future Work}
+We plan to extend the text-based interlinks with semantic-similarity-based ontology matching [5] to find more correspondencies between the textbooks.
+This will enable users to switch to another textbook when a certain concept is explained there in more detail.
+The SNIK toolset, SNIK Graph in particular, is under permanent improvement.
+For instance, we are considering alternatives to the shortest path that are the most preferred by different audiences and whether those paths can be (semi-)automatically generated using suitable node or edge weight functions.
+We are also investigating candidates for additional subontologies and as interlink targets.
 

### 2019-09-16 paper.tex
Continue paper.

@@
 \usepackage{aurl}
+\daurl{bb}{http://www.snik.eu/ontology/bb/}
 \lstset{language=SPARQL,breaklines=true}
-%\usepackage{cleveref}
+\usepackage{cleveref}
 %%%%%%%%%%% End of definitions
@@
 In order to integrate different knowledge sources and to provide the knowledge in a structured, machine-readable data format, we extracted knowledge about HIS management from three textbooks~\citep{bb,ob,he} and other sources.
-The combination of this knowledge~\citep{einsemantischesnetz,domaene,approachtosupport} results in SNIK, the Semantic Network of Information Management in Hospitals (“Krankenhaus” in German).
+The combination of this knowledge~\citep{einsemantischesnetz,domaene,approachtosupport} results in SNIK, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German).
 
@@
 
-The Meta Model (see Figure 1) provides a common vocabulary for the domain of HIS management and thus defines, which superclasses and relations can be used.
+The Meta Model (see \cref{fig:metamodel}) provides a common vocabulary for the domain of HIS management and thus defines, which superclasses and relations can be used.
 SNIK version 0.4.1 comprises five subontologies that are built upon the Meta Model, see Table 1.
-At the head of the class hierarchy is the “Top” class, which has exactly three disjunctive subclasses.
+At the head of the class hierarchy is the \enquote{Top} class, which has exactly three disjunctive subclasses.
 Following the Meta Model, each class has to be a subclass of exactly one of them.
-The correct superclass of a new concept can be found by answering the question: Who (“Role”) does what (“Function”) and which information (“EntityType”) is needed?
+The correct superclass of a new concept can be found by answering the question: Who (\enquote{Role}) does what (\enquote{Function}) and which information (\enquote{EntityType}) is needed?
 If a concept is neither of them, it cannot be modeled using the Meta Model.
@@
 Besides the subclass relationship, two classes can be connected with relations provided by the Meta Model.
-The generic “is associated with” relationship is possible between all classes but carries little information.
+The generic \enquote{is associated with} relationship is possible between all classes but carries little information.
 Which more specific relations between two classes are available depends on their superclasses.
-For example, a role and a function can be connected as “is involved in” “is responsible for” and “approves”.
-Relations that are neither of them can either be modeled by using the generic “is associated with” relation or by creating and using a new sub relation of “is associated with”
+For example, a role and a function can be connected as \enquote{is involved in} \enquote{is responsible for} and \enquote{approves}.
+Relations that are neither of them can either be modeled by using the generic \enquote{is associated with} relation or by creating and using a new sub relation of \enquote{is associated with}
 
@@
 
-\includegraphics[width=\columnwidth]{img/metamodel9s.pdf}
+\begin{figure*}
+\caption{The SNIK Meta Model}
+\label{fig:metamodel}
+\includegraphics[width=0.8\textwidth]{img/metamodel9s.pdf}
+\end{figure*}
+
 \includegraphics[width=\columnwidth]{img/hierarchy.pdf}
@@
 \section{Application}\label{sec:application}
-The SNIK ontologies were initially intended~\citep{domaene} as the data source for \textbf{software to support hospital CIOs}:
-(1) The requirements engineering decision support system TOREOnto~\citep{toreonto} and (2) the knowledge exploration and navigation visualization CIONo~\citep{ciono}.
-While those applications were designed and prototypes were implemented, they were ultimately not integrated into the informational infrastructure of the Uniklinikum Leipzig.
+SNIK was initially intended~\citep{domaene} as the data source for software to support hospital CIOs.
+Two applications were designed and implemented as prototypes: (1) the requirements engineering decision support system TOREOnto~\citep{toreonto} and (2) the knowledge exploration and navigation visualization CIONo~\citep{ciono}.\todo{add ciono to bib}
+Due to data protection regulations, efforts to integrated them into the informational infrastructure of the Uniklinikum Leipzig were stopped.
+Subsequent efforts shifted to ontology assisted teaching.
 
-Recent efforts focus on ontology assisted \textbf{teaching}. 
-SNIK Graph~\citep{snikgraph}
+\subsection{SNIK Graph}
+\todo{reference franziskas poster in both figures}
+\todo{get a higher quality version of overview graphic}
+
+\begin{figure}
+\caption{SNIK Graph Overview}
+\label{fig:snik-graph-overview}
+\includegraphics[width=\columnwidth]{img/snik-graph-overview.png}
+\end{figure}
+
+\begin{figure}
+\caption{A teacher prepares a lecture on strategic HIS directing using the \enquote{Circle Star} function of SNIK Graph.}
+\label{fig:snik-graph-circlestar}
+\includegraphics[width=\columnwidth]{img/snik-graph-circle-star.png}
+\end{figure}
+
+\begin{figure}
+\caption{Students learn new concepts about HIS quality by linking them to concepts already learned.
+A teacher asks a student to find out how the new concept \enquote{Quality of Data} is linked to the \enquote{Patient Identification Number}.
+The student connects the two concepts by using the “spiderworm“ visualization and learns that a patient identification number is associated with object identity.
+Object identity is a subclass of integrity of data.
+Besides integrity of data, there are also 13 other criteria for quality of data.}
+\label{fig:snik-graph-spiderworm}
+\includegraphics[width=\columnwidth]{img/snik-graph-spiderworm.png}
+\end{figure}
+
+SNIK Graph~\citep{snikgraph} is a web application\footnote{\url{http://www.snik.eu/graph}} that transforms the classes and triples of SNIK to nodes and edges of a graph.
+The graph is visualized using the Cytoscape.js~\citep{cytoscape} library with the spring-based Euler layout\todo{cite?}.
+The u~\cref{fig:snik-graph-overview}
+
+Exploration options like the \enquote{spider worm} (see Figure 4), which consists of the shortest path between a start node and an end node together with the end node’s neighbourhood, illustrate the context of a concept.
+%Existing Linked Data visualizations, such as Relfinder, LodView and LodLive (see Table 1) are also adapted to SNIK using the SPARQL endpoint as a data source.
+
+\subsection{Quiz}
+The DBpedia-powered Clover Quiz~\citep{cloverquiz} shows that ontologies can be used to automatically generate multiple-choice questions.
+We apply this approach to SNIK and generate 1231 English questions using the templates described in \cref{tab:templates}.
+It was used by students from the Universities of Amsterdam, Heidelberg and Leipzig during the international Frank van Swieten lectures in 2019.
+To provide difficult wrong answers, the so called \emph{distractors}, we use direct neighbours of the classes that represent the correct answers. 
+Due to the limited nesting capabilties of SPARQL, which does not support loops but only subqueries that cannot access variables declared outside their scope, querying for distinct sets of 4 neighbours, where exactly one was connected over a certain property to a certain object, was not possible with our Virtuoso SPARQL endpoint initially.
+We circumvented the timeout by calculating the neighbour-relation as a first step and uploading it to the endpoint.
+
+%SNIK Quiz is freely available as an open source web application.
+Because SNIK only contains the knowledge that the source textbooks describe, it does not contain the complete domain of Hospital Information Management.
+As such, negative questions are problematic, as the given relationship could hold in the real world but not be described in the textbook source of the class. 
+The same problem concerns the distractors of positive questions.
+So this problem cannot be avoided.
+However, this is one of the reasons that we don't ask count questions like \enquote{How many functions is the CIO responsible for?}, which don't help much for learning anyways. % improve writing
+Relationships could also hold implicitly through the subclass hierarchy.
+For example, \aurl{bb}{ChiefInformationOfficer} is a subclass of \aurl{bb}{InformationManagementStaff}, which is responsible for \aurl{bb}{OperationalInformationManagement}.
+It is unclear, whether the CIO is also responsible for operational information management.
+In the future, we will add more question types
+
+\begin{table*}
+\caption{Templates}
+\label{tab:templates}
+\begin{tabulary}{\textwidth}{lL}
+\toprule
+\textbf{Template}	&\textbf{Description}\\
+\midrule
+Definition		&Ask for the class that fits the given textbook definition.\\
+Example			&What is defined as \enquote{Examination of in and out patients in radiological department}?\\
+Distractors		&Labels of other classes (that have a path of length 2 or less to the correct class.)\\
+\midrule
+Subject			&Ask for the class that is related via a given relation to a given object.\\
+Example			&Who is \emph{involved in} a \emph{healthcare network}?\\
+Distractors		&Labels of other classes (of the same type) that \emph{are not} related via the same relation to the same object.\\
+\bottomrule
+\end{tabulary}
+\end{table*}
 
@@
 SNIK is based on the Meta Model from Figure \todo{reference}.
-Thus, SNIK can be considered as an archetype for ontologies describing for a given domain (here “HIS management”), what functions a certain role has to carry out and what information a person of this role needs and what information she/he provides while carrying out the functions.
+Thus, SNIK can be considered as an archetype for ontologies describing for a given domain (here \enquote{HIS management}), what functions a certain role has to carry out and what information a person of this role needs and what information she/he provides while carrying out the functions.
 This may be used for domains like the medical department of a hospital as well.
@@
 
-Figure 4– A “Spider Worm” between the classes “Change Management” and “HIS Quality” in the SNIK Graph visualization web application.
+Figure 4– A \enquote{Spider Worm} between the classes \enquote{Change Management} and \enquote{HIS Quality} in the SNIK Graph visualization web application.
 The content of SNIK is taken from textbooks which are protected by property rights of publishing houses or taken from an interview with a CIO holding property rights as well.
@@
 In general, missing approvals may prevent the creation of such open ontologies and thus the provision of textbook content as open knowledge.
-The approvals for SNIK were possible because for each class and relation being part of SNIK there is a clear indication of its source (see code “bb” and chapter information in Table 1).
+The approvals for SNIK were possible because for each class and relation being part of SNIK there is a clear indication of its source (see code \enquote{bb} and chapter information in Table 1).
 In future work we will use these indications for establishing hyperlinks into the sources such that users of SNIK may use SNIK as a smart index for retrieving more detailed knowledge and supporting illustrations from the original textbooks – if digitally available.
@@
 \section{Acknowledgments}
-We thank the publishers ... and ... for granting us permission to publish the subontologies based on the textbooks \citet{bb,ob,he} under an open license.
-The SNIK project is supported by the DFG (German Research Foundation) under the grant numbers 1605/7-1 and 1387/8-1.
+This work is supported by the DFG (German Research Foundation) under the Projects SNIK, grant no. 1605/7-1 and 1387/8-1, as well as HITO, grant no. WI 1605/11-1 and I3726-N31.
+%We thank the publishers ... and ... for granting us permission to publish the subontologies based on the textbooks \citet{bb,ob,he} under an open license.
 %\nocite{*} 
@@
 
+\section{Notes to Reviewers}
+This paper includes and extends text and figures from:
+- SNIK Graph: \todo{include medinfo poster reference}
+
 \end{document}

### 2019-09-17 paper.tex
Add snik references to bibliography.
Extend application.

@@
 \section{Application}\label{sec:application}
-SNIK was initially intended~\citep{domaene} as the data source for software to support hospital CIOs.
-Two applications were designed and implemented as prototypes: (1) the requirements engineering decision support system TOREOnto~\citep{toreonto} and (2) the knowledge exploration and navigation visualization CIONo~\citep{ciono}.\todo{add ciono to bib}
-Due to data protection regulations, efforts to integrated them into the informational infrastructure of the Uniklinikum Leipzig were stopped.
-Subsequent efforts shifted to ontology assisted teaching.
+SNIK was initially intended~\citep{domaene} both as the data source for software to support hospital CIOs and to support teaching.
+For the former goal, two applications were designed and implemented as prototypes~\citep{toreonto}: (1) the requirements engineering decision support system TOREOnto and (2) the knowledge exploration and navigation visualization CIONx.
+Due to internal regulations, efforts to integrate them into the informational infrastructure of the Uniklinikum Leipzig were stopped.
+Consequently, latter efforts concentrate on the teaching support:
 
@@
 
-SNIK Graph~\citep{snikgraph} is a web application\footnote{\url{http://www.snik.eu/graph}} that transforms the classes and triples of SNIK to nodes and edges of a graph.
-The graph is visualized using the Cytoscape.js~\citep{cytoscape} library with the spring-based Euler layout\todo{cite?}.
+SNIK Graph is a web application\footnote{\url{http://www.snik.eu/graph}} that transforms the classes and triples of SNIK to nodes and edges of a graph.
+The graph is visualized using the Cytoscape.js~\citep{cytoscape} library with the force-directed Euler layout.
 The u~\cref{fig:snik-graph-overview}
@@
 \bibliographystyle{ios1}
-\bibliography{paper}
+\bibliography{paper,snik}
 

### 2019-09-17 paper.tex
More SNIK graph.

@@
 Due to internal regulations, efforts to integrate them into the informational infrastructure of the Uniklinikum Leipzig were stopped.
-Consequently, latter efforts concentrate on the teaching support:
+Consequently, the other approaches are in the area of teaching support:
 
@@
 \caption{A teacher prepares a lecture on strategic HIS directing using the \enquote{Circle Star} function of SNIK Graph.}
-\label{fig:snik-graph-circlestar}
+\label{fig:snik-graph-circle-star}
 \includegraphics[width=\columnwidth]{img/snik-graph-circle-star.png}
@@
 SNIK Graph is a web application\footnote{\url{http://www.snik.eu/graph}} that transforms the classes and triples of SNIK to nodes and edges of a graph.
-The graph is visualized using the Cytoscape.js~\citep{cytoscape} library with the force-directed Euler layout.
-The u~\cref{fig:snik-graph-overview}
-
-Exploration options like the \enquote{spider worm} (see Figure 4), which consists of the shortest path between a start node and an end node together with the end node’s neighbourhood, illustrate the context of a concept.
-%Existing Linked Data visualizations, such as Relfinder, LodView and LodLive (see Table 1) are also adapted to SNIK using the SPARQL endpoint as a data source.
+The graph is visualized using the Cytoscape.js~\citep{cytoscape} library with the force-directed Euler layout, see~\cref{fig:snik-graph-overview}.
+With several thousand classes, specific parts of SNIK can be hard to discern.
+Thus, there are several options to view subgraphs of SNIK, for example to show only a specific chapter of a book to prepare a lecture about a specific topic.
+Users can also search for and restrict the view to a single class and then show the neighbourhood of that class (see \cref{fig:snik-graph-circle-star}) and subsequently the neighbourhood of selected neighbours of the previous step.
+SNIK Graph can also calculate the shortest path to between classes.
+We also tried to automatically find the most interesting path but this was not successfull as it is subjective and the depends on the goal of the user.
+Path and neighourhood operations are joined in the \enquote{spider worm} (see \cref{fig:snik-graph-spiderworm}), which consists of the shortest path between a start node and an end node together with the end node’s neighbourhood, illustrate the context of a concept.
 

### 2019-09-17 paper.tex
Start lifecycle.

@@
 \section{Data Set Description}\label{sec:dsd}
+\begin{table*}
+\caption{use of established vocabularies}
+\label{tab:templates}
+\begin{tabulary}{\textwidth}{lL}
+\toprule
+\textbf{Vocabulary}	&\textbf{Description}\\
+\midrule
+&\\
+&\\
+&\\
+\bottomrule
+\end{tabulary}
+\end{table*}
+
 The SNIK ontologies are available at \url{http://www.snik.eu/ontology/} under the Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International license.
 
-\section{Application}\label{sec:application}
+\section{Lifecycle}\label{sec:application}
+After the transformation~\citep{sniktec} of the spreadsheets using Tarql, the resulting RDF was refined in Protege and then uploaded to a Virtuoso SPARQL endpoint. 
+Besides issues of collaborative editing, this required a regular export and reupload, which delayed the effect of changes in SNIK to the user facing applications.
+- git repository (link): collaborative editing but protege can cause large diffs -> people who understand rdf enough to use text editor ttl/rdfxml 
+- use ontowiki: slow but user friendly, undo function,
+- SPARUL queries
+
+\section{Application}
 SNIK was initially intended~\citep{domaene} both as the data source for software to support hospital CIOs and to support teaching.

### 2019-09-18 paper.tex
Add todo
Finish lifecycle prototype.

--- 2019-09-17 paper.tex
@@
 \section{Data Set Description}\label{sec:dsd}
+\begin{table*}
+\caption{use of established vocabularies}
+\label{tab:templates}
+\begin{tabulary}{\textwidth}{lL}
+\toprule
+\textbf{Vocabulary}	&\textbf{Description}\\
+\midrule
+&\\
+&\\
+&\\
+\bottomrule
+\end{tabulary}
+\end{table*}
+
 The SNIK ontologies are available at \url{http://www.snik.eu/ontology/} under the Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International license.
 
-\section{Application}\label{sec:application}
+\section{Lifecycle}\label{sec:application}
+
+\begin{figure}[ht]
+\caption{The SNIK lifecycle}
+\label{fig:lifecycle}
+\includegraphics[width=\columnwidth]{cycle.pdf}
+\end{figure}
+
+When we add a new textbook to SNIK, we identify all concepts that fit the meta model and their relations to other concepts.
+Some project members are not proficient in RDF serializations, so we first fill out a meta model conforming spreadsheet template.
+We then use Tarql\footnote{\url{https://github.com/cygri/tarql}} with a mapping configuration file\footnote{\url{https://github.com/IMISE/snik-csv2rdf}} to converts that spreadsheet to RDF.
+Next, we refine the data in Protégé and upload it to a Virtuso SPARQL endpoint.
+
+After the initial publish phase, we describe the repeating steps of the reuse phrase with the SNIK lifecycle in \Cref{fig:lifecycle}:
+\begin{itemize}
+\item the Virtuoso SPARQL endpoint is used for querying and as data source for all our applications  
+\item an instance of the OntoWiki is used for punctual changes and additions by non-Semantic-Web-experts, supplemented by SPARUL for changes involving large numbers of classes 
+\item LIMES~\citep{limes} generates interlink candidates between the different textbooks, which can be approved or rejected by the users
+\item LodView\footnote{\url{https://lodview.it/}} lets the user browse among the classes of snik and view particular classes in detail
+\item SNIK Graph, detailed in \Cref{sec:snik-graph}, supplements the RDF browser by visualizing the relationships between classes.
+Users often discover incorrect modelling or possibilities for enhancement in SNIK graph.
+\item SNIK quality is a web application that uses SPARQL queries to find problems of varying degree, such as violations of domain, range or naming conventions, subclass cycles or missing definitions.
+\item Problems suggested by users and SNIK quality are verified, solved, and published on the SPARQL endpoint using SPARUL queries, at which point the circle begins anew. 
+\end{itemize}
+
+%After the initial extraction step, SNIK is published and users suggest problems or enhancements.%Besides issues of collaborative editing, this required a regular export and reupload, which delayed the effect of changes in SNIK to the user facing applications.
+%- git repository (link): collaborative editing but protege can cause large diffs -> people who understand rdf enough to use text editor ttl/rdfxml 
+%- use ontowiki: slow but user friendly, undo function,
+%- SPARUL queries
+The technical environment of those services is described in~\citet{sniktec}.
+Selected applications are described in the following section.
+
+\section{Applications}
 SNIK was initially intended~\citep{domaene} both as the data source for software to support hospital CIOs and to support teaching.
@@
 
-\subsection{SNIK Graph}
+\subsection{SNIK Graph}\label{sec:snik-graph}
 \todo{reference franziskas poster in both figures}
 \todo{get a higher quality version of overview graphic}
+\todo{reference \citet{visualizationoflargeontologies}}
 
@@
 \label{fig:snik-graph-overview}
-\includegraphics[width=\columnwidth]{img/snik-graph-overview.png}
+\includegraphics[width=\columnwidth]{img/snik-graph.png}
 \end{figure}

### 2019-09-18 paper.tex
Improve abstract

@@
 Textbooks contain abstract knowledge about a domain.
-In the case of Hospital Information Management, much of this knowledge consists of systems, their parts and their relations, making it a natural fit for RDF, RDFS and OWL.
-As the domain is large and highly relevant to students of Medical Informatics, modelling the knowledge is not only possible but also very useful.
+In the case of Hospital Information Management, much of this knowledge consists of systems, their parts and their relations.%, making it a natural fit for RDF, RDFS and OWL.
+%As the domain is large and highly relevant to students of Medical Informatics, modelling the knowledge is not only possible but also very useful.
 We present a common data model for the domain and the result of applying the data model to three textbooks.
@@
 
-\section{Data Quality}
-- extractors not always same people who wrote book/part of book
-- mistake when extracting
-- reasoning does not work (test)
-\begin{itemize}
-\item Multiple Subtops
-\item Inconsistent Subtop with Subclass
-\item SKOS Link to Different Subtop
-\item Subclass Cycles
-\item Missing superclass
-\item Undefined Objects
-\item Domain Violation
-\item Range Violation
-\item Class URL Naming Convention Violations
-\item Property URL Naming Convention Violations
-\item Missing Definition
-\item Literals with Semicolons
-\item Classes with too many subclasses
-\item No Restriction
-\item Non-HTTP URI
-\item OWL 2 DL: Undefined Properties
-\item OWL 2 DL: Multiply Defined Properties
-\item OWL 2 DL: Class Typing Constraints
-\end{itemize}
--> use SHAPE/SHACL ?
+%\section{Data Quality}
+%- extractors not always same people who wrote book/part of book
+%- mistake when extracting
+%- reasoning does not work (test)
+%\begin{itemize}
+%\item Multiple Subtops
+%\item Inconsistent Subtop with Subclass
+%\item SKOS Link to Different Subtop
+%\item Subclass Cycles
+%\item Missing superclass
+%\item Undefined Objects
+%\item Domain Violation
+%\item Range Violation
+%\item Class URL Naming Convention Violations
+%\item Property URL Naming Convention Violations
+%\item Missing Definition
+%\item Literals with Semicolons
+%\item Classes with too many subclasses
+%\item No Restriction
+%\item Non-HTTP URI
+%\item OWL 2 DL: Undefined Properties
+%\item OWL 2 DL: Multiply Defined Properties
+%\item OWL 2 DL: Class Typing Constraints
+%\end{itemize}
+%-> use SHAPE/SHACL ?
 
@@
 \begin{center}
-\begin{tabular*}{0.96\columnwidth}{lll}
+\begin{tabular*}{\columnwidth}{ll}
 \toprule
-\textbf{Ontology}				&\textbf{Prefix}&\textbf{Source}\\
+\textbf{Ontology}				&\textbf{Source}\\
 \midrule
-\url{http://www.snik.eu/ontology/meta}		&meta		&All Sources\\
-\url{http://www.snik.eu/ontology/bb}		&bb		&Textbook~\cite{bb}\\
-\url{http://www.snik.eu/ontology/ob}		&ob		&Textbook~\cite{ob}\\
-\url{http://www.snik.eu/ontology/he}		&he		&Textbook~\cite{he}\\
-\url{http://www.snik.eu/ontology/ciox}		&ciox		&CIO Interview\\
-\url{http://www.snik.eu/ontology/it4it}		&it4it		&Standard~\cite{it4it}\\
+\url{http://www.snik.eu/ontology/meta}		&Meta Model\\
+\url{http://www.snik.eu/ontology/bb}		&Textbook~\cite{bb}\\
+\url{http://www.snik.eu/ontology/ob}		&Textbook~\cite{ob}\\
+\url{http://www.snik.eu/ontology/he}		&Textbook~\cite{he}\\
+\url{http://www.snik.eu/ontology/ciox}		&CIO Interview\\
+\url{http://www.snik.eu/ontology/it4it}		&Standard~\cite{it4it}\\
 \bottomrule
@@
 \citet{he} explains information management beyond the scope of healthcare. 
-CIOX is based on an interview with the CIO of the Universitätsklinikum Leipzig.
-IT4IT is based on the IT4IT standard.\todo{Sebastian: Was über Standards und IT4IT schreiben}
+Other sources are interviews and standards.
+%CIOX is based on an interview with the CIO of the Universitätsklinikum Leipzig.
+%IT4IT is based on the IT4IT standard.\todo{Sebastian: Was über Standards und IT4IT schreiben}
 
@@
 Some project members are not proficient in RDF serializations, so we first fill out a meta model conforming spreadsheet template.
-We then use Tarql\footnote{\url{https://github.com/cygri/tarql}} with a mapping configuration file\footnote{\url{https://github.com/IMISE/snik-csv2rdf}} to converts that spreadsheet to RDF.
+We then use Tarql\footnote{\url{https://github.com/cygri/tarql}} with a mapping configuration file\footnote{\url{https://github.com/IMISE/snik-csv2rdf}} to convert that spreadsheet to RDF.
 Next, we refine the data in Protégé and upload it to a Virtuso SPARQL endpoint.
@@
 After the initial publish phase, we describe the repeating steps of the reuse phrase with the SNIK lifecycle in \Cref{fig:lifecycle}:
-\begin{itemize}
+\begin{enumerate}
 \item the Virtuoso SPARQL endpoint is used for querying and as data source for all our applications  
@@
 \item LIMES~\citep{limes} generates interlink candidates between the different textbooks, which can be approved or rejected by the users
-\item LodView\footnote{\url{https://lodview.it/}} lets the user browse among the classes of snik and view particular classes in detail
+\item LodView\footnote{\url{https://lodview.it/}} lets the user browse among the classes of SNIK and view particular classes in detail
 \item SNIK Graph, detailed in \Cref{sec:snik-graph}, supplements the RDF browser by visualizing the relationships between classes.
@@
 \item Problems suggested by users and SNIK quality are verified, solved, and published on the SPARQL endpoint using SPARUL queries, at which point the circle begins anew. 
-\end{itemize}
+\end{enumerate}
 

### 2019-11-25 paper.tex
Implement or integrate FJs comments.

@@
 Textbooks contain abstract knowledge about a domain.
-In the case of Hospital Information Management, much of this knowledge consists of systems, their parts and their relations.%, making it a natural fit for RDF, RDFS and OWL.
+In the case of Hospital Information Management, much of this knowledge consists of information management functions, roles executing these functions  and the information used or updated by these functions.
+%, making it a natural fit for RDF, RDFS and OWL.
 %As the domain is large and highly relevant to students of Medical Informatics, modelling the knowledge is not only possible but also very useful.
-We present a common data model for the domain and the result of applying the data model to three textbooks.
+We present a common data model for the domain and the result of applying the data model to three textbooks an interview and a standard.
 We publish the result over several interfaces that are useful for researchers, practitioners and students, depending on their objectives and their Semantic Web skills.
@@
 Given a care delivery organization, its health information system (HIS) can be defined as the part of the organization that processes and stores data, information, and knowledge.
-It usually consists of a large number of different application systems, computers, and network components~\citep{bb}.
+It usually consists of different application systems, computers, and network components~\citep{bb}.
 Managing an HIS comprises planning, monitoring and directing activities.
@@
 In order to integrate different knowledge sources and to provide the knowledge in a structured, machine-readable data format, we extracted knowledge about HIS management from three textbooks~\citep{bb,ob,he} and other sources.
-The combination of this knowledge~\citep{einsemantischesnetz,domaene,approachtosupport} results in SNIK, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German).
+The combination of this knowledge~\citep{einsemantischesnetz,domaene,approachtosupport} results in the SNIK ontology, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German).
 
 In order to encourage and enable other researchers, students and health informatics professionals to use available knowledge of HIS management, we answer the following questions in this paper:
-
-We discuss advantages and disadvantages of the interfaces and give recommendations on which ones present the best compromise for different use cases and target audiences.
+\todo{add questions}
+We introduce different user interfaces to the SNIK ontology and discuss their suitability for different use cases of the target audiences.
+%We discuss advantages and disadvantages of the interfaces and give recommendations on which ones present the best compromise for different use cases and target audiences.
 We conclude with plans for future work on interlinking and visualization.
 
-As the domain is large and highly relevant to students of Medical Informatics, modelling the knowledge is not only possible but also very useful.
+%As the domain is large and highly relevant to students of Medical Informatics, modelling the knowledge is not only possible but also very useful.
+Medical informatics students, who are trained for executive positions in information management departments of healthcare institutions, such as hospitals, need a clear terminology of their domain.
+Due to different frameworks and textbooks dealing with information management in healthcare, modelling the knowledge unravels the links between the different views on information management.
+These are only implicitly known or not known at all by experts in the field.
 
@@
 Using a data model of Information Management 
-Depending on the learning style and information need of the student, several tools are available to cosnume that knowledge.
+Depending on the learning style and information need of the student, several tools are available to consume that knowledge.
+\todo{FJ: Welchen Zweck hat dieser Teil in der Einleitung? Auch würde ich die Ziele vielleicht etwas kompakter als Liste beschreiben.}
 
@@
 
-The Meta Model (see \cref{fig:metamodel}) provides a common vocabulary for the domain of HIS management and thus defines, which superclasses and relations can be used.
-SNIK version 0.4.1 comprises five subontologies that are built upon the Meta Model, see Table 1.
+The meta model (see \cref{fig:metamodel}) provides a common vocabulary for the domain of HIS management and thus defines, which superclasses and property can be used.
+SNIK version 0.4.1 comprises five subontologies that are built upon the meta model, see Table 1.
 At the head of the class hierarchy is the \enquote{Top} class, which has exactly three disjunctive subclasses.
-Following the Meta Model, each class has to be a subclass of exactly one of them.
+Following the meta model, each class has to be a subclass of exactly one of them.
 The correct superclass of a new concept can be found by answering the question: Who (\enquote{Role}) does what (\enquote{Function}) and which information (\enquote{EntityType}) is needed?
-If a concept is neither of them, it cannot be modeled using the Meta Model.
+If a concept is neither of them, it cannot be modeled using the meta model.
 As the subclass relation is transitive, a new class can be placed further down the hierarchy and it can still be inferred, whether it is a Function, Role or EntityType (see Figure 3).
-Besides the subclass relationship, two classes can be connected with relations provided by the Meta Model.
-The generic \enquote{is associated with} relationship is possible between all classes but carries little information.
-Which more specific relations between two classes are available depends on their superclasses.
+Besides the subclass relationship, two classes can be connected with relations provided by the meta model.
+The generic \enquote{is associated with} relationship carries little information.
 For example, a role and a function can be connected as \enquote{is involved in} \enquote{is responsible for} and \enquote{approves}.
@@
 \section{Concessions to Practicality}
-Our main focus is that SNIK provides a net benefit to its users, many of which do not have Semantic Web expertise.
+SNIK is mainly used by students and researchers having no or little semantic web knowledge.
 This led to several concessions to usability and actual usage by sacrificing full conformance to the W3C OWL recommendation in the following cases:
@@
 The definitions of \aurl{rdfs}{domain} and \aurl{rdfs}{range} state that the subjects, respectively objects of all triples with that property are instances of that class.
-Because SNIK contains classes, which are subclasses, not instances of the subtopclasses, using those properties directly in SNIK is technically wrong and OWL property restrictions should be used.
+The classes of SNIK are subclasses, not instances, of the subtopclasses.
+This means that using those properties directly in SNIK is technically wrong and OWL property restrictions should be used.
 This presented the following problems:
@@
 It may even be the case that both of those statements are not what the author intended.
-\item Deleting or adding a connection between classes in the OntoWiki is no longer an atomic operation and may result in unreferenced blank nodes or partially restored undo results.
+\todo{FJ: Sollte man hier evtl.
+noch erklären, wie es denn tatsächlich gemeint ist?
+Eigentlich haben die in SNIK modellierten Aussagen jeweils eine Bedeutung wie: Laut Lehrbuch xy muss/sollte ein CEO im Krankenhaus i.
+d.
+R.
+an Projektreviews (innerhalb von IM-Projekten) beteiligt sein.
+Ob ein Krankenhaus einen oder oder mehrere CEOs hat, ist uns an der Stelle egal, ebenso wie viele Projektreviews das sind.
+Wir drücken hier sozusagen eher eine Regel (laut Lehrbuch) aus.
+}
+\item Deleting or adding a connection between classes in the OntoWiki\todo{needs to be introduced first} is no longer an atomic operation and may result in unreferenced blank nodes or partially restored undo results.
 \end{itemize}
@@
 \begin{figure*}
-\caption{The SNIK Meta Model}
+\caption{The SNIK meta model}
 \label{fig:metamodel}
@@
 The \textbf{Semantic Network of Information Management in Hospitals} (SNIK\footnote{Hospital means \enquote{Krankenhaus} in German.}) is a modular OWL 2 DL ontology.
-The \enquote{Meta Model} defines three basic disjunctive classes and their possible relations: Roles (who), Function (does what) and Entity Types (and which information is therefore needed).
+The \enquote{meta model} defines three basic disjunctive classes and their possible relations: Roles (who), Function (does what) and Entity Types (and which information is therefore needed).
 A set of modular subontologies define subclasses of those three classes and their relations as described by a certain knowledge source about information management in hospitals:
@@
 \item the Virtuoso SPARQL endpoint is used for querying and as data source for all our applications  
-\item an instance of the OntoWiki is used for punctual changes and additions by non-Semantic-Web-experts, supplemented by SPARUL for changes involving large numbers of classes 
+\item an instance of the OntoWiki~\citep{ontowiki} is used for punctual changes and additions by non-Semantic-Web-experts, supplemented by SPARUL for changes involving large numbers of classes 
 \item LIMES~\citep{limes} generates interlink candidates between the different textbooks, which can be approved or rejected by the users
@@
 \section{Applications}
-SNIK was initially intended~\citep{domaene} both as the data source for software to support hospital CIOs and to support teaching.
+SNIK was initially intended~\citep{domaene} both as  software to support hospital CIOs and to support teaching.
 For the former goal, two applications were designed and implemented as prototypes~\citep{toreonto}: (1) the requirements engineering decision support system TOREOnto and (2) the knowledge exploration and navigation visualization CIONx.
@@
 \end{figure}
-
+\todo{FJ: In diesem Abschnitt fehlt die Diskussion, wie es durch die verschiedenen Zielgruppen bei Use Cases genutzt werden kann.}
 SNIK Graph is a web application\footnote{\url{http://www.snik.eu/graph}} that transforms the classes and triples of SNIK to nodes and edges of a graph.
@@
 \section{Discussion}
-SNIK is based on the Meta Model from Figure \todo{reference}.
+SNIK is based on the meta model from Figure \todo{reference}.
 Thus, SNIK can be considered as an archetype for ontologies describing for a given domain (here \enquote{HIS management}), what functions a certain role has to carry out and what information a person of this role needs and what information she/he provides while carrying out the functions.
-This may be used for domains like the medical department of a hospital as well.
+This may be used for domains like the medical department of a hospital as well.\todo{FJ: Ich glaube, das müsste man ausführlicher erklären, das würde ich eher ans Ende der Diskussion nehmen.}
 This would be valuable knowledge, for instance as basis for systems analysis projects [1].
@@
 \section{Future Work}
+\todo{Connection with HITO}
 We plan to extend the text-based interlinks with semantic-similarity-based ontology matching [5] to find more correspondencies between the textbooks.
@@
 %\nocite{*} 
+\todo{FJ: Hmmm, ziemlich selbstreferenziell. Brauchen wir da wirklich so viele Publikationen von uns? Welche anderen könnte man noch einbringen? Ich schaue auch mal, was mir so über den Weg läuft.}
 \bibliographystyle{ios1}
 \bibliography{paper,snik}
-
 \section{Notes to Reviewers}

### 2019-12-06 paper.tex
Add missing images and reference to FJ's medinfo paper.

@@
 
-In order to encourage and enable other researchers, students and health informatics professionals to use available knowledge of HIS management, we answer the following questions in this paper:
-\todo{add questions}
-We introduce different user interfaces to the SNIK ontology and discuss their suitability for different use cases of the target audiences.
+In order to encourage and enable other researchers, students and health informatics professionals to use available knowledge of HIS management, We introduce different user interfaces to the SNIK ontology and discuss their suitability for different use cases of the target audiences.
 %We discuss advantages and disadvantages of the interfaces and give recommendations on which ones present the best compromise for different use cases and target audiences.
@@
 \begin{figure}
-\caption{SNIK Graph Overview}
+\caption{SNIK Graph Overview. Source: \protect\citet{snikgraph}.}
 \label{fig:snik-graph-overview}
@@
 \begin{figure}
-\caption{A teacher prepares a lecture on strategic HIS directing using the \enquote{Circle Star} function of SNIK Graph.}
+\caption{A teacher prepares a lecture on strategic HIS directing using the \enquote{Circle Star} function of SNIK Graph. Source: \protect\citet{snikgraph}.}
 \label{fig:snik-graph-circle-star}
@@
 This paper includes and extends text and figures from:
-- SNIK Graph: \todo{include medinfo poster reference}
+\citet{snikgraph}
 

### 2020-01-10 paper.tex
Update template to version from 2019/05/06

@@
 \citet{domaene} contains an initial structure of the meta model.
-\todo{Describe the meta model here}
-The \textbf{Semantic Network of Information Management in Hospitals} (SNIK\footnote{Hospital means \enquote{Krankenhaus} in German.}) is a modular OWL 2 DL ontology.
 The \enquote{meta model} defines three basic disjunctive classes and their possible relations: Roles (who), Function (does what) and Entity Types (and which information is therefore needed).
 A set of modular subontologies define subclasses of those three classes and their relations as described by a certain knowledge source about information management in hospitals:
+The \textbf{Semantic Network of Information Management in Hospitals} (SNIK\footnote{Hospital means \enquote{Krankenhaus} in German.}) is a modular OWL 2 DL ontology.
 %from different sources:% three textbooks, an interview and a standard.

### 2020-01-10 paper.tex
Shorten concessions to practicality.

@@
 \daurl{bb}{http://www.snik.eu/ontology/bb/}
+\daurl{meta}{http://www.snik.eu/ontology/meta/}
 \lstset{language=SPARQL,breaklines=true}
@@
 
-\section{Concessions to Practicality}
-SNIK is mainly used by students and researchers having no or little semantic web knowledge.
-This led to several concessions to usability and actual usage by sacrificing full conformance to the W3C OWL recommendation in the following cases:
-
-\paragraph{OWL Restrictions}
-The meta model offers several properties between the different subtopclasses.
-The definitions of \aurl{rdfs}{domain} and \aurl{rdfs}{range} state that the subjects, respectively objects of all triples with that property are instances of that class.
-The classes of SNIK are subclasses, not instances, of the subtopclasses.
-This means that using those properties directly in SNIK is technically wrong and OWL property restrictions should be used.
-This presented the following problems:
-\begin{itemize}
-\item Tools such as LodLive don't understand those OWL constructs and only show the RDF form with blank nodes and helper constructs, which is confusing to the users
-\item Communicating OWL restrictions to extractors and non-Semantic Web expert project members is difficult.
-\item Extractors only enter the name of the property, not whether it is a "someValuesFrom" or "allValuesFrom" restriction.
-For example, it is unclear whether each CEO is involved in at least one project review, or whether each CEO is only involved in project reviews.
-It may even be the case that both of those statements are not what the author intended.
-\todo{FJ: Sollte man hier evtl.
-noch erklären, wie es denn tatsächlich gemeint ist?
-Eigentlich haben die in SNIK modellierten Aussagen jeweils eine Bedeutung wie: Laut Lehrbuch xy muss/sollte ein CEO im Krankenhaus i.
-d.
-R.
-an Projektreviews (innerhalb von IM-Projekten) beteiligt sein.
-Ob ein Krankenhaus einen oder oder mehrere CEOs hat, ist uns an der Stelle egal, ebenso wie viele Projektreviews das sind.
-Wir drücken hier sozusagen eher eine Regel (laut Lehrbuch) aus.
-}
-\item Deleting or adding a connection between classes in the OntoWiki\todo{needs to be introduced first} is no longer an atomic operation and may result in unreferenced blank nodes or partially restored undo results.
-\end{itemize}
-
-To avoid those problems, we use direct triples between the classes of SNIK as if they were instances, not classes.
+As a concession to practicality, we express each geleral rule extracted from a textbook as a single triple using classes as subject and object, and a property of the meta model.
+For example, the rule "the CEO is involved in project reviews" is modelled as \enquote{:Ceo \aurl{meta}{isInvolvedIn} :ProjectReview.}, where \texttt{:Ceo} is a subclass of \aurl{meta}{Role} and \aurl{meta}{ProjectReview} is a subclass of \aurl{meta}{Function}.
+While an OWL restrictions using \enquote{someValuesFrom} and \enquote{allValuesFrom} would be technically correct, the knowledge is not expressed that specifically in the textbooks.
+Additionally, encoding each fact as a single triple facilitates tool support and prevents inconsistencies through non-atomic changes.
 

### 2020-06-11 paper.tex
Extend paper.

@@
 %\pretitle{}
-\title{Textbook Knowledge on Hospital Information Management as Linked Data}
+\title{SNIK: Textbook Knowledge on Hospital Information Management as Linked Data}
 %\runningtitle{The SNIK Ontology of Hospital Information Management}
@@
 \begin{abstract}
-Textbooks contain abstract knowledge about a domain.
-In the case of Hospital Information Management, much of this knowledge consists of information management functions, roles executing these functions  and the information used or updated by these functions.
+%Textbooks contain abstract knowledge about a domain.
+Textbooks about information management in hospitals describe the planning, monitoring and directing of a hospital's information system.
+SNIK consists of the manually transformed content of three such textbooks as Linked Open Data.
+The data model describes information management functions, roles executing these functions and the information used or updated by these functions.
+%SNIK provides applications that are used to teach students internationally.
+%and the result of applying the data model to three textbooks, an interview and a standard.
+%To compare, 
 %, making it a natural fit for RDF, RDFS and OWL.
 %As the domain is large and highly relevant to students of Medical Informatics, modelling the knowledge is not only possible but also very useful.
-We present a common data model for the domain and the result of applying the data model to three textbooks an interview and a standard.
-We publish the result over several interfaces that are useful for researchers, practitioners and students, depending on their objectives and their Semantic Web skills.
+We publish the result over several different interfaces that are useful for researchers, information system administrators or students, depending on their objectives and their capabilities.
 \end{abstract}
@@
 \section{Introduction}
-Given a care delivery organization, its health information system (HIS) can be defined as the part of the organization that processes and stores data, information, and knowledge.
-It usually consists of different application systems, computers, and network components~\citep{bb}.
-Managing an HIS comprises planning, monitoring and directing activities.
-Due to the complexity and the unique conditions in health care, HIS management is an exceptionally challenging task.
-There is an enormous amount of frameworks, textbooks and articles describing the scope of HIS management from the perspective of medical informatics.
+A health information system (HIS) is the socio-technical subsystem of a care delivery organization~\citep{bb}.
+This includes its different application systems, computers, and network components as well as their users.
+A HIS processes data, information, and knowledge and its management involves planning, monitoring and directing of those activities.
+Due to the complexity and the unique conditions in health care, HIS management is an challenging task.
+There are many different frameworks, textbooks and articles describing the scope of HIS management from the perspective of medical informatics.
 %However, the disciplines of business informatics and information systems (IS) provide an even broader view on information systems and their management.
-A structured representation of the different perspectives leads to a holistic view on HIS management and helps help researchers and students connect their existing knowledge with further knowledge from other sources during research and learning.
+%A structured representation of the different perspectives leads to a holistic view on HIS management and helps help researchers and students connect their existing knowledge with further knowledge from other sources during research and learning.
 In order to integrate different knowledge sources and to provide the knowledge in a structured, machine-readable data format, we extracted knowledge about HIS management from three textbooks~\citep{bb,ob,he} and other sources.
-The combination of this knowledge~\citep{einsemantischesnetz,domaene,approachtosupport} results in the SNIK ontology, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German).
+The combination of this knowledge~\citep{einsemantischesnetz,domaene,approachtosupport} results in SNIK, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German).
 
@@
 %We discuss advantages and disadvantages of the interfaces and give recommendations on which ones present the best compromise for different use cases and target audiences.
-We conclude with plans for future work on interlinking and visualization.
+%We conclude with plans for future work on interlinking and visualization.
 
@@
 Publishing textbook knowledge as Linked Data enables different ways of teaching.
-The data is the result of 
-The extraction effort includes 
 The data is continously revised: users constantly report wrong or missing data in the visualization, which is then corrected, respectively revised, by the researchers of the project.
 
-This paper describes the current state of digitalized Hospital Information Management textbook knowledge.
-
-Using a data model of Information Management 
-Depending on the learning style and information need of the student, several tools are available to consume that knowledge.
 \todo{FJ: Welchen Zweck hat dieser Teil in der Einleitung? Auch würde ich die Ziele vielleicht etwas kompakter als Liste beschreiben.}
@@
 
+\section{Data Quality}
+
 \section{Applications}

### 2020-07-14 paper.tex
Update paper.

@@
 \usepackage{cleveref}
+\newcommand{\snikversion}{1.0}
+\newcommand{\sniktriples}{\num{71852}}
+\newcommand{\snikclasses}{\num{2655}}
+\newcommand{\snikproperties}{\num{339}}
+\newcommand{\sniklinks}{\num{581}}
 %%%%%%%%%%% End of definitions
 
-\pubyear{2019}
+\pubyear{2020}
 \volume{0}
@@
 \author[A]{\fnms{Alfred} \snm{Winter}\ead[label=e6]{alfred.winter@imise.uni-leipzig.de}}
-\runningauthor{K. Höffner et al.}
+%\runauthor{K. Höffner et al.}
 \address[A]{Institute for Medical Informatics, Statistics and Epidemiology (IMISE),
@@
 In order to integrate different knowledge sources and to provide the knowledge in a structured, machine-readable data format, we extracted knowledge about HIS management from three textbooks~\citep{bb,ob,he} and other sources.
-The combination of this knowledge~\citep{einsemantischesnetz,domaene,approachtosupport} results in SNIK, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German).
+The combination of this knowledge~\citep{semantischesnetz,domaene,approachtosupport} results in SNIK, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German).
 
@@
 Publishing textbook knowledge as Linked Data enables different ways of teaching.
-The data is continously revised: users constantly report wrong or missing data in the visualization, which is then corrected, respectively revised, by the researchers of the project.
-
-\todo{FJ: Welchen Zweck hat dieser Teil in der Einleitung? Auch würde ich die Ziele vielleicht etwas kompakter als Liste beschreiben.}
+%The data is continously revised: users constantly report wrong or missing data in the visualization, which is then corrected, respectively revised, by the researchers of the project.
 
@@
 We thus call our data model the \enquote{meta model}, as it is an ontology.
-
+meta model in accordance with its definition in the domain as a shared modelling language~\citep[p.~8]{ob}.
 The meta model (see \cref{fig:metamodel}) provides a common vocabulary for the domain of HIS management and thus defines, which superclasses and property can be used.
-SNIK version 0.4.1 comprises five subontologies that are built upon the meta model, see Table 1.
+SNIK version \snikversion{} comprises five subontologies that are built upon the meta model, see Table 1.
 At the head of the class hierarchy is the \enquote{Top} class, which has exactly three disjunctive subclasses.
@@
 As a concession to practicality, we express each geleral rule extracted from a textbook as a single triple using classes as subject and object, and a property of the meta model.
-For example, the rule "the CEO is involved in project reviews" is modelled as \enquote{:Ceo \aurl{meta}{isInvolvedIn} :ProjectReview.}, where \texttt{:Ceo} is a subclass of \aurl{meta}{Role} and \aurl{meta}{ProjectReview} is a subclass of \aurl{meta}{Function}.
-While an OWL restrictions using \enquote{someValuesFrom} and \enquote{allValuesFrom} would be technically correct, the knowledge is not expressed that specifically in the textbooks.
+For example, the rule "the CEO is involved in project reviews" is modelled as \enquote{\texttt{:Ceo} \aurl{meta}{isInvolvedIn} \texttt{:ProjectReview}.}, where \texttt{:Ceo} is a subclass of \aurl{meta}{Role} and \aurl{meta}{ProjectReview} is a subclass of \aurl{meta}{Function}.
+While an OWL restrictions using \aurl{owl}{someValuesFrom} and \aurl{owl}{allValuesFrom} would be technically correct, the knowledge is not expressed that specifically in the textbooks.
 Additionally, encoding each fact as a single triple facilitates tool support and prevents inconsistencies through non-atomic changes.
@@
 URL		&\url{http://www.snik.eu/ontology}\\
-Version		&2018-10-10, 0.4.1\\
+Version		&2018-10-10, \snikversion{}\\
 License		&CC BY-NC-SA 4.0\\
@@
 RDF Browser	&\url{http://www.snik.eu/ontology}\\
-Download	&\url{https://github.com/IMISE/snik-ontology/releases/download/0.8.0/snik-0.8.zip}\\
-\# Triples	&\num{112747}\\
-\# Classes	&\num{4729}\\
-\# Properties	&\num{329}\\
-\# Interlinks	&\num{713}\\
+Download	&\url{https://github.com/IMISE/snik-ontology/releases/download/1.0.0/snik.zip}\\
+\# Triples	&\sniktriples{}\\
+\# Classes	&\snikclasses{}\\
+\# Properties	&\snikproperties{}\\
+\# Interlinks	&\sniklinks{}\\
 %Subontologies	&\url{http://www.snik.eu/ontology/bb}\\
@@
 \section{Data Model}\label{sec:architecture}
-\citet{domaene} contains an initial structure of the meta model.
+The initial structure of the meta model is defined in \citet{domaene}.
 The \enquote{meta model} defines three basic disjunctive classes and their possible relations: Roles (who), Function (does what) and Entity Types (and which information is therefore needed).

### 2020-07-15 paper.tex
Update triple count including DBpedia links.
Update statistics and wrap its tabulary in a table.

@@
 \newcommand{\snikversion}{1.0}
-\newcommand{\sniktriples}{\num{71852}}
+\newcommand{\sniktriples}{\num{72046}}
 \newcommand{\snikclasses}{\num{2655}}
@@
 
+\begin{table}
+\caption{Statistics and Details}
+\label{tab:namespaces}
 \begin{tabulary}{\columnwidth}{lL}
@@
 URL		&\url{http://www.snik.eu/ontology}\\
-Version		&2018-10-10, \snikversion{}\\
+Version		&\snikversion{}\\
 License		&CC BY-NC-SA 4.0\\
 SPARQL Endpoint	&\url{http://www.snik.eu/sparql}\\
+LOD Cloud	&\url{https://lod-cloud.net/dataset/snik}\\
+Datahub		&\url{https://datahub.io/imise/snik/v/1}\\
 Visualization	&\url{http://www.snik.eu/graph}\\
@@
 \# Properties	&\snikproperties{}\\
-\# Interlinks	&\sniklinks{}\\
+\# Internal Links	&\num{360}\\
+\# Links to HITO  	&\num{260}\\
+\# Links to DBpedia 	&\sniklinks{}\\
 %Subontologies	&\url{http://www.snik.eu/ontology/bb}\\
@@
 \end{tabulary}%The source code for the services is available at \url{https://github.com/imise}
+\end{table}
 

### 2020-08-21 paper.tex
Expand interlinking.
Improve and split details and statistics table.
Add performance statements to SNIK Graph.

@@
 \newcommand{\snikversion}{1.0}
-\newcommand{\sniktriples}{\num{72046}}
-\newcommand{\snikclasses}{\num{2655}}
-\newcommand{\snikproperties}{\num{339}}
-\newcommand{\sniklinks}{\num{581}}
+\newcommand{\sniktriples}{72046}
+\newcommand{\snikclasses}{2655}
+\newcommand{\snikproperties}{339}
+\newcommand{\sniklinks}{581}
 %%%%%%%%%%% End of definitions
@@
 \begin{table}
-\caption{Statistics and Details}
-\label{tab:namespaces}
+\caption{Access Options and Details}
+\label{tab:access}
 \begin{tabulary}{\columnwidth}{lL}
@@
 Download	&\url{https://github.com/IMISE/snik-ontology/releases/download/1.0.0/snik.zip}\\
-\# Triples	&\sniktriples{}\\
-\# Classes	&\snikclasses{}\\
-\# Properties	&\snikproperties{}\\
-\# Internal Links	&\num{360}\\
-\# Links to HITO  	&\num{260}\\
-\# Links to DBpedia 	&\sniklinks{}\\
+\bottomrule
+\end{tabulary}%The source code for the services is available at \url{https://github.com/imise}
+\end{table}
+
+\begin{table}
+\caption{Statistics}
+\label{tab:statistics}
+\begin{tabulary}{\columnwidth}{lS}
+\toprule
+Nr. of Triples			&\sniktriples{}\\
+Nr. of Classes			&\snikclasses{}\\
+Nr. of Properties		&\snikproperties{}\\
+Nr. of Internal Links		&360\\
+Nr. of Links to HITO  		&260\\
+Nr. of Links to DBpedia 	&\sniklinks{}\\
 %Subontologies	&\url{http://www.snik.eu/ontology/bb}\\
@@
 
+\section{Data Transformation Process}
+\subsection{Interlinking}
+%As shown in \cref{...}, we aligned our dataset to several others:
+We manually aligned our dataset to HITO, the Health IT Ontology~\citep{hito}
+We automatically aligned our dataset to:
+- DBpedia~\citep{dbpedia}, as  
+- 
 \section{Lifecycle}\label{sec:application}
@@
 \subsection{SNIK Graph}\label{sec:snik-graph}
+\paragraperformance
+due to the limitations of JavaScript (e.g.
+single threaded), which are hard to work around, which leads to performance problems with the library Cytoscape.js which is the base of the graph visualization, on very large ontologies / graphs.
+right now we have 2641 nodes and 4904 edges and the performance is ok in Chrome but not amazing.
+with 10000 nodes it would probably be too slow on most machines.
+
 \todo{reference franziskas poster in both figures}

### 2021-08-20 paper.tex
Update paper.

@@
 \section{Future Work}
-\todo{Connection with HITO}
 We plan to extend the text-based interlinks with semantic-similarity-based ontology matching [5] to find more correspondencies between the textbooks.
@@
 We are also investigating candidates for additional subontologies and as interlink targets.
+
+Past efforts of automatic textbook extractions were unsuccessfull but recent advances in natural language processing, such as GPT-3~\citep{gpt3}, motivate further investigation.
 

### 2021-11-21 paper.tex
Comment out undefined command to get it to compile again.
Integrate content from SNIK Graph paper.

@@
 % add. options: [seceqn,secthm,crcready,onecolumn]
-\documentclass[sw]{iosart2x}
-
+%\documentclass[sw]{iosart2x}
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+\usepackage[utf8]{inputenc}
+\usepackage[english]{babel}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 %\usepackage{dcolumn}
@@
 \usepackage{tabulary}
+\usepackage{tabularx}
 \usepackage{csquotes}
 \usepackage{siunitx}
-%\usepackage{graphicx}
 \usepackage{todonotes}
 %\usepackage{natbib} % does not seem to work with ios 1
-\renewcommand{\citet}{\cite}% citet is not defined without natbib
-\renewcommand{\citep}{\cite}% citep is not defined without natbib
+\newcommand{\citet}{\cite}% citet is not defined without natbib
+\newcommand{\citep}{\cite}% citep is not defined without natbib
 \setlength{\marginparwidth}{2.1cm}% enough space for todonotes
@@
 \usepackage{aurl}
+\daurl{ob}{http://www.snik.eu/ontology/bb/}
 \daurl{bb}{http://www.snik.eu/ontology/bb/}
@@
 \lstset{language=SPARQL,breaklines=true}
-\usepackage{cleveref}
-\newcommand{\snikversion}{1.0}
+\usepackage[noabbrev,capitalize]{cleveref}
+\newcommand{\snikversion}{1.3.0}
 \newcommand{\sniktriples}{72046}
@@
 
-\pubyear{2020}
-\volume{0}
-\firstpage{1}
-\lastpage{1}
+%\pubyear{2020}
+%\volume{0}
+%\firstpage{1}
+%\lastpage{1}
 
@@
 
-\begin{frontmatter}
+%\begin{frontmatter}
 
@@
 %\subtitle{}
-
+\iffalse
 % Two or more authors:
@@
 %Härtelstraße 16--18, D-04107 Leipzig
+\fi
 
+\author{\IEEEauthorblockN{Konrad Höffner}
+\IEEEauthorblockA{\textit{Institute for Medical Informatics, Statistics and Epidemiology} \\
+\textit{Leipzig University}\\
+Leipzig, Germany \\
+\url{https://orcid.org/0000-0001-7358-3217}}
+\and
+\IEEEauthorblockN{Franziska Jahn}
+\IEEEauthorblockA{\textit{Institute for Medical Informatics, Statistics and Epidemiology} \\
+\textit{Leipzig University}\\
+Leipzig, Germany \\
+\url{https://orcid.org/0000-0002-7687-8544}}
+\and
+\and
+\IEEEauthorblockN{Alfred Winter}
+\IEEEauthorblockA{\textit{Institute for Medical Informatics, Statistics and Epidemiology} \\
+\textit{Leipzig University}\\
+Leipzig, Germany \\
+%alfred.winter@imise.uni-leipzig.de}
+\url{https://orcid.org/0000-0003-0179-954X}}
+\and
+\IEEEauthorblockN{Thomas Pause}
+\IEEEauthorblockA{\textit{Institute for Medical Informatics, Statistics and Epidemiology} \\
+\textit{Leipzig University}\\
+Leipzig, Germany \\
+\url{https://orcid.org/0000-0001-5832-4890}}
+}
+\maketitle
 \begin{abstract}
@@
 
-\begin{keyword}
-\kwd{information management, information systems, hospital information management}
-\end{keyword}
+% SWJ style
+%\begin{keyword}
+%\kwd{information management, information systems, hospital information management}
+%\end{keyword}
 
-\end{frontmatter}
-
-\begin{table}
-\caption{}
-\label{tab:namespaces}
-\begin{tabular}{ll}
-\toprule
-\textbf{x}	&\textbf{y}\\
-\midrule
-\bottomrule
-\end{tabular}
-\end{table}
+% ICSCR style
+\begin{IEEEkeywords}
+linked open data visualization,health information systems,information management
+\end{IEEEkeywords}
+%\end{frontmatter}
 
@@
 %A structured representation of the different perspectives leads to a holistic view on HIS management and helps help researchers and students connect their existing knowledge with further knowledge from other sources during research and learning.
-In order to integrate different knowledge sources and to provide the knowledge in a structured, machine-readable data format, we extracted knowledge about HIS management from three textbooks~\citep{bb,ob,he} and other sources.
+In order to integrate different knowledge sources and to provide the knowledge in a structured, machine-readable data format, we extracted knowledge about HIS management from three textbooks~\citep{bb,ob,he} and other sources, see \cref{tab:source}.
 The combination of this knowledge~\citep{semantischesnetz,domaene,approachtosupport} results in SNIK, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German).
 
-In order to encourage and enable other researchers, students and health informatics professionals to use available knowledge of HIS management, We introduce different user interfaces to the SNIK ontology and discuss their suitability for different use cases of the target audiences.
+In order to encourage and enable other researchers, students and health informatics professionals to use available knowledge of HIS management, we introduce different user interfaces to the SNIK ontology and discuss their suitability for different use cases of the target audiences.
 %We discuss advantages and disadvantages of the interfaces and give recommendations on which ones present the best compromise for different use cases and target audiences.
@@
 
+\begin{table}[t!]
+\caption{The different subontologies of SNIK and their sources.
+The URL of each subontology is equal to its graph name, the namespace has an additional slash \enquote{/} suffix.
+%All graphs are combined in the graph group \lowercase{\url{http://www.snik.eu/ontology}}.
+}
+\label{tab:source}
+\begin{center}
+%\begin{tabular}{\columnwidth}{lll}
+\begin{tabular}{lll}
+\toprule
+\textbf{Ontology}				&\textbf{Source}					&\textbf{Prefix}\\
+\midrule
+\url{http://www.snik.eu/ontology/meta}		&Meta Model				&meta\\
+\url{http://www.snik.eu/ontology/bb}		&Textbook~\cite{bb}		&bb\\
+\url{http://www.snik.eu/ontology/ob}		&Textbook~\cite{ob}		&ob\\
+\url{http://www.snik.eu/ontology/he}		&Textbook~\cite{he}		&he\\
+\url{http://www.snik.eu/ontology/ciox}		&CIO Interview			&ciox\\
+\url{http://www.snik.eu/ontology/it4it}		&Standard~\cite{it4it}	&it4it\\
+\url{http://www.snik.eu/ontology}			&Combined Graph Group	&sniko\\
+\bottomrule
+\end{tabular}
+\end{center}
+\end{table}
+
 \section{Data Model}
@@
 Because processed textbooks contain abstract knowledge instead of information about any specific hospital, all concepts are modelled as classes.
-We thus call our data model the \enquote{meta model}, as it is an ontology.
-meta model in accordance with its definition in the domain as a shared modelling language~\citep[p.~8]{ob}.
+We thus call our data model the \enquote{meta model} in accordance with the term's definition as a shared modelling language~\citep[p.~8]{ob}.
 The meta model (see \cref{fig:metamodel}) provides a common vocabulary for the domain of HIS management and thus defines, which superclasses and property can be used.
 SNIK version \snikversion{} comprises five subontologies that are built upon the meta model, see Table 1.
-At the head of the class hierarchy is the \enquote{Top} class, which has exactly three disjunctive subclasses.
+At the head of the class hierarchy (see \cref{fig:hierarchy}) is the \enquote{Top} class, which has exactly three disjunctive subclasses.
+
 Following the meta model, each class has to be a subclass of exactly one of them.
@@
 
-\section{Unsorted}
-
 \begin{table}
-\caption{Access Options and Details}
+\caption{Access Options and Details.%
+%KBox usage example: \texttt{pip install kbox \&\& kbox -kb http://www.snik.eu/ontology -install -server}.%doesn't work right now
+}
 \label{tab:access}
@@
 License					&CC BY-NC-SA 4.0\\
+Ontology Documentation	&\url{https://snikproject.github.io/ontology/}\\
 SPARQL Endpoint			&\url{http://www.snik.eu/sparql}\\
 LOD Cloud				&\url{https://lod-cloud.net/dataset/snik}\\
-Datahub		&\url{https://datahub.io/imise/snik/v/1}\\
+%Datahub				&\url{https://datahub.io/imise/snik/v/1}\\%offline
+%Docker
+KBox~\cite{kbox} Graph	&\url{http://www.snik.eu/ontology}\\
+RDF Browser				&\url{http://www.snik.eu/ontology}\\
 Visualization			&\url{http://www.snik.eu/graph}\\
-RDF Browser	&\url{http://www.snik.eu/ontology}\\
-Download	&\url{https://github.com/IMISE/snik-ontology/releases/download/1.0.0/snik.zip}\\
+Download				&\url{https://github.com/snikproject/ontology/archive/refs/tags/\snikversion.zip}\\
+Figshare				&\url{https://doi.org/10.6084/m9.figshare.17057084.v1}\\
 \bottomrule
@@
 \label{tab:statistics}
+\centering
 \begin{tabulary}{\columnwidth}{lS}
 \toprule
-Nr. of Triples			&\sniktriples{}\\
-Nr. of Classes			&\snikclasses{}\\
-Nr. of Properties		&\snikproperties{}\\
-Nr. of Internal Links		&360\\
-Nr. of Links to HITO  		&260\\
-Nr. of Links to DBpedia 	&\sniklinks{}\\
+\textbf{Category}		&\textbf{Count}\\
+\midrule
+Triples				&\sniktriples{}\\
+Classes				&\snikclasses{}\\
+Properties			&\snikproperties{}\\
+Internal Links		&360\\
+Links to HITO  		&260\\
+Links to DBpedia 	&\sniklinks{}\\
 %Subontologies	&\url{http://www.snik.eu/ontology/bb}\\
@@
 \label{fig:metamodel}
-\includegraphics[width=0.8\textwidth]{img/metamodel9s.pdf}
+\centering
+\includegraphics[width=\textwidth]{img/metamodel9s.pdf}
 \end{figure*}
-
-\includegraphics[width=\columnwidth]{img/hierarchy.pdf}
-\begin{center}
-\begin{tabular*}{\columnwidth}{ll}
-\toprule
-\textbf{Ontology}				&\textbf{Source}\\
-\midrule
-\url{http://www.snik.eu/ontology/meta}		&Meta Model\\
-\url{http://www.snik.eu/ontology/bb}		&Textbook~\cite{bb}\\
-\url{http://www.snik.eu/ontology/ob}		&Textbook~\cite{ob}\\
-\url{http://www.snik.eu/ontology/he}		&Textbook~\cite{he}\\
-\url{http://www.snik.eu/ontology/ciox}		&CIO Interview\\
-\url{http://www.snik.eu/ontology/it4it}		&Standard~\cite{it4it}\\
-\bottomrule
-\end{tabular*}
-\end{center}
 
@@
 
-\section{Data Model}\label{sec:architecture}
-The initial structure of the meta model is defined in \citet{domaene}.
-The \enquote{meta model} defines three basic disjunctive classes and their possible relations: Roles (who), Function (does what) and Entity Types (and which information is therefore needed).
-A set of modular subontologies define subclasses of those three classes and their relations as described by a certain knowledge source about information management in hospitals:
-The \textbf{Semantic Network of Information Management in Hospitals} (SNIK\footnote{Hospital means \enquote{Krankenhaus} in German.}) is a modular OWL 2 DL ontology.
+%\section{Data Model}\label{sec:architecture}
+%The initial structure of the meta model is defined in \citet{domaene}.
+%The \enquote{meta model} defines three basic disjunctive classes and their possible relations: Roles (who), Function (does what) and Entity Types (and which information is therefore needed).
+%A set of modular subontologies define subclasses of those three classes and their relations as described by a certain knowledge source about information management in hospitals:
+%The \textbf{Semantic Network of Information Management in Hospitals} (SNIK\footnote{Hospital means \enquote{Krankenhaus} in German.}) is a modular OWL 2 DL ontology.
 %from different sources:% three textbooks, an interview and a standard.
 
+\iffalse
 \section{Data Set Description}\label{sec:dsd}
@@
 \end{table*}
-
-The SNIK ontologies are available at \url{http://www.snik.eu/ontology/} under the Creative Commons Attribution-NonCommercial-ShareAlike 4.0 International license.
+\fi
 
@@
 %As shown in \cref{...}, we aligned our dataset to several others:
-We manually aligned our dataset to HITO, the Health IT Ontology~\citep{hito}
+We manually aligned our dataset to HITO, the Health IT Ontology~\citep{hitometh2}
 We automatically aligned our dataset to:
@@
 
-\section{Data Quality}
-
 \section{Applications}
@@
 Due to internal regulations, efforts to integrate them into the informational infrastructure of the Uniklinikum Leipzig were stopped.
-Consequently, the other approaches are in the area of teaching support:
+Consequently, the other approaches are in the area of teaching support.
+
+%\subsection{Reuse of Standard Components}
+%To focus the software development effort on the components described below, we reuse standard components for common use cases:
+%Virtuoso Opensource 7 serves as the SPARQL endpoint.
 
 \subsection{SNIK Graph}\label{sec:snik-graph}
-\paragraperformance
-due to the limitations of JavaScript (e.g.
-single threaded), which are hard to work around, which leads to performance problems with the library Cytoscape.js which is the base of the graph visualization, on very large ontologies / graphs.
-right now we have 2641 nodes and 4904 edges and the performance is ok in Chrome but not amazing.
-with 10000 nodes it would probably be too slow on most machines.
+%\paragraperformance
+
+Graph-based visualizations allow teachers to intuitively convey relationships between a selection of those concepts and also allow students to explore the domain on their own~\cite{ontologybased}.
+There are several existing graph-based Linked Data visualizations~\cite{linkeddatavisualization}, which visualize RDF resources (classes or instanes) as nodes and their relationships as edges, but they do not fit our requirements~\cite{visualizationoflargeontologies}.
+SNIK Graph is a web-based client-side open source web application\footnote{Published at \url{http://www.snik.eu/graph}.} that offers Linked Data visualization for both classes and instances and according to the requirements of SNIK and its users.
+With the thousands of classes of SNIK, graph-based visualization suffers from overplotting.
+Thus, SNIK Graph offers several options to select and layout subgraphs, for example to show only a specific chapter of a book to prepare a lecture about a specific topic.
+Users can also iteratively explore SNIK starting at a single class using neighbourhood and path operations. %(see \cref{fig:snik-graph-circle-star})
+
+\paragraph{Design Goals and Requirements}
+Our main goal is to visualize knowledge extracted from text books to users that may not have any Semantic Web experience.
+The time and cognitive load of the users required to install, learn and operate the application should be as low as possible, so that they can focus on the data at hand and experience a benefit compared to only reading the textbook, such as when studying for an exam.
+A web application is easiest to set up for the user, as it is operating system agnostic and does not need to be installed.
+
+\paragraph{Implementation and Setup}
+As none of the investigated existing approaches support all our use cases and design goals and provide adequate performance on more than 4000 nodes, we implemented SNIK Graph as a web application using JavaScript (EcmaScript 2015) based on the graph visualization library Cytoscape.js~\cite{cytoscape}.
+SNIK Graph is freely available as open source software\footnote{\url{https://github.com/imise/snik-cytoscape.js}}.
+As a client-side tool\footnotemark{}, SNIK Graph can be adapted for other knowledge bases and ontologies and published on any HTML web server by cloning the repository, installing the NPM dependencies and adapting the configuration file settings including the SPARQL endpoint URL.
+\footnotetext{Aside from a SPARQL endpoint.}%
+An installation containing SNIK is published at \url{https://www.snik.eu/graph}.
+
+%\paragraph{Features}
+\paragraph{Tabs and Containers}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{img/tabs.png}
+    \caption{A session containing the default view and freely partitioned subgraphs containers.}
+	\label{fig:tabs}
+\end{figure}
+\vspace{-3pt}
+%For this subgraphs of the ontology have to be extracted and organized.
+%A feature of SNIK Graph, tabs, allows multi-instance usage.
+%One major issue in working with SNIK Graph was the missing ability of organizing teaching scenarios.
+To organize teaching scenarios, independent SNIK Graph instances are loaded into containers\footnotemark{} that can be organized around the browser window.
+\footnotetext{Implemented using the JavaScript GoldenLayout framework (\url{https://golden-layout.com/}).} 
+These containers can either be accessed as tabs or freely partitioned see \cref{fig:tabs}.
+The content and of the all containers can be saved and loaded as a session to continue developing a teaching scenario at a later point.
+
+%This feature allows better workflows in preparation and presentation of teaching scenarios.
+%We use JSON formatted files for export and import.
+
+\paragraph{Compound Layout}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{img/combine.pdf}
+    \caption{Grouping of equivalent concepts from different sources in SNIK Graph (color coded).}% using the \enquote{combine matches} option.}
+	\label{fig:combine}
+\end{figure}
+\vspace{-3pt}
+%\noindent For citations of references, we prefer the use of square brackets and consecutive numbers, e.g.
+as shown by Author et al.
+[2], [3, pp.
+5–10], as mentioned earlier [1], [3], [9].
+The following bibliography provides the basic formats as a reference list with entries for journal articles [1], book chapter [2], as well as a URL [5].
+For further guidance please refer to the \href{https://ieeeauthorcenter.ieee.org/wp-content/uploads/IEEE-Reference-Guide.pdf}{IEEE-Reference-Guide}.
+
+Users can group interlinked nodes together and either display them side by side or on top of each other, see \cref{fig:combine}.
+As SNIK consists of more than 4000 classes, viewing large parts of it at once does not convey much information.%, see \cref{fig:much}.
+To alleviate this, SNIK offers multiple exploration methods described in the following.
+%\begin{figure}[h!]
+%    \centering
+%    \includegraphics[width=\linewidth]{much.pdf}
+%    \caption{Excerpt of SNIK using the default force directed layout.}\label{fig:much}
+%\end{figure}
+
+\paragraph{Role Use}
+\iffalse
+\todo{konrad an alle: was mache ich mit der sparql query?
+rausschmeissen, drin lassen oder umbauen?
+TP: drinlassen, aber direkt ans Bild hängen..
+Sowas wie aus der Query folgt dieses Bild.
+Ich finde sowas sollte schon rein ins Paper}
+\begin{figure}[h]
+\begin{lstlisting}
+SELECT DISTINCT ?inner ?middle ?outer ?outerx {
+ <role> (rdfs:subClassOf|skos:closeMatch|^skos:closeMatch)* ?inner.
+ ?inner meta:subTopClass meta:Role.
+ OPTIONAL {
+  ?inner ?p ?middle.
+  ?middle meta:subTopClass meta:Function.
+  OPTIONAL {
+   ?middle ?q ?outer.
+   ?outer meta:subTopClass meta:EntityType.
+   OPTIONAL {?outer (skos:closeMatch|^skos:closeMatch|^rdfs:subClassOf)+ ?outerx.}
+}}}
+\end{lstlisting}
+\caption{SPARQL query to generate the \emph{role use} visualization.}
+\label{lst:roleuse}
+\end{figure}
+\vspace{-3pt}
+\fi
+A frequent question is, what a given role does and which information is needed for those functions represented by the entity types connected to those functions.
+This question is visually answered by the \enquote{role use} feature, which arranges roles, functions and entity types in concentric circles, see \Cref{fig:classuse}:
+
+\begin{enumerate}[align=left, labelwidth=1ex]
+\item The first ring consists of the given role and all roles that are connected by a path of \aurl{rdfs}{subClassOf} and \aurl{skos}{closeMatch} edges.
+\item The function ring consists of all functions adjacent to roles in the inner ring.
+\item The third ring consists of all entity types adjacent to functions in the middle ring.
+\item The final ring consists of all entity types that are connected to entity types from the third ring by a path of \aurl{skos}{closeMatch} and reverse \aurl{rdfs}{subClassOf} edges.
+\end{enumerate}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\columnwidth]{img/class-use-project-manager.pdf}
+    \caption{\emph{Role Use} of \aurl{bb}{ProjectManager}. Outermost layer ommitted due to space limitiations.}\label{fig:classuse}
+\end{figure}
+
+%\section{Approach}
+%Due to its large size, visualizing all of SNIK at once  overplotting.
+%At the core of SNIK Graph is incremental exploration, which starts at one or more nodes and gradually expands them to related ones.
+
+\paragraph{Search}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\columnwidth]{img/search.pdf}
+    \caption{Visual search results for \enquote{Chief Information Officer} in a part of SNIK.}\label{fig:search}
+\end{figure}
+
+Due to the large amount of resources, exploration often begins with a search.
+The search index is populated from the target SPARQL endpoint and is implemented using the Fuse.js library that is based on the Baeza-Yates--Gonnet algorithm~\cite{textsearching}.
+Fuse.js\footnotemark{} is a light-weight, purely client side JavaScript library and presents an alternative to backend-driven indexes like Lucene.%
+\footnotetext{\url{https://fusejs.io/}, \url{https://github.com/krisk/Fuse}}
+This enables fast fuzzy search on any dataset loaded via SPARQL endpoint but requires waiting for index initialization on the first search of each user session.
+Search results presented to the user are color coded in three categories: visible (green), invisible (yellow)%
+\footnote{The resource is either \emph{filtered} or \emph{hidden}.}% \cref{sec:visibility}
+ and not loaded (red)%
+\footnote{The resource is included in the search index built from the SPARQL endpoint but either deleted in the graph or not loaded in the first place, such as due to configuration.}.
+Each search entry of a class contains the label values of \aurl{rdfs}{label} (long form) and \aurl{skos}{altLabel} (short and alternative forms) with a weight of 0.7.
+Textbook definitions are included with a weight of 0.3.
+Labels of resources connected via \aurl{skos}{closeMatch} interlinks in either direction are included as well, because those resources are defined as semantically close, so we assume their labels are synonyms.
+Fuzzy matching is enabled with a threshold of \num{0.25}.
+The resource \aurl{bb}{ChiefInformationOfficer} can thus be found by entering either \enquote{CIO} (alternative label), \enquote{vice president} (part of the definition) or \enquote{Leiter des Rechenzentrums} (German alternative label of \aurl{ob}{ChiefInformationOfficer}).
+Search results are then selected in the graph, see~\cref{fig:search}, which allows further exploration by chaining the path and neighbourhood operations described in following.
+
+\paragraph{Shortest Paths}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=\columnwidth]{img/path.png}
+    \caption{The shortest non-trivial path between \aurl{bb}{PatientIdentificationNumber} and \aurl{bb}{QualityOfData} in SNIK Graph.}
+	\label{fig:path}
+\end{figure}
+\vspace{-3pt}
+The shortest path is the most basic method of visualizing the relationship between two resources, see \cref{fig:path}.
+We treat SNIK as an undirected graph for the calculation because the direction is often arbitrary.
+Roles can be nested using \aurl{meta}{roleComponent} (\emph{has part}), which could have just as well been modelled as a \emph{part of} relation with reversed subject and object.
+Another reason is that a set of resources that forms a graph connected by some symmetric property implies a complete subgraph, which negatively effects speed, memory and visibility of SNIK Graph.
+For this reason, \aurl{skos}{closeMatch} and other symmetrical properties are often only sparsely modelled, and the other triples are implied.
+As SPARQL endpoints, such as Virtuoso SPARQL employed by SNIK, do not perform reasoning to infer such implied triples such as those generated by symmetric properties, this would prevent the shortest path from including resources where only the opposite direction is explicitly modelled.%
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\columnwidth]{img/hierarchy.pdf}
+    \caption{Excerpt of the SNIK class hierarchy. Source: \cite{snikposter}.}
+	\label{fig:hierarchy}
+\end{figure}
+\vspace{-3pt}
+We also don't explicitly store triples inferred by transitive properties such as \aurl{rdfs}{subClassOf}, which prevents paths involving such properties from being squashed, such as the path between \emph{Project Manager} and \emph{HIS Stakeholder} in \cref{fig:hierarchy}.
+
+%We ignore the direction of edges because, that is, which resource is the subject and the object 
+The largest problem of shortest paths is however, that they are not necessarily informative to the user. 
+For example, any two roles, such as \aurl{bb}{ChiefInformationOfficer} and \aurl{bb}{ChiefExecutiveOfficer}, are connected by a path of length 2 as they are all subclasses of \aurl{meta}{Role}, which the user already knows given the triangular shapes representing roles in SNIK Graph.
+To exclude such trivial paths,  the meta model is not shown by default.
+Furthermore, filtering, such as by knowledge source or type of relation or resource, allows further tuning of the resources that are shown and eliglible for paths.
+%Further research into informative paths has been done in two bachelor theses~\cite{}.
+An general approach to solve the problem of informative shortest paths is \emph{weighted shortest paths for RDF graphs (WiSP)}~\cite{wisp}.
+Future work includes adopting this approach to SNIK Graph and evaluating it with its users.
+Another approach is to show all short paths under a given length and let the user remove unwanted ones manually, as employed by \emph{Relfinder}~\cite{relfinder}.
+
+\subsection{Neighbourhood Operations}
+Exploration using neighbours, that is the successive uncovering of nodes adjacent to a starting node given by a user, is a common feature of tools such as \cite{lodlive} and \cite{vizlod}.%
+\subsection{Star}
+The directed and undirected star operations show nodes connected to selected nodes via all paths that contain at most one property other than \aurl{skos}{closeMatch}.
+The \emph{circle star} also rearranges the nodes using the force-directed layout locally on the currently visible subgraph.
+\Cref{fig:star} shows a mind map of strategic information management, created by an undirected star, which can be used by a teacher to prepare a lecture about the topic.
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\linewidth]{img/strategic-im.pdf}
+    \caption{Undirected \emph{Star} of \aurl{bb}{StrategicInformationManagement}.}
+	\label{fig:star}
+\end{figure}
+\vspace{-3pt}
+
+%\todo{Konrad: Nimm den Star um "ob/ProjectManagement".} % ist inkludiert in \cref{fig:tabs}, refer?
+The concept \emph{project management} both exists in a textbook~\cite{ob} and in the knowledge about information management in a German university hospital.
+Grouping both concepts with the help of the star function the user detects that the terms have a different meaning according to their neighboorhood terms.
+Whereas the text book deals with the management of single projects, in the CIO's world \enquote{project management} means managing multiple projects, as shown in the lower right container in \cref{fig:tabs}.
+As all neighbourhood and path operations automatically select all resulting nodes, repeated application of a star without clearing the selection results in an incremental uncovering of the whole connected graph around the starting node.
+
+%\subsection{Mixed Operations}
+\paragraph{Spiderworm}
+A \emph{spiderworm} is a path from node \emph{A} to node \emph{B} combined with a \emph{star} of \emph{B}.
+\Cref{fig:spiderworm} shows how we use a spiderworm to teach a student how the new concept \enquote{quality of data} is connected the the already introduced concept \enquote{patient identification number.}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.5\linewidth]{img/spiderworm.pdf}
+    \caption{\emph{Spiderworm} from \aurl{bb}{PatientIdentificationNumber} to \aurl{bb}{QualityOfData}~\cite{snikgraphposter}.}
+	\label{fig:spiderworm}
+\end{figure}
+\vspace{-3pt}
+
+\iffalse
+Due to the limitations of JavaScript, such as a single thread, and the used library library Cytoscape.js which is the base of the graph visualization, there are performance problems on large graphs.
+Right now we have 2641 nodes and 4904 edges and the performance is ok in Chrome but not amazing.
+With 10000 nodes it would probably be too slow on most machines.
 
@@
 \begin{figure}
-\caption{SNIK Graph Overview. Source: \protect\citet{snikgraph}.}
+\caption{SNIK Graph Overview. Source: \protect\citet{snikgraphposter}.}
 \label{fig:snik-graph-overview}
@@
 Path and neighourhood operations are joined in the \enquote{spider worm} (see \cref{fig:snik-graph-spiderworm}), which consists of the shortest path between a start node and an end node together with the end node’s neighbourhood, illustrate the context of a concept.
-
+\fi
 \subsection{Quiz}
@@
 It is unclear, whether the CIO is also responsible for operational information management.
-In the future, we will add more question types
+An ongoing bachelor's thesis~\citep{snikquiz} is investigating different strategies of generating complex questions, answers and distrators.
 
@@
 \label{tab:templates}
+\centering
 \begin{tabulary}{\textwidth}{lL}
@@
 \todo{FJ: Hmmm, ziemlich selbstreferenziell. Brauchen wir da wirklich so viele Publikationen von uns? Welche anderen könnte man noch einbringen? Ich schaue auch mal, was mir so über den Weg läuft.}
-\bibliographystyle{ios1}
-\bibliography{paper,snik}
-\section{Notes to Reviewers}
-This paper includes and extends text and figures from:
-\citet{snikgraph}
-
+%\bibliographystyle{ios1}
+\bibliographystyle{IEEEtran}
+\bibliography{paper,snik,relatedwork}
+%\section{Notes to Reviewers}
+%This paper includes and extends text and figures from:
+%\citet{snikgraph}
+%\todo{KH: Extending a poster should be allowed for a journal but maybe not for a conference.
+replace text and figures from the poster.
+should only be the snik graph chapter.
+use snik graph paper instead.}
+%Update: All the text from the paper should now be commented out or properly referenced.
 \end{document}

### 2021-11-21 paper.tex
Nearly finished alpha.

@@
 \usepackage{siunitx}
-\usepackage{todonotes}
+\usepackage[disable]{todonotes}
 %\usepackage{natbib} % does not seem to work with ios 1
@@
 Links to HITO  						&260\\
-Links to DBpedia 	&\sniklinks{}\\
+Links to DBpedia~\citep{dbpedia} 	&\sniklinks{}\\
 %Subontologies	&\url{http://www.snik.eu/ontology/bb}\\
@@
 
-\section{Data Transformation Process}
-\subsection{Interlinking}
+%\section{Data Transformation Process}
+%\subsection{Interlinking}
 %As shown in \cref{...}, we aligned our dataset to several others:
-We manually aligned our dataset to HITO, the Health IT Ontology~\citep{hitometh2}
-We automatically aligned our dataset to:
-- DBpedia~\citep{dbpedia}, as  
-- 
+%We manually aligned our dataset to HITO, the Health IT Ontology~\citep{hitometh2}
+%We automatically aligned our dataset to:
+%- DBpedia~\citep{dbpedia}, as  
+%- 
 \section{Lifecycle}\label{sec:application}
@@
     \centering
-    \includegraphics[width=\linewidth]{img/combine.pdf}
+    \includegraphics[width=\columnwidth]{img/combine.pdf}
     \caption{Grouping of equivalent concepts from different sources in SNIK Graph (color coded).}% using the \enquote{combine matches} option.}
@@
     \centering
-    \includegraphics[width=0.9\linewidth]{img/strategic-im.pdf}
+    \includegraphics[width=\columnwidth]{img/strategic-im.pdf}
     \caption{Undirected \emph{Star} of \aurl{bb}{StrategicInformationManagement}.}
@@
     \centering
-    \includegraphics[width=0.5\linewidth]{img/spiderworm.pdf}
+    \includegraphics[width=\columnwidth]{img/spiderworm.pdf}
     \caption{\emph{Spiderworm} from \aurl{bb}{PatientIdentificationNumber} to \aurl{bb}{QualityOfData}~\cite{snikgraphposter}.}
@@
 \subsection{Quiz}
+
+\begin{table}[t]
+\caption{Initial Quiz question templates.}
+\label{tab:templates}
+\centering
+\begin{tabulary}{\columnwidth}{lL}
+\toprule
+\textbf{Template}	&\textbf{Description}\\
+\midrule
+Definition		&Ask for the class that fits the given textbook definition.\\
+Example			&What is defined as \enquote{Examination of in and out patients in radiological department}?\\
+Distractors		&Labels of other classes (that have a path of length 2 or less to the correct class.)\\
+\midrule
+Subject			&Ask for the class that is related via a given relation to a given object.\\
+Example			&Who is \emph{involved in} a \emph{healthcare network}?\\
+Distractors		&Labels of other classes (of the same type) that \emph{are not} related via the same relation to the same object.\\
+\bottomrule
+\end{tabulary}
+\end{table}
+
 The DBpedia-powered Clover Quiz~\citep{cloverquiz} shows that ontologies can be used to automatically generate multiple-choice questions.
@@
 
-\begin{table*}
-\caption{Templates}
-\label{tab:templates}
-\centering
-\begin{tabulary}{\textwidth}{lL}
-\toprule
-\textbf{Template}	&\textbf{Description}\\
-\midrule
-Definition		&Ask for the class that fits the given textbook definition.\\
-Example			&What is defined as \enquote{Examination of in and out patients in radiological department}?\\
-Distractors		&Labels of other classes (that have a path of length 2 or less to the correct class.)\\
-\midrule
-Subject			&Ask for the class that is related via a given relation to a given object.\\
-Example			&Who is \emph{involved in} a \emph{healthcare network}?\\
-Distractors		&Labels of other classes (of the same type) that \emph{are not} related via the same relation to the same object.\\
-\bottomrule
-\end{tabulary}
-\end{table*}
-
 %The technical environment of SNIK is characterized in previous work [11].
+\iffalse
 \section{Discussion}
@@
 For example, the IT management of a hospital can use SNIK as a vocabulary to integrate data from different formats that result from application components from different vendors.
-Using the SPARQL1 query language, experts can also integrate applications with the SNIK ontology.
+Using the SPARQL query language, experts can also integrate applications with the SNIK ontology.
  Experts can also use the SPARQL query editor to answer specific questions, see Table 2.
@@
 Thus, it might be attractive for publishing houses.
-
+\fi
 \todo{verweis auf existierende papers}

### 2021-11-21 paper.tex
Finished alpha version and shortened to exactly 7 pages.

@@
 \title{SNIK: Textbook Knowledge on Hospital Information Management as Linked Data}
+%\thanks{This work is supported by the DFG (German Research Foundation) under the Project SNIK, Grant no. 1605/7-1 and 1387/8-1.}%bug: will create an empty first page
 %\runningtitle{The SNIK Ontology of Hospital Information Management}
@@
 %As the domain is large and highly relevant to students of Medical Informatics, modelling the knowledge is not only possible but also very useful.
-We publish the result over several different interfaces that are useful for researchers, information system administrators or students, depending on their objectives and their capabilities.
+We publish the result over several interfaces that are useful for researchers, administrators or students, depending on their objectives and capabilities.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-linked open data visualization,health information systems,information management
+linked open data, information management
 \end{IEEEkeywords}
@@
 %A structured representation of the different perspectives leads to a holistic view on HIS management and helps help researchers and students connect their existing knowledge with further knowledge from other sources during research and learning.
-In order to integrate different knowledge sources and to provide the knowledge in a structured, machine-readable data format, we extracted knowledge about HIS management from three textbooks~\citep{bb,ob,he} and other sources, see \cref{tab:source}.
+In order to integrate different knowledge sources and to provide the knowledge in a structured, machine-readable data format, we extracted knowledge about HIS management from several sources, see \cref{tab:source}.
+%three textbooks~\citep{bb,ob,he} and other sources, see \cref{tab:source}.
+%\section{Sources}\label{sec:sources}
+Three textbooks provide different views on the domain of Hospital Information Management:
+\citet{bb} presents a broad view on \enquote{typical architectures of health information systems and their systematic strategic management}.
+\citep{ob} concentrates on the \emph{tactical} management of information systems in general and on healthcare in particular.
+The focus of tactical management lies on the planning and operation of projects.
+\citet{he} explains information management beyond the scope of healthcare.
+Other sources are interviews and standards.
+%CIOX is based on an interview with the CIO of the Universitätsklinikum Leipzig.
+%IT4IT is based on the IT4IT standard.\todo{Sebastian: Was über Standards und IT4IT schreiben}
+
 The combination of this knowledge~\citep{semantischesnetz,domaene,approachtosupport} results in SNIK, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German).
@@
 These are only implicitly known or not known at all by experts in the field.
-
+%
 We present a common data model for the domain and the result of applying the data model to three textbooks.
 We publish the result over several interfaces that are useful for researchers, practitioners and students, depending on their objectives and their Semantic Web skills.
-
+%
 Publishing textbook knowledge as Linked Data enables different ways of teaching.
@@
 At the head of the class hierarchy (see \cref{fig:hierarchy}) is the \enquote{Top} class, which has exactly three disjunctive subclasses.
-
+%
 Following the meta model, each class has to be a subclass of exactly one of them.
@@
 Relations that are neither of them can either be modeled by using the generic \enquote{is associated with} relation or by creating and using a new sub relation of \enquote{is associated with}
-
+%
 As a concession to practicality, we express each geleral rule extracted from a textbook as a single triple using classes as subject and object, and a property of the meta model.
@@
 Additionally, encoding each fact as a single triple facilitates tool support and prevents inconsistencies through non-atomic changes.
-
-\paragraph{}
 
@@
 License					&CC BY-NC-SA 4.0\\
-Ontology Documentation	&\url{https://snikproject.github.io/ontology/}\\
+Documentation	&\url{https://snikproject.github.io/ontology/}\\
 SPARQL Endpoint			&\url{http://www.snik.eu/sparql}\\
@@
 Internal Links						&360\\
-Links to HITO  						&260\\
+Links to HITO~\citep{hitometh2}  	&260\\
 Links to DBpedia~\citep{dbpedia} 	&\sniklinks{}\\
@@
 \centering
-\includegraphics[width=\textwidth]{img/metamodel9s.pdf}
+\includegraphics[width=0.9\textwidth]{img/metamodel9s.pdf}
+%\includegraphics[width=\textwidth]{img/metamodel9s.pdf}
 \end{figure*}
-
-\section{Sources}\label{sec:sources}
-Three textbooks provide different views on the domain of Hospital Information Management:
-\citet{bb} presents a broad view on \enquote{typical architectures of health information systems and their systematic strategic management}.
-\citep{ob} concentrates on the \emph{tactical} management of information systems in general and on healthcare in particular.
-The focus of tactical management lies on the planning and operation of projects.
-\citet{he} explains information management beyond the scope of healthcare.
-Other sources are interviews and standards.
-%CIOX is based on an interview with the CIO of the Universitätsklinikum Leipzig.
-%IT4IT is based on the IT4IT standard.\todo{Sebastian: Was über Standards und IT4IT schreiben}
 
@@
 
-\begin{enumerate}[align=left, labelwidth=1ex]
+\begin{enumerate}%[align=left, labelwidth=1ex]
 \item The first ring consists of the given role and all roles that are connected by a path of \aurl{rdfs}{subClassOf} and \aurl{skos}{closeMatch} edges.
@@
 
-\subsection{Neighbourhood Operations}
+\paragraph{Neighbourhood Operations}
 Exploration using neighbours, that is the successive uncovering of nodes adjacent to a starting node given by a user, is a common feature of tools such as \cite{lodlive} and \cite{vizlod}.%
-\subsection{Star}
+\paragraph{Star}
 The directed and undirected star operations show nodes connected to selected nodes via all paths that contain at most one property other than \aurl{skos}{closeMatch}.
@@
 We circumvented the timeout by calculating the neighbour-relation as a first step and uploading it to the endpoint.
-
+%
 %SNIK Quiz is freely available as an open source web application.
@@
 
-\section{Acknowledgments}
-This work is supported by the DFG (German Research Foundation) under the Projects SNIK, grant no. 1605/7-1 and 1387/8-1, as well as HITO, grant no. WI 1605/11-1 and I3726-N31.
+%\section{Acknowledgments}
+%This work is supported by the DFG (German Research Foundation) under the Projects SNIK, grant no. 1605/7-1 and 1387/8-1, as well as HITO, grant no. WI 1605/11-1 and I3726-N31.
 %We thank the publishers ... and ... for granting us permission to publish the subontologies based on the textbooks \citet{bb,ob,he} under an open license.
@@
 %\bibliographystyle{ios1}
+\bstctlcite{IEEE:BSTcontrol}
 \bibliographystyle{IEEEtran}

### 2021-11-23 paper.tex
Small corrections and improvements for camera ready version after submission.

@@
 This includes its different application systems, computers, and network components as well as their users.
-A HIS processes data, information, and knowledge and its management involves planning, monitoring and directing of those activities.
+A HIS processes data, information, and knowledge and its management involves planning, monitoring and directing those activities.
 Due to the complexity and the unique conditions in health care, HIS management is an challenging task.
@@
 %IT4IT is based on the IT4IT standard.\todo{Sebastian: Was über Standards und IT4IT schreiben}
-
+%
 The combination of this knowledge~\citep{semantischesnetz,domaene,approachtosupport} results in SNIK, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German).
@@
 Version					&\snikversion{}\\
-License					&CC BY-NC-SA 4.0\\
+%License					&CC BY-NC-SA 4.0\\
+License					&Attribution-NonCommercial-ShareAlike 4.0 International (CC BY-NC-SA 4.0)\\
 Documentation	&\url{https://snikproject.github.io/ontology/}\\
@@
 \centering
-\includegraphics[width=0.9\textwidth]{img/metamodel9s.pdf}
-%\includegraphics[width=\textwidth]{img/metamodel9s.pdf}
+%\includegraphics[width=0.9\textwidth]{img/metamodel9s.pdf}%7 page version
+\includegraphics[width=\textwidth]{img/metamodel9s.pdf}
 \end{figure*}
@@
 Some project members are not proficient in RDF serializations, so we first fill out a meta model conforming spreadsheet template.
-We then use Tarql\footnote{\url{https://github.com/cygri/tarql}} with a mapping configuration file\footnote{\url{https://github.com/IMISE/snik-csv2rdf}} to convert that spreadsheet to RDF.
+We then use Tarql\footnote{\url{https://github.com/cygri/tarql}} with a mapping configuration file\footnote{\url{https://github.com/snikproject/snik-csv2rdf}} to convert that spreadsheet to RDF.
 Next, we refine the data in Protégé and upload it to a Virtuso SPARQL endpoint.
@@
 \item an instance of the OntoWiki~\citep{ontowiki} is used for punctual changes and additions by non-Semantic-Web-experts, supplemented by SPARUL for changes involving large numbers of classes 
-\item LIMES~\citep{limes} generates interlink candidates between the different textbooks, which can be approved or rejected by the users
+\item LIMES~\citep{limes} generates interlink candidates between the different textbooks, which can be approved or rejected by the users.
+\aurl{skos}{closeMatch} is used instead of \aurl{owl}{equivalentClass} to emphasize that not only the names of concepts may differ between textbooks but there may be subtle differences in meaning as well.
+We do not intent to express different treatment, such as reasoning, due to the theoretical difference between \aurl{skos}{Concept} and \aurl{owl}{Class}.
+Links to HITO and DBpedia are created as well, see \cref{tab:statistics}.
 \item LodView\footnote{\url{https://lodview.it/}} lets the user browse among the classes of SNIK and view particular classes in detail
@@
 Graph-based visualizations allow teachers to intuitively convey relationships between a selection of those concepts and also allow students to explore the domain on their own~\cite{ontologybased}.
-There are several existing graph-based Linked Data visualizations~\cite{linkeddatavisualization}, which visualize RDF resources (classes or instanes) as nodes and their relationships as edges, but they do not fit our requirements~\cite{visualizationoflargeontologies}.
+There are several existing graph-based Linked Data visualizations~\cite{linkeddatavisualization}, which visualize RDF resources (classes or instanes) as nodes and their relationships as edges, but they do not fit the requirements we have previously identified regarding performance, maintenance, usability, functionality and up-to-dateness~\cite{visualizationoflargeontologies}.
 SNIK Graph is a web-based client-side open source web application\footnote{Published at \url{http://www.snik.eu/graph}.} that offers Linked Data visualization for both classes and instances and according to the requirements of SNIK and its users.
@@
 As none of the investigated existing approaches support all our use cases and design goals and provide adequate performance on more than 4000 nodes, we implemented SNIK Graph as a web application using JavaScript (EcmaScript 2015) based on the graph visualization library Cytoscape.js~\cite{cytoscape}.
-SNIK Graph is freely available as open source software\footnote{\url{https://github.com/imise/snik-cytoscape.js}}.
-As a client-side tool\footnotemark{}, SNIK Graph can be adapted for other knowledge bases and ontologies and published on any HTML web server by cloning the repository, installing the NPM dependencies and adapting the configuration file settings including the SPARQL endpoint URL.
+SNIK Graph is freely available as open source software\footnote{Git repository at \url{https://github.com/snikproject/snik-graph}, licensed under CC BY-NC-SA 4.0}.
+As a client-side tool\footnotemark{}, SNIK Graph can be adapted to other knowledge bases and ontologies and published on any HTML web server by cloning the repository, installing the NPM dependencies and adapting the configuration file settings including the SPARQL endpoint URL.
 \footnotetext{Aside from a SPARQL endpoint.}%
@@
 \end{lstlisting}
-\caption{SPARQL query to generate the \emph{role use} visualization.}
+\caption{SPARQL query to generate the \emph{class use} visualization.}
 \label{lst:roleuse}
@@
 \end{figure}
+In SNIK Graph, this operation is called \emph{class use} and allows users to also put a function or entity type in the center.
 
@@
 \vspace{-3pt}
-We also don't explicitly store triples inferred by transitive properties such as \aurl{rdfs}{subClassOf}, which prevents paths involving such properties from being squashed, such as the path between \emph{Project Manager} and \emph{HIS Stakeholder} in \cref{fig:hierarchy}.
+We also do not explicitly store triples inferred by transitive properties such as \aurl{rdfs}{subClassOf}, which prevents paths involving such properties from being squashed, such as the path between \emph{Project Manager} and \emph{HIS Stakeholder} in \cref{fig:hierarchy}.
 
@@
 \paragraph{Neighbourhood Operations}
-Exploration using neighbours, that is the successive uncovering of nodes adjacent to a starting node given by a user, is a common feature of tools such as \cite{lodlive} and \cite{vizlod}.%
+Exploration using neighbours, that is the successive uncovering of nodes adjacent to a starting node given by a user, is a common feature of tools such as LodLive~\cite{lodlive} and VizLOD~\cite{vizlod}.%
 \paragraph{Star}
@@
 
-The DBpedia-powered Clover Quiz~\citep{cloverquiz} shows that ontologies can be used to automatically generate multiple-choice questions.
+Clover Quiz~\citep{cloverquiz} shows that ontologies can be used to automatically generate multiple-choice questions over DBpedia~\citep{dbpedia}.
 We apply this approach to SNIK and generate 1231 English questions using the templates described in \cref{tab:templates}.
@@
 So this problem cannot be avoided.
-However, this is one of the reasons that we don't ask count questions like \enquote{How many functions is the CIO responsible for?}, which don't help much for learning anyways. % improve writing
+However, this is one of the reasons that we do not ask count questions like \enquote{How many functions is the CIO responsible for?}, which do not help much for learning anyways. % improve writing
 Relationships could also hold implicitly through the subclass hierarchy.

### 2021-11-28 paper.tex
Further corrections.

@@
 \newcommand{\snikversion}{1.3.0}
-\newcommand{\sniktriples}{72046}
-\newcommand{\snikclasses}{2655}
-\newcommand{\snikproperties}{339}
-\newcommand{\sniklinks}{581}
+\newcommand{\snikversionlink}{\href{https://github.com/snikproject/ontology/releases/tag/\snikversion}{\snikversion}}
+\newcommand{\sniktriples}{81499}
+% SELECT (COUNT(DISTINCT *) AS ?count) {?s a owl:Class. FILTER(STRSTARTS(STR(?s),"http://www.snik.eu/ontology/"))}
+\newcommand{\snikclasses}{4107}
+% SELECT (COUNT(DISTINCT ?s) AS ?count) {{?s a rdf:Property.} UNION {?s a owl:ObjectProperty.} UNION {?s a owl:DataTypeProperty.} FILTER(STRSTARTS(STR(?s),"http://www.snik.eu/ontology/"))}
+\newcommand{\snikproperties}{65}
+%SELECT (COUNT(DISTINCT *) AS ?links) {?s ?p ?o. FILTER(STRSTARTS(STR(?o),"http://dbpedia")) }
+\newcommand{\sniklinks}{579}
 %%%%%%%%%%% End of definitions
@@
 
-\begin{table}[t!]
-\caption{The different subontologies of SNIK and their sources.
-The URL of each subontology is equal to its graph name, the namespace has an additional slash \enquote{/} suffix.
+\begin{table}[tbh]
+\caption{The different ontologies and graphs of SNIK and their sources.
+The namespace of each ontology is is equal to its URL with a slash suffix.
+The graph group URL has no associated ontology and is the entry point of the RDF browser.
 %All graphs are combined in the graph group \lowercase{\url{http://www.snik.eu/ontology}}.
@@
 \toprule
-\textbf{Ontology}				&\textbf{Source}					&\textbf{Prefix}\\
+\textbf{Prefix}	&\textbf{Ontology and Graph URL}			&\textbf{Source}\\		
 \midrule
-\url{http://www.snik.eu/ontology/meta}		&Meta Model				&meta\\
-\url{http://www.snik.eu/ontology/bb}		&Textbook~\cite{bb}		&bb\\
-\url{http://www.snik.eu/ontology/ob}		&Textbook~\cite{ob}		&ob\\
-\url{http://www.snik.eu/ontology/he}		&Textbook~\cite{he}		&he\\
-\url{http://www.snik.eu/ontology/ciox}		&CIO Interview			&ciox\\
-\url{http://www.snik.eu/ontology/it4it}		&Standard~\cite{it4it}	&it4it\\
-\url{http://www.snik.eu/ontology}			&Combined Graph Group	&sniko\\
+meta			&\url{http://www.snik.eu/ontology/meta}		&Meta Model\\
+bb				&\url{http://www.snik.eu/ontology/bb}		&Textbook~\cite{bb}\\
+ob				&\url{http://www.snik.eu/ontology/ob}		&Textbook~\cite{ob}\\
+he				&\url{http://www.snik.eu/ontology/he}		&Textbook~\cite{he}\\
+ciox			&\url{http://www.snik.eu/ontology/ciox}		&CIO Interview\\
+it4it			&\url{http://www.snik.eu/ontology/it4it}	&Standard~\cite{it4it}\\
+\midrule
+sniko			&\url{http://www.snik.eu/ontology}			&Combined Graph Group\\
 \bottomrule
@@
 \section{Data Model}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.8\columnwidth]{img/hierarchy.pdf}
+    \caption{Excerpt of the SNIK class hierarchy. Source: \cite{snikposter}.}
+	\label{fig:hierarchy}
+\end{figure}
+\vspace{-3pt}
 In order to specify, which information should be extracted from the books and to facilitate comparisons, we use a common data model.
@@
 The meta model (see \cref{fig:metamodel}) provides a common vocabulary for the domain of HIS management and thus defines, which superclasses and property can be used.
-SNIK version \snikversion{} comprises five subontologies that are built upon the meta model, see Table 1.
+SNIK version \snikversionlink{} comprises five subontologies that are built upon the meta model, see Table 1.
 At the head of the class hierarchy (see \cref{fig:hierarchy}) is the \enquote{Top} class, which has exactly three disjunctive subclasses.
@@
 URL						&\url{http://www.snik.eu/ontology}\\
-Version					&\snikversion{}\\
+Version					&\snikversionlink{}\\
 %License					&CC BY-NC-SA 4.0\\
-License					&Attribution-NonCommercial-ShareAlike 4.0 International (CC BY-NC-SA 4.0)\\
+License					&Attribution-NonCommercial-ShareAlike 4.0 International (\href{https://creativecommons.org/licenses/by-nc-sa/4.0/}{CC BY-NC-SA 4.0})\\
 Documentation	&\url{https://snikproject.github.io/ontology/}\\
@@
 As SPARQL endpoints, such as Virtuoso SPARQL employed by SNIK, do not perform reasoning to infer such implied triples such as those generated by symmetric properties, this would prevent the shortest path from including resources where only the opposite direction is explicitly modelled.%
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.8\columnwidth]{img/hierarchy.pdf}
-    \caption{Excerpt of the SNIK class hierarchy. Source: \cite{snikposter}.}
-	\label{fig:hierarchy}
-\end{figure}
-\vspace{-3pt}
 We also do not explicitly store triples inferred by transitive properties such as \aurl{rdfs}{subClassOf}, which prevents paths involving such properties from being squashed, such as the path between \emph{Project Manager} and \emph{HIS Stakeholder} in \cref{fig:hierarchy}.

### 2022-01-21 paper.tex
Add todo.
Shorten and adapt to MIE.

@@
-% add. options: [seceqn,secthm,crcready,onecolumn]
-%\documentclass[sw]{iosart2x}
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-\usepackage[utf8]{inputenc}
-\usepackage[english]{babel}
+\documentclass{IOS-Book-Article}     %[seceqn,secfloat,secthm]
+\usepackage{mathptmx}
+\usepackage{soul}\setuldepth{article}
+
+%\normalfont
+%\usepackage[T1]{fontenc}
+%\usepackage{times}
+%\usepackage[mtplusscr,mtbold]{mathtime}
+\def\hb{\hbox to 11.5 cm{}}
+%\usepackage[utf8]{inputenc}
+%\usepackage[english]{babel}
 \usepackage{graphicx}
 \usepackage{textcomp}
-\usepackage{xcolor}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-%\usepackage{dcolumn}
-%\usepackage{endnotes}
+%\usepackage{xcolor}
 \usepackage{booktabs}
@@
 \newcommand{\sniklinks}{579}
-%%%%%%%%%%% End of definitions
-
-%\pubyear{2020}
-%\volume{0}
-%\firstpage{1}
-%\lastpage{1}
 
@@
 
-%\begin{frontmatter}
+\pagestyle{headings}
+\def\thepage{}
 
-%\pretitle{}
+\begin{frontmatter}              % The preamble begins here.
+
+%\pretitle{Pretitle}
 \title{SNIK: Textbook Knowledge on Hospital Information Management as Linked Data}
 %\thanks{This work is supported by the DFG (German Research Foundation) under the Project SNIK, Grant no. 1605/7-1 and 1387/8-1.}%bug: will create an empty first page
-%\runningtitle{The SNIK Ontology of Hospital Information Management}
+
+\markboth{}{May 2022\hb}
 %\subtitle{}
-\iffalse
-% Two or more authors:
-\author[A]{\inits {K.}\fnms{Konrad} \snm{Höffner}\ead[label=e1]{konrad.hoeffner@imise.uni-leipzig.de}%
-\thanks{Corresponding author. \printead{e1}.}},
-\author[A]{\fnms{Franziska} \snm{Jahn}\ead[label=e2]{franziska.jahn@imise.uni-leipzig.de}},
-\author[A]{\fnms{Birgit} \snm{Schneider}\ead[label=e3]{birgit.schneider@imise.uni-leipzig.de}},
-%\author[A]{\fnms{Anna} \snm{Lörke}\ead[label=e4]{anna.loerke@imise.uni-leipzig.de}},
-\author[A]{\fnms{Thomas} \snm{Pause}\ead[label=e5]{thomas.pause@imise.uni-leipzig.de}},
-\author[A]{\fnms{Alfred} \snm{Winter}\ead[label=e6]{alfred.winter@imise.uni-leipzig.de}}
-%\runauthor{K. Höffner et al.}
-\address[A]{Institute for Medical Informatics, Statistics and Epidemiology (IMISE),
-\orgname{University of Leipzig}, \cny{Germany}\printead[presep={\\}]{e1,e2,e3,e5,e6}}
+
+\author[A]{\fnms{Konrad} \snm{Höffner}%
+%\thanks{Corresponding author: Konrad Höffner, Institute for Medical Informatics, Statistics and Epidemiology, Leipzig University,
+\thanks{Corresponding author: Konrad Höffner, IMISE, Leipzig University,
+Härtelstraße 16--18, 04107 Leipzig, Germany; E-mail: konrad.hoeffner@imise.uni-leipzig.de.}},
+\author[A]{\fnms{Franziska} \snm{Jahn}},
+\author[A]{\fnms{Birgit} \snm{Schneider}},
+\author[A]{\fnms{Thomas} \snm{Pause}},
+\author[A]{\fnms{Alfred} \snm{Winter}}
+\runningauthor{K. Höffner et al.}
+%\address[A]{Institute for Medical Informatics, Statistics and Epidemiology (IMISE), Leipzig University, Germany}
+\address[A]{IMISE, Leipzig University, Germany}
 %Medical Informatics, Management of Health Information Systemsi
 %Härtelstraße 16--18, D-04107 Leipzig
-\fi
 
+\iffalse
 \author{\IEEEauthorblockN{Konrad Höffner}
@@
 }
-\maketitle
+\fi
 \begin{abstract}
@@
 %\end{keyword}
+% ICSCR style
+%\begin{IEEEkeywords}
+%linked open data, information management
+%\end{IEEEkeywords}
 
-% ICSCR style
-\begin{IEEEkeywords}
-linked open data, information management
-\end{IEEEkeywords}
-%\end{frontmatter}
+% MIE style
+\begin{keyword}
+linked open data \sep information management
+\end{keyword}
+\end{frontmatter}
 
@@
 
-\section{Data Model}
+\section{Methods}
+\iffalse
 \begin{figure}[h]
     \centering
-    \includegraphics[width=0.8\columnwidth]{img/hierarchy.pdf}
+    \includegraphics[width=0.5\columnwidth]{img/hierarchy.pdf}
     \caption{Excerpt of the SNIK class hierarchy. Source: \cite{snikposter}.}
@@
 \end{figure}
+\fi
 \vspace{-3pt}
@@
 SNIK version \snikversionlink{} comprises five subontologies that are built upon the meta model, see Table 1.
-At the head of the class hierarchy (see \cref{fig:hierarchy}) is the \enquote{Top} class, which has exactly three disjunctive subclasses.
+%At the head of the class hierarchy (see \cref{fig:hierarchy}) is the \enquote{Top} class, which has exactly three disjunctive subclasses.
+At the head of the class hierarchy is the \enquote{Top} class, which has exactly three disjunctive subclasses.
 %
@@
 
+\iffalse
 \begin{table}
@@
 \end{table}
+\fi
 
@@
 %\includegraphics[width=0.9\textwidth]{img/metamodel9s.pdf}%7 page version
-\includegraphics[width=\textwidth]{img/metamodel9s.pdf}
+\includegraphics[width=0.7\textwidth]{img/metamodel9s.pdf}
 \end{figure*}
@@
 %- 
-\section{Lifecycle}\label{sec:application}
-
+%\section{Lifecycle}\label{sec:application}
+\section{Results}\label{sec:application}
+\iffalse
 \begin{figure}[ht]
@@
 \label{fig:lifecycle}
-\includegraphics[width=\columnwidth]{cycle.pdf}
+\includegraphics[width=0.5\columnwidth]{cycle.pdf}
 \end{figure}
-
+\fi
 When we add a new textbook to SNIK, we identify all concepts that fit the meta model and their relations to other concepts.
@@
 Next, we refine the data in Protégé and upload it to a Virtuso SPARQL endpoint.
-
-After the initial publish phase, we describe the repeating steps of the reuse phrase with the SNIK lifecycle in \Cref{fig:lifecycle}:
+%After the initial publish phase, we describe the repeating steps of the reuse phrase with the SNIK lifecycle in \Cref{fig:lifecycle}:
 \begin{enumerate}
@@
 We do not intent to express different treatment, such as reasoning, due to the theoretical difference between \aurl{skos}{Concept} and \aurl{owl}{Class}.
-Links to HITO and DBpedia are created as well, see \cref{tab:statistics}.
+Links to HITO and DBpedia are created as well.%, see \cref{tab:statistics}.
 \item LodView\footnote{\url{https://lodview.it/}} lets the user browse among the classes of SNIK and view particular classes in detail
@@
 The technical environment of those services is described in~\citet{sniktec}.
-Selected applications are described in the following section.
+Selected applications are described in the following.
 
-\section{Applications}
+%\section{Applications}
 SNIK was initially intended~\citep{domaene} both as  software to support hospital CIOs and to support teaching.
@@
 \subsection{SNIK Graph}\label{sec:snik-graph}
+%TODO: Search auto complete and suggestions
 %\paragraperformance
@@
 \paragraph{Tabs and Containers}
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{img/tabs.png}
-    \caption{A session containing the default view and freely partitioned subgraphs containers.}
-	\label{fig:tabs}
-\end{figure}
-\vspace{-3pt}
+%\begin{figure}[h]
+%    \centering
+%    \includegraphics[width=0.5\columnwidth]{img/tabs.png}
+%    \caption{A session containing the default view and freely partitioned subgraphs containers.}
+%	\label{fig:tabs}
+%\end{figure}
+%\vspace{-3pt}
+
 %For this subgraphs of the ontology have to be extracted and organized.
@@
 \footnotetext{Implemented using the JavaScript GoldenLayout framework (\url{https://golden-layout.com/}).} 
-These containers can either be accessed as tabs or freely partitioned see \cref{fig:tabs}.
+These containers can either be accessed as tabs or freely partitioned.% see \cref{fig:tabs}.
 The content and of the all containers can be saved and loaded as a session to continue developing a teaching scenario at a later point.
@@
 \paragraph{Compound Layout}
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\columnwidth]{img/combine.pdf}
-    \caption{Grouping of equivalent concepts from different sources in SNIK Graph (color coded).}% using the \enquote{combine matches} option.}
-	\label{fig:combine}
-\end{figure}
-\vspace{-3pt}
+%\begin{figure}[h]
+%    \centering
+%    \includegraphics[width=0.5\columnwidth]{img/combine.pdf}
+%    \caption{Grouping of equivalent concepts from different sources in SNIK Graph (color coded).}% using the \enquote{combine matches} option.}
+%	\label{fig:combine}
+%\end{figure}
+%\vspace{-3pt}
+
 %\noindent For citations of references, we prefer the use of square brackets and consecutive numbers, e.g.
@@
 
-Users can group interlinked nodes together and either display them side by side or on top of each other, see \cref{fig:combine}.
+Users can group interlinked nodes together and either display them side by side or on top of each other.%, see \cref{fig:combine}.
 As SNIK consists of more than 4000 classes, viewing large parts of it at once does not convey much information.%, see \cref{fig:much}.
@@
 \paragraph{Role Use}
-\iffalse
-\todo{konrad an alle: was mache ich mit der sparql query?
+
+%\todo{konrad an alle: was mache ich mit der sparql query?
 rausschmeissen, drin lassen oder umbauen?
@@
 Ich finde sowas sollte schon rein ins Paper}
-\begin{figure}[h]
-\begin{lstlisting}
-SELECT DISTINCT ?inner ?middle ?outer ?outerx {
- <role> (rdfs:subClassOf|skos:closeMatch|^skos:closeMatch)* ?inner.
- ?inner meta:subTopClass meta:Role.
- OPTIONAL {
-  ?inner ?p ?middle.
-  ?middle meta:subTopClass meta:Function.
-  OPTIONAL {
-   ?middle ?q ?outer.
-   ?outer meta:subTopClass meta:EntityType.
-   OPTIONAL {?outer (skos:closeMatch|^skos:closeMatch|^rdfs:subClassOf)+ ?outerx.}
-}}}
-\end{lstlisting}
-\caption{SPARQL query to generate the \emph{class use} visualization.}
-\label{lst:roleuse}
-\end{figure}
-\vspace{-3pt}
-\fi
+%\begin{figure}[h]
+%\begin{lstlisting}
+%SELECT DISTINCT ?inner ?middle ?outer ?outerx {
+% <role> (rdfs:subClassOf|skos:closeMatch|^skos:closeMatch)* ?inner.
+% ?inner meta:subTopClass meta:Role.
+% OPTIONAL {
+%  ?inner ?p ?middle.
+%  ?middle meta:subTopClass meta:Function.
+%  OPTIONAL {
+%   ?middle ?q ?outer.
+%   ?outer meta:subTopClass meta:EntityType.
+%   OPTIONAL {?outer (skos:closeMatch|^skos:closeMatch|^rdfs:subClassOf)+ ?outerx.}
+%}}}
+%\end{lstlisting}
+%\caption{SPARQL query to generate the \emph{class use} visualization.}
+%\label{lst:roleuse}
+%\end{figure}
+%\vspace{-3pt}
+
 A frequent question is, what a given role does and which information is needed for those functions represented by the entity types connected to those functions.
-This question is visually answered by the \enquote{role use} feature, which arranges roles, functions and entity types in concentric circles, see \Cref{fig:classuse}:
+This question is visually answered by the \enquote{role use} feature, which arranges roles, functions and entity types in concentric circles.%, see \Cref{fig:classuse}:
 
@@
 \end{enumerate}
+
+\iffalse
 \begin{figure}[h!]
     \centering
-    \includegraphics[width=\columnwidth]{img/class-use-project-manager.pdf}
+    \includegraphics[width=0.5\columnwidth]{img/class-use-project-manager.pdf}
     \caption{\emph{Role Use} of \aurl{bb}{ProjectManager}. Outermost layer ommitted due to space limitiations.}\label{fig:classuse}
 \end{figure}
+\fi
+
 In SNIK Graph, this operation is called \emph{class use} and allows users to also put a function or entity type in the center.
@@
 \paragraph{Search}
+\iffalse
 \begin{figure}[h!]
     \centering
-    \includegraphics[width=\columnwidth]{img/search.pdf}
+    \includegraphics[width=0.5\columnwidth]{img/search.pdf}
     \caption{Visual search results for \enquote{Chief Information Officer} in a part of SNIK.}\label{fig:search}
 \end{figure}
-
+\fi
 Due to the large amount of resources, exploration often begins with a search.
@@
 The resource \aurl{bb}{ChiefInformationOfficer} can thus be found by entering either \enquote{CIO} (alternative label), \enquote{vice president} (part of the definition) or \enquote{Leiter des Rechenzentrums} (German alternative label of \aurl{ob}{ChiefInformationOfficer}).
-Search results are then selected in the graph, see~\cref{fig:search}, which allows further exploration by chaining the path and neighbourhood operations described in following.
+%Search results are then selected in the graph, see~\cref{fig:search}, which allows further exploration by chaining the path and neighbourhood operations described in following.
+Search results are then selected in the graph, which allows further exploration by chaining the path and neighbourhood operations described in following.
 
@@
 The \emph{circle star} also rearranges the nodes using the force-directed layout locally on the currently visible subgraph.
-\Cref{fig:star} shows a mind map of strategic information management, created by an undirected star, which can be used by a teacher to prepare a lecture about the topic.
+%\Cref{fig:star} shows a mind map of strategic information management, created by an undirected star, which can be used by a teacher to prepare a lecture about the topic.
+\iffalse
 \begin{figure}[h!]
@@
 \vspace{-3pt}
+\fi
 
@@
 A \emph{spiderworm} is a path from node \emph{A} to node \emph{B} combined with a \emph{star} of \emph{B}.
-\Cref{fig:spiderworm} shows how we use a spiderworm to teach a student how the new concept \enquote{quality of data} is connected the the already introduced concept \enquote{patient identification number.}
+%\Cref{fig:spiderworm} shows how we use a spiderworm to teach a student how the new concept \enquote{quality of data} is connected the the already introduced concept \enquote{patient identification number.}
 
+\iffalse
 \begin{figure}[h!]
@@
 
-\section{Conclusions}
+\section{Discussion}
 We showed that knowledge on the management of information systems in medicine and health care is made publicly available using open standards over several interfaces with different compromises between expressivity and accessibility.
@@
 
-\section{Future Work}
-We plan to extend the text-based interlinks with semantic-similarity-based ontology matching [5] to find more correspondencies between the textbooks.
+%\section{Future Work}
+\section{Conclusions}
+In the future, we plan to extend the text-based interlinks with semantic-similarity-based ontology matching [5] to find more correspondencies between the textbooks.
 This will enable users to switch to another textbook when a certain concept is explained there in more detail.
@@
 We are also investigating candidates for additional subontologies and as interlink targets.
-
 Past efforts of automatic textbook extractions were unsuccessfull but recent advances in natural language processing, such as GPT-3~\citep{gpt3}, motivate further investigation.
@@
 %\bibliographystyle{ios1}
-\bstctlcite{IEEE:BSTcontrol}
-\bibliographystyle{IEEEtran}
+\bibliographystyle{vancouver}
 \bibliography{paper,snik,relatedwork}
@@
 %\citet{snikgraph}
-%\todo{KH: Extending a poster should be allowed for a journal but maybe not for a conference.
-replace text and figures from the poster.
-should only be the snik graph chapter.
-use snik graph paper instead.}
-%Update: All the text from the paper should now be commented out or properly referenced.
 \end{document}

### 2022-01-21 paper.tex
Shorten paper.

@@
 %A structured representation of the different perspectives leads to a holistic view on HIS management and helps help researchers and students connect their existing knowledge with further knowledge from other sources during research and learning.
-In order to integrate different knowledge sources and to provide the knowledge in a structured, machine-readable data format, we extracted knowledge about HIS management from several sources, see \cref{tab:source}.
+In order to integrate different knowledge sources and to provide the knowledge in a structured, machine-readable data format, we extracted knowledge about HIS management from several sources.%, see \cref{tab:source}.
 %three textbooks~\citep{bb,ob,he} and other sources, see \cref{tab:source}.

### 2022-01-21 paper.tex
Add initial SNIK Quiz paper draft.

@@
 We thus call our data model the \enquote{meta model} in accordance with the term's definition as a shared modelling language~\citep[p.~8]{ob}.
-The meta model (see \cref{fig:metamodel}) provides a common vocabulary for the domain of HIS management and thus defines, which superclasses and property can be used.
+The meta model (see \cref{fig:metamodel}) provides a common vocabulary for the domain of HIS management and thus defines, which superclasses and properties can be used.
 SNIK version \snikversionlink{} comprises five subontologies that are built upon the meta model, see Table 1.

### 2022-01-21 quiz.tex
Add initial SNIK Quiz paper draft.

--- 2022-01-21 paper.tex
@@
 %\pretitle{Pretitle}
-\title{SNIK: Textbook Knowledge on Hospital Information Management as Linked Data}
+\title{SNIK Quiz: Automatic Generation of Multiple Choice Questionnaires about Information Management in Hospitals}
 %\thanks{This work is supported by the DFG (German Research Foundation) under the Project SNIK, Grant no. 1605/7-1 and 1387/8-1.}%bug: will create an empty first page
@@
 Härtelstraße 16--18, 04107 Leipzig, Germany; E-mail: konrad.hoeffner@imise.uni-leipzig.de.}},
+\author[A]{\fnms{Arne} \snm{Roszeitis}},
+\author[A]{\fnms{Max Niclas} \snm{Wächtler}},
 \author[A]{\fnms{Franziska} \snm{Jahn}},
-\author[A]{\fnms{Birgit} \snm{Schneider}},
-\author[A]{\fnms{Thomas} \snm{Pause}},
 \author[A]{\fnms{Alfred} \snm{Winter}}
@@
 %\address[A]{Institute for Medical Informatics, Statistics and Epidemiology (IMISE), Leipzig University, Germany}
+%\address[A]{Institute for Medical Informatics, Statistics and Epidemiology, Leipzig University, Germany}
 \address[A]{IMISE, Leipzig University, Germany}
@@
 %Textbooks contain abstract knowledge about a domain.
-Textbooks about information management in hospitals describe the planning, monitoring and directing of a hospital's information system.
-SNIK consists of the manually transformed content of three such textbooks as Linked Open Data.
-The data model describes information management functions, roles executing these functions and the information used or updated by these functions.
+%Textbooks about information management in hospitals describe the planning, monitoring and directing of a hospital's information system.
+
+SNIK consists of the manually transformed content of three textbooks as Linked Open Data.
+SNIK Quiz...
+%The data model describes information management functions, roles executing these functions and the information used or updated by these functions.
+
 %SNIK provides applications that are used to teach students internationally.
@@
 %As the domain is large and highly relevant to students of Medical Informatics, modelling the knowledge is not only possible but also very useful.
-We publish the result over several interfaces that are useful for researchers, administrators or students, depending on their objectives and capabilities.
+%We publish the result over several interfaces that are useful for researchers, administrators or students, depending on their objectives and capabilities.
 \end{abstract}
@@
 \section{Introduction}
-A health information system (HIS) is the socio-technical subsystem of a care delivery organization~\citep{bb}.
-This includes its different application systems, computers, and network components as well as their users.
-A HIS processes data, information, and knowledge and its management involves planning, monitoring and directing those activities.
-Due to the complexity and the unique conditions in health care, HIS management is an challenging task.
-There are many different frameworks, textbooks and articles describing the scope of HIS management from the perspective of medical informatics.
+%A health information system (HIS) is the socio-technical subsystem of a care delivery organization~\citep{bb}.
+%This includes its different application systems, computers, and network components as well as their users.
+Medical informatics students, who are trained for executive positions in information management departments of healthcare institutions, such as hospitals, need a clear terminology of their domain.
+This terminology is offered by SNIK~\citep{semantischesnetz,sniktec}, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German), which integrates knowledge extracted from three textbooks~\citep{bb,ob,he} in the form of Linked Open Data.
+
+%A HIS processes data, information, and knowledge and its management involves planning, monitoring and directing those activities.
+%Due to the complexity and the unique conditions in health care, HIS management is an challenging task.
+%There are many different frameworks, textbooks and articles describing the scope of HIS management from the perspective of medical informatics.
 %However, the disciplines of business informatics and information systems (IS) provide an even broader view on information systems and their management.
 %A structured representation of the different perspectives leads to a holistic view on HIS management and helps help researchers and students connect their existing knowledge with further knowledge from other sources during research and learning.
-In order to integrate different knowledge sources and to provide the knowledge in a structured, machine-readable data format, we extracted knowledge about HIS management from several sources, see \cref{tab:source}.
 %three textbooks~\citep{bb,ob,he} and other sources, see \cref{tab:source}.
 %\section{Sources}\label{sec:sources}
-Three textbooks provide different views on the domain of Hospital Information Management:
-\citet{bb} presents a broad view on \enquote{typical architectures of health information systems and their systematic strategic management}.
-\citep{ob} concentrates on the \emph{tactical} management of information systems in general and on healthcare in particular.
-The focus of tactical management lies on the planning and operation of projects.
-\citet{he} explains information management beyond the scope of healthcare.
-Other sources are interviews and standards.
+%Three textbooks provide different views on the domain of Hospital Information Management:
+%\citet{bb} presents a broad view on \enquote{typical architectures of health information systems and their systematic strategic management}.
+%\citep{ob} concentrates on the \emph{tactical} management of information systems in general and on healthcare in particular.
+%The focus of tactical management lies on the planning and operation of projects.
+%\citet{he} explains information management beyond the scope of healthcare.
+%Other sources are interviews and standards.
 %CIOX is based on an interview with the CIO of the Universitätsklinikum Leipzig.
@@
 %
-The combination of this knowledge~\citep{semantischesnetz,domaene,approachtosupport} results in SNIK, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German).
-
-In order to encourage and enable other researchers, students and health informatics professionals to use available knowledge of HIS management, we introduce different user interfaces to the SNIK ontology and discuss their suitability for different use cases of the target audiences.
 %We discuss advantages and disadvantages of the interfaces and give recommendations on which ones present the best compromise for different use cases and target audiences.
 %We conclude with plans for future work on interlinking and visualization.
-
 %As the domain is large and highly relevant to students of Medical Informatics, modelling the knowledge is not only possible but also very useful.
-Medical informatics students, who are trained for executive positions in information management departments of healthcare institutions, such as hospitals, need a clear terminology of their domain.
-Due to different frameworks and textbooks dealing with information management in healthcare, modelling the knowledge unravels the links between the different views on information management.
-These are only implicitly known or not known at all by experts in the field.
+%Due to different frameworks and textbooks dealing with information management in healthcare, modelling the knowledge unravels the links between the different views on information management.
+%These are only implicitly known or not known at all by experts in the field.
 %
-We present a common data model for the domain and the result of applying the data model to three textbooks.
-We publish the result over several interfaces that are useful for researchers, practitioners and students, depending on their objectives and their Semantic Web skills.
 %
-Publishing textbook knowledge as Linked Data enables different ways of teaching.
+%Publishing textbook knowledge as Linked Data enables different ways of teaching.
 %The data is continously revised: users constantly report wrong or missing data in the visualization, which is then corrected, respectively revised, by the researchers of the project.
 
+\iffalse
 \begin{table}[tbh]
@@
 \end{table}
+\fi
 
 \section{Methods}
+
+\begin{figure*}
+\caption{The SNIK meta model}
+\label{fig:metamodel}
+\centering
+\includegraphics[width=0.7\textwidth]{img/metamodel9s.pdf}
+\end{figure*}
+
 \iffalse
@@
 \end{figure}
+\vspace{-3pt}
 \fi
-\vspace{-3pt}
-In order to specify, which information should be extracted from the books and to facilitate comparisons, we use a common data model.
-Because processed textbooks contain abstract knowledge instead of information about any specific hospital, all concepts are modelled as classes.
-We thus call our data model the \enquote{meta model} in accordance with the term's definition as a shared modelling language~\citep[p.~8]{ob}.
-The meta model (see \cref{fig:metamodel}) provides a common vocabulary for the domain of HIS management and thus defines, which superclasses and property can be used.
-SNIK version \snikversionlink{} comprises five subontologies that are built upon the meta model, see Table 1.
+
+SNIK describes information management functions, roles executing these functions and entity types, the information used or updated by these functions.
+Because the textbooks describe those concepts in the abstract sense, and not those of any concrete hospital, those entities are modelled as classes and individuals at the same time, using OWL 2 DL \emph{punning}.
+Thus, \emph{function}, \emph{role} and \emph{entity type} are metaclasses and the ontology of SNIK is called the \emph{meta model}, see \cref{fig:metamodel}.
+%In order to specify, which information should be extracted from the books and to facilitate comparisons, we use a common data model.
+%Because processed textbooks contain abstract knowledge instead of information about any specific hospital, all concepts are modelled as classes.
+%We thus call our data model the \enquote{meta model} in accordance with the term's definition as a shared modelling language~\citep[p.~8]{ob}.
+%The meta model (see \cref{fig:metamodel}) provides a common vocabulary for the domain of HIS management and thus defines, which superclasses and properties can be used.
+%SNIK version \snikversionlink{} comprises five subontologies that are built upon the meta model, see Table 1.
 %At the head of the class hierarchy (see \cref{fig:hierarchy}) is the \enquote{Top} class, which has exactly three disjunctive subclasses.
-At the head of the class hierarchy is the \enquote{Top} class, which has exactly three disjunctive subclasses.
+%At the head of the class hierarchy is the \enquote{Top} class, which has exactly three disjunctive subclasses.
 %
-Following the meta model, each class has to be a subclass of exactly one of them.
-The correct superclass of a new concept can be found by answering the question: Who (\enquote{Role}) does what (\enquote{Function}) and which information (\enquote{EntityType}) is needed?
+%Following the meta model, each class has to be a subclass of exactly one of them.
+%The correct superclass of a new concept can be found by answering the question: Who (\enquote{Role}) does what (\enquote{Function}) and which information (\enquote{EntityType}) is needed?
 If a concept is neither of them, it cannot be modeled using the meta model.
-As the subclass relation is transitive, a new class can be placed further down the hierarchy and it can still be inferred, whether it is a Function, Role or EntityType (see Figure 3).
-Besides the subclass relationship, two classes can be connected with relations provided by the meta model.
-The generic \enquote{is associated with} relationship carries little information.
-For example, a role and a function can be connected as \enquote{is involved in} \enquote{is responsible for} and \enquote{approves}.
-Relations that are neither of them can either be modeled by using the generic \enquote{is associated with} relation or by creating and using a new sub relation of \enquote{is associated with}
+%As the subclass relation is transitive, a new class can be placed further down the hierarchy and it can still be inferred, whether it is a Function, Role or EntityType (see Figure 3).
+%Besides the subclass relationship, two classes can be connected with relations provided by the meta model.
+%The generic \enquote{is associated with} relationship carries little information.
+%For example, a role and a function can be connected as \enquote{is involved in} \enquote{is responsible for} and \enquote{approves}.
+%Relations that are neither of them can either be modeled by using the generic \enquote{is associated with} relation or by creating and using a new sub relation of \enquote{is associated with}
 %
-As a concession to practicality, we express each geleral rule extracted from a textbook as a single triple using classes as subject and object, and a property of the meta model.
-For example, the rule "the CEO is involved in project reviews" is modelled as \enquote{\texttt{:Ceo} \aurl{meta}{isInvolvedIn} \texttt{:ProjectReview}.}, where \texttt{:Ceo} is a subclass of \aurl{meta}{Role} and \aurl{meta}{ProjectReview} is a subclass of \aurl{meta}{Function}.
-While an OWL restrictions using \aurl{owl}{someValuesFrom} and \aurl{owl}{allValuesFrom} would be technically correct, the knowledge is not expressed that specifically in the textbooks.
-Additionally, encoding each fact as a single triple facilitates tool support and prevents inconsistencies through non-atomic changes.
+%As a concession to practicality, we express each geleral rule extracted from a textbook as a single triple using classes as subject and object, and a property of the meta model.
+%For example, the rule "the CEO is involved in project reviews" is modelled as \enquote{\texttt{:Ceo} \aurl{meta}{isInvolvedIn} \texttt{:ProjectReview}.}, where \texttt{:Ceo} is a subclass of \aurl{meta}{Role} and \aurl{meta}{ProjectReview} is a subclass of \aurl{meta}{Function}.
+%While an OWL restrictions using \aurl{owl}{someValuesFrom} and \aurl{owl}{allValuesFrom} would be technically correct, the knowledge is not expressed that specifically in the textbooks.
+%Additionally, encoding each fact as a single triple facilitates tool support and prevents inconsistencies through non-atomic changes.
 
@@
 %-> use SHAPE/SHACL ?
-
-\begin{table}
-\caption{Access Options and Details.%
-%KBox usage example: \texttt{pip install kbox \&\& kbox -kb http://www.snik.eu/ontology -install -server}.%doesn't work right now
-}
-\label{tab:access}
-\begin{tabulary}{\columnwidth}{lL}
-\toprule
-URL						&\url{http://www.snik.eu/ontology}\\
-Version					&\snikversionlink{}\\
-%License					&CC BY-NC-SA 4.0\\
-License					&Attribution-NonCommercial-ShareAlike 4.0 International (\href{https://creativecommons.org/licenses/by-nc-sa/4.0/}{CC BY-NC-SA 4.0})\\
-Documentation	&\url{https://snikproject.github.io/ontology/}\\
-SPARQL Endpoint			&\url{http://www.snik.eu/sparql}\\
-LOD Cloud				&\url{https://lod-cloud.net/dataset/snik}\\
-%Datahub				&\url{https://datahub.io/imise/snik/v/1}\\%offline
-%Docker
-KBox~\cite{kbox} Graph	&\url{http://www.snik.eu/ontology}\\
-RDF Browser				&\url{http://www.snik.eu/ontology}\\
-Visualization			&\url{http://www.snik.eu/graph}\\
-Download				&\url{https://github.com/snikproject/ontology/archive/refs/tags/\snikversion.zip}\\
-Figshare				&\url{https://doi.org/10.6084/m9.figshare.17057084.v1}\\
-\bottomrule
-\end{tabulary}%The source code for the services is available at \url{https://github.com/imise}
-\end{table}
-
-\iffalse
-\begin{table}
-\caption{Statistics}
-\label{tab:statistics}
-\centering
-\begin{tabulary}{\columnwidth}{lS}
-\toprule
-\textbf{Category}					&\textbf{Count}\\
-\midrule
-Triples								&\sniktriples{}\\
-Classes								&\snikclasses{}\\
-Properties							&\snikproperties{}\\
-Internal Links						&360\\
-Links to HITO~\citep{hitometh2}  	&260\\
-Links to DBpedia~\citep{dbpedia} 	&\sniklinks{}\\
-%Subontologies	&\url{http://www.snik.eu/ontology/bb}\\
-%		&\url{http://www.snik.eu/ontology/ciox,}\\
-%		&\url{http://www.snik.eu/ontology/ob}\\
-%		&\url{http://www.snik.eu/ontology/he}\\
-%		&\url{http://www.snik.eu/ontology/it4it}\\
-\bottomrule
-\end{tabulary}%The source code for the services is available at \url{https://github.com/imise}
-\end{table}
-\fi
-
-\begin{figure*}
-\caption{The SNIK meta model}
-\label{fig:metamodel}
-\centering
-%\includegraphics[width=0.9\textwidth]{img/metamodel9s.pdf}%7 page version
-\includegraphics[width=0.7\textwidth]{img/metamodel9s.pdf}
-\end{figure*}
 
@@
 %\section{Lifecycle}\label{sec:application}
-\section{Results}\label{sec:application}
-\iffalse
-\begin{figure}[ht]
-\caption{The SNIK lifecycle}
-\label{fig:lifecycle}
-\includegraphics[width=0.5\columnwidth]{cycle.pdf}
-\end{figure}
-\fi
-When we add a new textbook to SNIK, we identify all concepts that fit the meta model and their relations to other concepts.
-Some project members are not proficient in RDF serializations, so we first fill out a meta model conforming spreadsheet template.
-We then use Tarql\footnote{\url{https://github.com/cygri/tarql}} with a mapping configuration file\footnote{\url{https://github.com/snikproject/snik-csv2rdf}} to convert that spreadsheet to RDF.
-Next, we refine the data in Protégé and upload it to a Virtuso SPARQL endpoint.
-%After the initial publish phase, we describe the repeating steps of the reuse phrase with the SNIK lifecycle in \Cref{fig:lifecycle}:
-\begin{enumerate}
-\item the Virtuoso SPARQL endpoint is used for querying and as data source for all our applications  
-\item an instance of the OntoWiki~\citep{ontowiki} is used for punctual changes and additions by non-Semantic-Web-experts, supplemented by SPARUL for changes involving large numbers of classes 
-\item LIMES~\citep{limes} generates interlink candidates between the different textbooks, which can be approved or rejected by the users.
-\aurl{skos}{closeMatch} is used instead of \aurl{owl}{equivalentClass} to emphasize that not only the names of concepts may differ between textbooks but there may be subtle differences in meaning as well.
-We do not intent to express different treatment, such as reasoning, due to the theoretical difference between \aurl{skos}{Concept} and \aurl{owl}{Class}.
-Links to HITO and DBpedia are created as well.%, see \cref{tab:statistics}.
-\item LodView\footnote{\url{https://lodview.it/}} lets the user browse among the classes of SNIK and view particular classes in detail
-\item SNIK Graph, detailed in \Cref{sec:snik-graph}, supplements the RDF browser by visualizing the relationships between classes.
-Users often discover incorrect modelling or possibilities for enhancement in SNIK graph.
-\item SNIK quality is a web application that uses SPARQL queries to find problems of varying degree, such as violations of domain, range or naming conventions, subclass cycles or missing definitions.
-\item Problems suggested by users and SNIK quality are verified, solved, and published on the SPARQL endpoint using SPARUL queries, at which point the circle begins anew.
-\end{enumerate}
+%Virtuoso SPARQL endpoint is used for querying and as data source for all our applications
 
@@
 %- SPARUL queries
-The technical environment of those services is described in~\citet{sniktec}.
-Selected applications are described in the following.
+%The technical environment of those services is described in~\citet{sniktec}.
 
 %\section{Applications}
-SNIK was initially intended~\citep{domaene} both as  software to support hospital CIOs and to support teaching.
-For the former goal, two applications were designed and implemented as prototypes~\citep{toreonto}: (1) the requirements engineering decision support system TOREOnto and (2) the knowledge exploration and navigation visualization CIONx.
-Due to internal regulations, efforts to integrate them into the informational infrastructure of the Uniklinikum Leipzig were stopped.
-Consequently, the other approaches are in the area of teaching support.
-
-%\subsection{Reuse of Standard Components}
-%To focus the software development effort on the components described below, we reuse standard components for common use cases:
-%Virtuoso Opensource 7 serves as the SPARQL endpoint.
-
-\subsection{SNIK Graph}\label{sec:snik-graph}
-%TODO: Search auto complete and suggestions
-%\paragraperformance
-
-Graph-based visualizations allow teachers to intuitively convey relationships between a selection of those concepts and also allow students to explore the domain on their own~\cite{ontologybased}.
-There are several existing graph-based Linked Data visualizations~\cite{linkeddatavisualization}, which visualize RDF resources (classes or instanes) as nodes and their relationships as edges, but they do not fit the requirements we have previously identified regarding performance, maintenance, usability, functionality and up-to-dateness~\cite{visualizationoflargeontologies}.
-SNIK Graph is a web-based client-side open source web application\footnote{Published at \url{http://www.snik.eu/graph}.} that offers Linked Data visualization for both classes and instances and according to the requirements of SNIK and its users.
-With the thousands of classes of SNIK, graph-based visualization suffers from overplotting.
-Thus, SNIK Graph offers several options to select and layout subgraphs, for example to show only a specific chapter of a book to prepare a lecture about a specific topic.
-Users can also iteratively explore SNIK starting at a single class using neighbourhood and path operations. %(see \cref{fig:snik-graph-circle-star})
-
-\paragraph{Design Goals and Requirements}
-Our main goal is to visualize knowledge extracted from text books to users that may not have any Semantic Web experience.
-The time and cognitive load of the users required to install, learn and operate the application should be as low as possible, so that they can focus on the data at hand and experience a benefit compared to only reading the textbook, such as when studying for an exam.
-A web application is easiest to set up for the user, as it is operating system agnostic and does not need to be installed.
-
-\paragraph{Implementation and Setup}
-As none of the investigated existing approaches support all our use cases and design goals and provide adequate performance on more than 4000 nodes, we implemented SNIK Graph as a web application using JavaScript (EcmaScript 2015) based on the graph visualization library Cytoscape.js~\cite{cytoscape}.
-SNIK Graph is freely available as open source software\footnote{Git repository at \url{https://github.com/snikproject/snik-graph}, licensed under CC BY-NC-SA 4.0}.
-As a client-side tool\footnotemark{}, SNIK Graph can be adapted to other knowledge bases and ontologies and published on any HTML web server by cloning the repository, installing the NPM dependencies and adapting the configuration file settings including the SPARQL endpoint URL.
-\footnotetext{Aside from a SPARQL endpoint.}%
-An installation containing SNIK is published at \url{https://www.snik.eu/graph}.
-
-%\paragraph{Features}
-\paragraph{Tabs and Containers}
-%\begin{figure}[h]
-%    \centering
-%    \includegraphics[width=0.5\columnwidth]{img/tabs.png}
-%    \caption{A session containing the default view and freely partitioned subgraphs containers.}
-%	\label{fig:tabs}
-%\end{figure}
-%\vspace{-3pt}
-
-%For this subgraphs of the ontology have to be extracted and organized.
-%A feature of SNIK Graph, tabs, allows multi-instance usage.
-%One major issue in working with SNIK Graph was the missing ability of organizing teaching scenarios.
-To organize teaching scenarios, independent SNIK Graph instances are loaded into containers\footnotemark{} that can be organized around the browser window.
-\footnotetext{Implemented using the JavaScript GoldenLayout framework (\url{https://golden-layout.com/}).} 
-These containers can either be accessed as tabs or freely partitioned.% see \cref{fig:tabs}.
-The content and of the all containers can be saved and loaded as a session to continue developing a teaching scenario at a later point.
-
-%This feature allows better workflows in preparation and presentation of teaching scenarios.
-%We use JSON formatted files for export and import.
-
-\paragraph{Compound Layout}
-%\begin{figure}[h]
-%    \centering
-%    \includegraphics[width=0.5\columnwidth]{img/combine.pdf}
-%    \caption{Grouping of equivalent concepts from different sources in SNIK Graph (color coded).}% using the \enquote{combine matches} option.}
-%	\label{fig:combine}
-%\end{figure}
-%\vspace{-3pt}
-
-%\noindent For citations of references, we prefer the use of square brackets and consecutive numbers, e.g.
-as shown by Author et al.
-[2], [3, pp.
-5–10], as mentioned earlier [1], [3], [9].
-The following bibliography provides the basic formats as a reference list with entries for journal articles [1], book chapter [2], as well as a URL [5].
-For further guidance please refer to the \href{https://ieeeauthorcenter.ieee.org/wp-content/uploads/IEEE-Reference-Guide.pdf}{IEEE-Reference-Guide}.
-
-Users can group interlinked nodes together and either display them side by side or on top of each other.%, see \cref{fig:combine}.
-As SNIK consists of more than 4000 classes, viewing large parts of it at once does not convey much information.%, see \cref{fig:much}.
-To alleviate this, SNIK offers multiple exploration methods described in the following.
-%\begin{figure}[h!]
-%    \centering
-%    \includegraphics[width=\linewidth]{much.pdf}
-%    \caption{Excerpt of SNIK using the default force directed layout.}\label{fig:much}
-%\end{figure}
-
-\paragraph{Role Use}
-
-%\todo{konrad an alle: was mache ich mit der sparql query?
-rausschmeissen, drin lassen oder umbauen?
-TP: drinlassen, aber direkt ans Bild hängen..
-Sowas wie aus der Query folgt dieses Bild.
-Ich finde sowas sollte schon rein ins Paper}
-%\begin{figure}[h]
-%\begin{lstlisting}
-%SELECT DISTINCT ?inner ?middle ?outer ?outerx {
-% <role> (rdfs:subClassOf|skos:closeMatch|^skos:closeMatch)* ?inner.
-% ?inner meta:subTopClass meta:Role.
-% OPTIONAL {
-%  ?inner ?p ?middle.
-%  ?middle meta:subTopClass meta:Function.
-%  OPTIONAL {
-%   ?middle ?q ?outer.
-%   ?outer meta:subTopClass meta:EntityType.
-%   OPTIONAL {?outer (skos:closeMatch|^skos:closeMatch|^rdfs:subClassOf)+ ?outerx.}
-%}}}
-%\end{lstlisting}
-%\caption{SPARQL query to generate the \emph{class use} visualization.}
-%\label{lst:roleuse}
-%\end{figure}
-%\vspace{-3pt}
-
-A frequent question is, what a given role does and which information is needed for those functions represented by the entity types connected to those functions.
-This question is visually answered by the \enquote{role use} feature, which arranges roles, functions and entity types in concentric circles.%, see \Cref{fig:classuse}:
-
-\begin{enumerate}%[align=left, labelwidth=1ex]
-\item The first ring consists of the given role and all roles that are connected by a path of \aurl{rdfs}{subClassOf} and \aurl{skos}{closeMatch} edges.
-\item The function ring consists of all functions adjacent to roles in the inner ring.
-\item The third ring consists of all entity types adjacent to functions in the middle ring.
-\item The final ring consists of all entity types that are connected to entity types from the third ring by a path of \aurl{skos}{closeMatch} and reverse \aurl{rdfs}{subClassOf} edges.
-\end{enumerate}
-
-\iffalse
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.5\columnwidth]{img/class-use-project-manager.pdf}
-    \caption{\emph{Role Use} of \aurl{bb}{ProjectManager}. Outermost layer ommitted due to space limitiations.}\label{fig:classuse}
-\end{figure}
-\fi
-
-In SNIK Graph, this operation is called \emph{class use} and allows users to also put a function or entity type in the center.
-
-%\section{Approach}
-%Due to its large size, visualizing all of SNIK at once  overplotting.
-%At the core of SNIK Graph is incremental exploration, which starts at one or more nodes and gradually expands them to related ones.
-
-\paragraph{Search}
-\iffalse
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.5\columnwidth]{img/search.pdf}
-    \caption{Visual search results for \enquote{Chief Information Officer} in a part of SNIK.}\label{fig:search}
-\end{figure}
-\fi
-Due to the large amount of resources, exploration often begins with a search.
-The search index is populated from the target SPARQL endpoint and is implemented using the Fuse.js library that is based on the Baeza-Yates--Gonnet algorithm~\cite{textsearching}.
-Fuse.js\footnotemark{} is a light-weight, purely client side JavaScript library and presents an alternative to backend-driven indexes like Lucene.%
-\footnotetext{\url{https://fusejs.io/}, \url{https://github.com/krisk/Fuse}}
-This enables fast fuzzy search on any dataset loaded via SPARQL endpoint but requires waiting for index initialization on the first search of each user session.
-Search results presented to the user are color coded in three categories: visible (green), invisible (yellow)%
-\footnote{The resource is either \emph{filtered} or \emph{hidden}.}% \cref{sec:visibility}
- and not loaded (red)%
-\footnote{The resource is included in the search index built from the SPARQL endpoint but either deleted in the graph or not loaded in the first place, such as due to configuration.}.
-Each search entry of a class contains the label values of \aurl{rdfs}{label} (long form) and \aurl{skos}{altLabel} (short and alternative forms) with a weight of 0.7.
-Textbook definitions are included with a weight of 0.3.
-Labels of resources connected via \aurl{skos}{closeMatch} interlinks in either direction are included as well, because those resources are defined as semantically close, so we assume their labels are synonyms.
-Fuzzy matching is enabled with a threshold of \num{0.25}.
-The resource \aurl{bb}{ChiefInformationOfficer} can thus be found by entering either \enquote{CIO} (alternative label), \enquote{vice president} (part of the definition) or \enquote{Leiter des Rechenzentrums} (German alternative label of \aurl{ob}{ChiefInformationOfficer}).
-%Search results are then selected in the graph, see~\cref{fig:search}, which allows further exploration by chaining the path and neighbourhood operations described in following.
-Search results are then selected in the graph, which allows further exploration by chaining the path and neighbourhood operations described in following.
-
-\paragraph{Shortest Paths}
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\columnwidth]{img/path.png}
-    \caption{The shortest non-trivial path between \aurl{bb}{PatientIdentificationNumber} and \aurl{bb}{QualityOfData} in SNIK Graph.}
-	\label{fig:path}
-\end{figure}
-\vspace{-3pt}
-The shortest path is the most basic method of visualizing the relationship between two resources, see \cref{fig:path}.
-We treat SNIK as an undirected graph for the calculation because the direction is often arbitrary.
-Roles can be nested using \aurl{meta}{roleComponent} (\emph{has part}), which could have just as well been modelled as a \emph{part of} relation with reversed subject and object.
-Another reason is that a set of resources that forms a graph connected by some symmetric property implies a complete subgraph, which negatively effects speed, memory and visibility of SNIK Graph.
-For this reason, \aurl{skos}{closeMatch} and other symmetrical properties are often only sparsely modelled, and the other triples are implied.
-As SPARQL endpoints, such as Virtuoso SPARQL employed by SNIK, do not perform reasoning to infer such implied triples such as those generated by symmetric properties, this would prevent the shortest path from including resources where only the opposite direction is explicitly modelled.%
-We also do not explicitly store triples inferred by transitive properties such as \aurl{rdfs}{subClassOf}, which prevents paths involving such properties from being squashed, such as the path between \emph{Project Manager} and \emph{HIS Stakeholder} in \cref{fig:hierarchy}.
-
-%We ignore the direction of edges because, that is, which resource is the subject and the object 
-The largest problem of shortest paths is however, that they are not necessarily informative to the user. 
-For example, any two roles, such as \aurl{bb}{ChiefInformationOfficer} and \aurl{bb}{ChiefExecutiveOfficer}, are connected by a path of length 2 as they are all subclasses of \aurl{meta}{Role}, which the user already knows given the triangular shapes representing roles in SNIK Graph.
-To exclude such trivial paths,  the meta model is not shown by default.
-Furthermore, filtering, such as by knowledge source or type of relation or resource, allows further tuning of the resources that are shown and eliglible for paths.
-%Further research into informative paths has been done in two bachelor theses~\cite{}.
-An general approach to solve the problem of informative shortest paths is \emph{weighted shortest paths for RDF graphs (WiSP)}~\cite{wisp}.
-Future work includes adopting this approach to SNIK Graph and evaluating it with its users.
-Another approach is to show all short paths under a given length and let the user remove unwanted ones manually, as employed by \emph{Relfinder}~\cite{relfinder}.
-
-\paragraph{Neighbourhood Operations}
-Exploration using neighbours, that is the successive uncovering of nodes adjacent to a starting node given by a user, is a common feature of tools such as LodLive~\cite{lodlive} and VizLOD~\cite{vizlod}.%
-\paragraph{Star}
-The directed and undirected star operations show nodes connected to selected nodes via all paths that contain at most one property other than \aurl{skos}{closeMatch}.
-The \emph{circle star} also rearranges the nodes using the force-directed layout locally on the currently visible subgraph.
-%\Cref{fig:star} shows a mind map of strategic information management, created by an undirected star, which can be used by a teacher to prepare a lecture about the topic.
-\iffalse
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\columnwidth]{img/strategic-im.pdf}
-    \caption{Undirected \emph{Star} of \aurl{bb}{StrategicInformationManagement}.}
-	\label{fig:star}
-\end{figure}
-\vspace{-3pt}
-\fi
-
-%\todo{Konrad: Nimm den Star um "ob/ProjectManagement".} % ist inkludiert in \cref{fig:tabs}, refer?
-The concept \emph{project management} both exists in a textbook~\cite{ob} and in the knowledge about information management in a German university hospital.
-Grouping both concepts with the help of the star function the user detects that the terms have a different meaning according to their neighboorhood terms.
-Whereas the text book deals with the management of single projects, in the CIO's world \enquote{project management} means managing multiple projects, as shown in the lower right container in \cref{fig:tabs}.
-As all neighbourhood and path operations automatically select all resulting nodes, repeated application of a star without clearing the selection results in an incremental uncovering of the whole connected graph around the starting node.
-
-%\subsection{Mixed Operations}
-\paragraph{Spiderworm}
-A \emph{spiderworm} is a path from node \emph{A} to node \emph{B} combined with a \emph{star} of \emph{B}.
-%\Cref{fig:spiderworm} shows how we use a spiderworm to teach a student how the new concept \enquote{quality of data} is connected the the already introduced concept \enquote{patient identification number.}
-
-\iffalse
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=\columnwidth]{img/spiderworm.pdf}
-    \caption{\emph{Spiderworm} from \aurl{bb}{PatientIdentificationNumber} to \aurl{bb}{QualityOfData}~\cite{snikgraphposter}.}
-	\label{fig:spiderworm}
-\end{figure}
-\vspace{-3pt}
-\fi
-
-\iffalse
-Due to the limitations of JavaScript, such as a single thread, and the used library library Cytoscape.js which is the base of the graph visualization, there are performance problems on large graphs.
-Right now we have 2641 nodes and 4904 edges and the performance is ok in Chrome but not amazing.
-With 10000 nodes it would probably be too slow on most machines.
-
-\todo{reference franziskas poster in both figures}
-\todo{get a higher quality version of overview graphic}
-\todo{reference \citet{visualizationoflargeontologies}}
-
-\begin{figure}
-\caption{SNIK Graph Overview. Source: \protect\citet{snikgraphposter}.}
-\label{fig:snik-graph-overview}
-\includegraphics[width=\columnwidth]{img/snik-graph.png}
-\end{figure}
-
-\begin{figure}
-\caption{A teacher prepares a lecture on strategic HIS directing using the \enquote{Circle Star} function of SNIK Graph. Source: \protect\citet{snikgraph}.}
-\label{fig:snik-graph-circle-star}
-\includegraphics[width=\columnwidth]{img/snik-graph-circle-star.png}
-\end{figure}
-
-\begin{figure}
-\caption{Students learn new concepts about HIS quality by linking them to concepts already learned.
-A teacher asks a student to find out how the new concept \enquote{Quality of Data} is linked to the \enquote{Patient Identification Number}.
-The student connects the two concepts by using the “spiderworm“ visualization and learns that a patient identification number is associated with object identity.
-Object identity is a subclass of integrity of data.
-Besides integrity of data, there are also 13 other criteria for quality of data.}
-\label{fig:snik-graph-spiderworm}
-\includegraphics[width=\columnwidth]{img/snik-graph-spiderworm.png}
-\end{figure}
-\todo{FJ: In diesem Abschnitt fehlt die Diskussion, wie es durch die verschiedenen Zielgruppen bei Use Cases genutzt werden kann.}
-SNIK Graph is a web application\footnote{\url{http://www.snik.eu/graph}} that transforms the classes and triples of SNIK to nodes and edges of a graph.
-The graph is visualized using the Cytoscape.js~\citep{cytoscape} library with the force-directed Euler layout, see~\cref{fig:snik-graph-overview}.
-With several thousand classes, specific parts of SNIK can be hard to discern.
-Thus, there are several options to view subgraphs of SNIK, for example to show only a specific chapter of a book to prepare a lecture about a specific topic.
-Users can also search for and restrict the view to a single class and then show the neighbourhood of that class (see \cref{fig:snik-graph-circle-star}) and subsequently the neighbourhood of selected neighbours of the previous step.
-SNIK Graph can also calculate the shortest path to between classes.
-We also tried to automatically find the most interesting path but this was not successfull as it is subjective and the depends on the goal of the user.
-Path and neighourhood operations are joined in the \enquote{spider worm} (see \cref{fig:snik-graph-spiderworm}), which consists of the shortest path between a start node and an end node together with the end node’s neighbourhood, illustrate the context of a concept.
-\fi
-\subsection{Quiz}
 
@@
 Clover Quiz~\citep{cloverquiz} shows that ontologies can be used to automatically generate multiple-choice questions over DBpedia~\citep{dbpedia}.
-We apply this approach to SNIK and generate 1231 English questions using the templates described in \cref{tab:templates}.
-It was used by students from the Universities of Amsterdam, Heidelberg and Leipzig during the international Frank van Swieten lectures in 2019.
 To provide difficult wrong answers, the so called \emph{distractors}, we use direct neighbours of the classes that represent the correct answers.
@@
 It is unclear, whether the CIO is also responsible for operational information management.
-An ongoing bachelor's thesis~\citep{snikquiz} is investigating different strategies of generating complex questions, answers and distrators.
+%An ongoing bachelor's thesis~\citep{snikquiz} is investigating different strategies of generating complex questions, answers and distrators.
+
+\subsection{Question Types}
+%\paragraph{}
+
+\paragraph{intertwined}
+
+\begin{figure}
+\includegraphics[width=0.7\textwidth]{img/intertwined_cml.png} 
+\caption{Left to right: correct answer, difficult and easy distractor for the \emph{intertwined} question type.}
+\label{fig:intertwined}
+\end{figure}
+
+Intertwined focuses on the connectivity of the knowledge graph.
+In each answer, three related concepts are given.
+The correct answer is the one where all concepts are directly connected with each other, that is, they form a triangle in the graph, see \cref{fig:intertwined}.
+The distractors should then not form this triangle.
+In the difficult variation of the distractor, one of the three concept pairs is directly connected, one has a shortest path of length 2 and the third one has neither.
+In this easier variation, two concepts pairs are connected with a shortest path of length 2 instead.
+
+\paragraph{closeMatch}
+\begin{table}
+\begin{tabular}{p{0.8\textwidth}  p{0.1\textwidth}}
+\toprule
+question	& correct\\
+\midrule
+In the Tactical Information Management, the Project Execution uses the Resource, while in the Strategic Information Management, the Project Execution uses the Resource.											&True \\
+In the Strategic Information Management, the Project Management Board is responsible for  the Project Monitoring, while in the Tactical Information Management, the Project Steering Committee uses the Project.	&False \\
+\bottomrule
+\end{tabular}
+\caption{Examples of generated \emph{closeMatch}-questions}
+\label{tab:closematch}
+\end{table}
+
+This question type tests the connection between the textbooks \cite{ob} (tactical information management) and \cite{bb} (strategic information management) based on the existing \aurl{skos}{closeMatch} interlinks.
+The aim is to help students transfer knowledge in one textbook to the other.
+The questions have the form: In [Tactical / Strategic IM], the [subject] [predicate] the [object], while in [Strategic / Tactical IM], the [subject] [predicate] the [object] with 2 possible answers: true and false.
+Examplarly questions are given in \cref{tab:closematch}.
+
+%Dies beinhaltet dann Begriffe, welche in beiden Büchern auftreten.
+Ausgehend davon sollen dann ebenfalls in beiden Büchern auftretende Begriffe gesucht werden, welche verschiedene Grundklassen aufweisen: im einen Buch als Aufgabe und im anderen als Subjekt.
+Damit sollen dann Sätze gebildet werden, welche eine Gegenüberstellung in den Büchern erlauben, um so Brücken zu bauen und das Wissen des einen Buches effektiv auf das andere transferieren zu können.
+Ein Fragesatz könnte dann lauten: \enquote{In \textit{[Tactical / Strategic IM]}, the \textit{[Subjekt] [Prädikat]} the \textit{[Objekt]}, while in \textit{[Strategic / Tactical IM]}, the \textit{[Subjekt] [Prädikat]} the \textit{[Objekt]}} mit 2 Antwortmöglichkeiten, wahr und falsch.
+
+\paragraph{Occurence}
+Hier handelt es sich um eine Spezialisierung von close-by, bei der ent-
+schieden werden soll, ob ein Begriff in keinem Buch, in Ammenwerth
+u. a., 2014, in Winter u. a., 2011 oder in beiden Büchern auftritt. Eine
+mögliche Frage ist hier: \enquote{In which books does the term \emph{Chief Information Officer} occur?}.
+Auf diese Frage kann dann noch die Definition des Begriffes folgen, um noch mehr Kontext zu liefern.
+
+\begin{table}
+\begin{tabular}{p{0.5\textwidth}  p{0.3\textwidth}}
+\toprule
+Frage  & korrekt \\
+\midrule
+In which contexts does the term \enquote{Automated Observation} occur? & Tactical Information Management \\
+In which contexts does the term \enquote{Health Insurance Company} occur? & Strategic Information Management \\
+In which contexts does the term \enquote{System Analysis} occur? & Both Contexts \\
+In which contexts does the term \enquote{System Administrator} occur? & Neither\\
+\bottomrule
+\end{tabular}
+\caption[occurence-Ergebnisbeispiele]{Beispiele für generierte Fragen der \textit{occurence}-Strategie}
+\label{tabelle:bsp_occurence}
+\end{table}
+
+\section{Results}\label{sec:application}
+We apply this approach to SNIK and generate 1231 English questions using the templates described in \cref{tab:templates}.
+It was used by students from the Universities of Amsterdam, Heidelberg and Leipzig during the international Frank van Swieten lectures in 2019.
 
 %The technical environment of SNIK is characterized in previous work [11].
-\iffalse
-\section{Discussion}
-SNIK is based on the meta model from Figure \todo{reference}.
-Thus, SNIK can be considered as an archetype for ontologies describing for a given domain (here \enquote{HIS management}), what functions a certain role has to carry out and what information a person of this role needs and what information she/he provides while carrying out the functions.
-This may be used for domains like the medical department of a hospital as well.\todo{FJ: Ich glaube, das müsste man ausführlicher erklären, das würde ich eher ans Ende der Diskussion nehmen.}
-This would be valuable knowledge, for instance as basis for systems analysis projects [1].
-Different user groups can use the interfaces of SNIK that are most suitable to them.
-For example, the IT management of a hospital can use SNIK as a vocabulary to integrate data from different formats that result from application components from different vendors.
-Using the SPARQL query language, experts can also integrate applications with the SNIK ontology.
- Experts can also use the SPARQL query editor to answer specific questions, see Table 2.
-Students and teachers, on the other hand, benefit more from SNIK Graph, which intuitively presents knowledge without the need for technical skills.
-
-Figure 4– A \enquote{Spider Worm} between the classes \enquote{Change Management} and \enquote{HIS Quality} in the SNIK Graph visualization web application.
-The content of SNIK is taken from textbooks which are protected by property rights of publishing houses or taken from an interview with a CIO holding property rights as well.
-Making the SNIK ontology openly available may therefore conflict with these property rights.
-In case of SNIK, we appreciate the approvals of the publishing houses Springer, de Gruyter and Schattauer and of the CIO.
-In general, missing approvals may prevent the creation of such open ontologies and thus the provision of textbook content as open knowledge.
-The approvals for SNIK were possible because for each class and relation being part of SNIK there is a clear indication of its source (see code \enquote{bb} and chapter information in Table 1).
-In future work we will use these indications for establishing hyperlinks into the sources such that users of SNIK may use SNIK as a smart index for retrieving more detailed knowledge and supporting illustrations from the original textbooks – if digitally available.
-We are convinced that this way an ontology like SNIK is not a competitor of textbooks but a smart guide for using a textbook.
-Thus, it might be attractive for publishing houses.
-\fi
-\todo{verweis auf existierende papers}
-\todo{benutzung durch braunschweig}
-\todo{benutzung durch amsterdam}
 
 \section{Discussion}
+
 We showed that knowledge on the management of information systems in medicine and health care is made publicly available using open standards over several interfaces with different compromises between expressivity and accessibility.
@@
 \section{Conclusions}
-In the future, we plan to extend the text-based interlinks with semantic-similarity-based ontology matching [5] to find more correspondencies between the textbooks.
-This will enable users to switch to another textbook when a certain concept is explained there in more detail.
-The SNIK toolset, SNIK Graph in particular, is under permanent improvement.
-For instance, we are considering alternatives to the shortest path that are the most preferred by different audiences and whether those paths can be (semi-)automatically generated using suitable node or edge weight functions.
-We are also investigating candidates for additional subontologies and as interlink targets.
-Past efforts of automatic textbook extractions were unsuccessfull but recent advances in natural language processing, such as GPT-3~\citep{gpt3}, motivate further investigation.
 

### 2022-01-21 quiz.tex
More quiz.

@@
 SNIK Quiz...
+More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract 
+More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract 
+More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract 
+More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract 
 %The data model describes information management functions, roles executing these functions and the information used or updated by these functions.
@@
 This terminology is offered by SNIK~\citep{semantischesnetz,sniktec}, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German), which integrates knowledge extracted from three textbooks~\citep{bb,ob,he} in the form of Linked Open Data.
-
+More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction 
+More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction 
+More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction 
 %A HIS processes data, information, and knowledge and its management involves planning, monitoring and directing those activities.
@@
 It is unclear, whether the CIO is also responsible for operational information management.
-%An ongoing bachelor's thesis~\citep{snikquiz} is investigating different strategies of generating complex questions, answers and distrators.
+
+% TODO: reformulate now that its not future work anymore but part of the paper
+An ongoing bachelor's thesis~\citep{snikquiz} is investigating different strategies of generating complex questions, answers and distrators.
 
 \subsection{Question Types}
-%\paragraph{}
+
+TODO: Add the old question types, but only the b-versions (so no suffix is required in this paper)
+
+\paragraph{definition}
+describe definition-b...
+
+\paragraph{label-definition}
+describe label-definition...
+
+\paragraph{contains}
+describe contains...
 
@@
 \paragraph{closeMatch}
-\begin{table}
-\begin{tabular}{p{0.8\textwidth}  p{0.1\textwidth}}
+\begin{table}[h]
+\begin{tabulary}{\textwidth}{Ll}
 \toprule
@@
 \bottomrule
-\end{tabular}
+\end{tabulary}
 \caption{Examples of generated \emph{closeMatch}-questions}
@@
 The aim is to help students transfer knowledge in one textbook to the other.
-The questions have the form: In [Tactical / Strategic IM], the [subject] [predicate] the [object], while in [Strategic / Tactical IM], the [subject] [predicate] the [object] with 2 possible answers: true and false.
+The questions have the form: \emph{\enquote{In [Tactical / Strategic IM], the [subject] [predicate] the [object], while in [Strategic / Tactical IM], the [subject] [predicate] the [object].}} with two possible answers: true and false.
 Examplarly questions are given in \cref{tab:closematch}.
@@
 
-\paragraph{Occurence}
-Hier handelt es sich um eine Spezialisierung von close-by, bei der ent-
-schieden werden soll, ob ein Begriff in keinem Buch, in Ammenwerth
-u. a., 2014, in Winter u. a., 2011 oder in beiden Büchern auftritt. Eine
-mögliche Frage ist hier: \enquote{In which books does the term \emph{Chief Information Officer} occur?}.
-Auf diese Frage kann dann noch die Definition des Begriffes folgen, um noch mehr Kontext zu liefern.
-
-\begin{table}
-\begin{tabular}{p{0.5\textwidth}  p{0.3\textwidth}}
+\paragraph{occurence}
+\begin{table}[h]
+\begin{tabulary}{\textwidth}{LL}
 \toprule
-Frage  & korrekt \\
+question	&correct\\
 \midrule
-In which contexts does the term \enquote{Automated Observation} occur? & Tactical Information Management \\
-In which contexts does the term \enquote{Health Insurance Company} occur? & Strategic Information Management \\
-In which contexts does the term \enquote{System Analysis} occur? & Both Contexts \\
-In which contexts does the term \enquote{System Administrator} occur? & Neither\\
+In which contexts does the term \enquote{Automated Observation} occur?		&Tactical Information Management \\
+In which contexts does the term \enquote{Health Insurance Company} occur?	&Strategic Information Management \\
+In which contexts does the term \enquote{System Analysis} occur?			&Both Contexts \\
+In which contexts does the term \enquote{System Administrator} occur?		&Neither\\
 \bottomrule
-\end{tabular}
-\caption[occurence-Ergebnisbeispiele]{Beispiele für generierte Fragen der \textit{occurence}-Strategie}
-\label{tabelle:bsp_occurence}
+\end{tabulary}
+\caption{Exemplary \emph{occurrence}-questions and correct answers.}
+\label{tab:occurence}
 \end{table}
 
+The \emph{occurence} question type is a specialization of \emph{closeMatch}, where the student has to choose whether a term occurs in~\cite{ob},~\cite{bb}, in both textbooks or neither of them.
+Exemplary questions are given in \cref{tab:occurence}.
+%mögliche Frage ist hier: \enquote{In which books does the term \emph{Chief Information Officer} occur?}.
+The context of the questions can be augmented with the definitions of the concepts.
+
 \section{Results}\label{sec:application}
-We apply this approach to SNIK and generate 1231 English questions using the templates described in \cref{tab:templates}.
+The question types definition-a and contains-a were used to generate 1231 English questions, see \cref{tab:templates}.
 It was used by students from the Universities of Amsterdam, Heidelberg and Leipzig during the international Frank van Swieten lectures in 2019.
@@
 
+\subsection{Evaluation}
+\begin{table}[h]
+\begin{tabularx}{\textwidth}{Xccc}
+\toprule
+Type				&rating interview 1	&rating interview 2	&rating interview 3\\
+\midrule
+label-definition	&3				&4				&3\\
+definition-b		&4				&6				&3\\
+contains-b			&5--6			&7				&6\\
+\midrule
+intertwined			&7				&7--8			&8\\
+closeMatch			&8				&7--8			&8\\
+occurence			&7				&5				&7\\
+\bottomrule
+\end{tabularx}
+\caption{Rating given by interviewees based on the complexity of the questions on a scale between 1 and 10 with 10 representing maximum complexity.}
+\label{tabelle:eval_komplex}
+\end{table}
+
 \section{Discussion}
 
-We showed that knowledge on the management of information systems in medicine and health care is made publicly available using open standards over several interfaces with different compromises between expressivity and accessibility.
-It can be combined with other knowledge in biomedical and health informatics and in other disciplines.
+%We showed that knowledge on the management of information systems in medicine and health care is made publicly available using open standards over several interfaces with different compromises between expressivity and accessibility.
+%It can be combined with other knowledge in biomedical and health informatics and in other disciplines.
 

### 2022-01-22 quiz.tex
More quiz.

@@
 \newcommand{\snikversionlink}{\href{https://github.com/snikproject/ontology/releases/tag/\snikversion}{\snikversion}}
-\newcommand{\sniktriples}{81499}
-% SELECT (COUNT(DISTINCT *) AS ?count) {?s a owl:Class. FILTER(STRSTARTS(STR(?s),"http://www.snik.eu/ontology/"))}
-\newcommand{\snikclasses}{4107}
+\newcommand{\sniktriples}{\num{81499}}
+% SELECT (COUNT(DISTINCT ?s) AS ?count) {?s a owl:Class. FILTER(STRSTARTS(STR(?s),"http://www.snik.eu/ontology/"))}
+\newcommand{\snikclasses}{\num{4719}}
+% SELECT (COUNT(DISTINCT ?s) AS ?roles) {?s meta:subTopClass meta:Role.}
+\newcommand{\snikroles}{260}
+% SELECT (COUNT(DISTINCT ?s) AS ?functions) {?s meta:subTopClass meta:Function.}
+\newcommand{\snikfunctions}{\num{1310}}
+% SELECT (COUNT(DISTINCT ?s) AS ?entiytypes) {?s meta:subTopClass meta:EntityTypes.}
+\newcommand{\snikentitytypes}{\num{3182}}
 % SELECT (COUNT(DISTINCT ?s) AS ?count) {{?s a rdf:Property.} UNION {?s a owl:ObjectProperty.} UNION {?s a owl:DataTypeProperty.} FILTER(STRSTARTS(STR(?s),"http://www.snik.eu/ontology/"))}
@@
 %Textbooks about information management in hospitals describe the planning, monitoring and directing of a hospital's information system.
-
-SNIK consists of the manually transformed content of three textbooks as Linked Open Data.
-SNIK Quiz...
+SNIK is a knowledge base about the management of health information systems generated by extracting Linked Data from textbooks and other sources.
+%SNIK consists of the manually transformed content of three textbooks as Linked Open Data.
+SNIK Quiz demonstrates 
 More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract 
@@
 \begin{keyword}
-linked open data \sep information management
+linked open data \sep information management \sep multiple choice quiz
 \end{keyword}
@@
 Medical informatics students, who are trained for executive positions in information management departments of healthcare institutions, such as hospitals, need a clear terminology of their domain.
-This terminology is offered by SNIK~\citep{semantischesnetz,sniktec}, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German), which integrates knowledge extracted from three textbooks~\citep{bb,ob,he} in the form of Linked Open Data.
-More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction 
-More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction 
-More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction More introduction 
+This terminology is offered by SNIK~\citep{semantischesnetz,sniktec}, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German), which integrates knowledge extracted from three textbooks~\citep{bb,ob,he} and other sources such as interviews in the form of Linked Open Data.
+SNIK describes information management functions, roles executing these functions and entity types, the information used or updated by these functions.
+Because the textbooks describe those concepts in the abstract sense, and not those of any concrete hospital, those entities are modelled as classes and individuals at the same time, using OWL 2 DL punning.
+Thus, \emph{function}, \emph{role} and \emph{entity type} are metaclasses and the ontology of SNIK is a meta model as depicted in \cref{fig:metamodel}.
+SNIK version \snikversionlink{} contains \sniktriples{} triples describing \snikproperties{} properties and \snikclasses{} classes, of which there are \snikroles{} roles, \snikfunctions{} functions and \snikentitytypes{} entity types.
 %A HIS processes data, information, and knowledge and its management involves planning, monitoring and directing those activities.
@@
 \centering
-\includegraphics[width=0.7\textwidth]{img/metamodel9s.pdf}
+\includegraphics[width=\textwidth,trim={0 0 0 135mm},clip]{img/metamodel9s.pdf}
 \end{figure*}
@@
 
-SNIK describes information management functions, roles executing these functions and entity types, the information used or updated by these functions.
-Because the textbooks describe those concepts in the abstract sense, and not those of any concrete hospital, those entities are modelled as classes and individuals at the same time, using OWL 2 DL \emph{punning}.
-Thus, \emph{function}, \emph{role} and \emph{entity type} are metaclasses and the ontology of SNIK is called the \emph{meta model}, see \cref{fig:metamodel}.
 %In order to specify, which information should be extracted from the books and to facilitate comparisons, we use a common data model.
@@
 
+\begin{table}
+\begin{tabulary}{\textwidth}{LLLL}
+\toprule
+Question			&correct answer	&distractor 1	&distractor 2\\ 
+\midrule
+In the context of Strategic Information Management, which one of the following triples belongs together the most?
+&Chief Information Officer -- Department of Information management -- Information Management Staff & Chief Information Officer -- Strategic Gap -- Corporate Strategy & Chief Information Officer -- Ticket Evaluation -- Project Monitoring \\
+\midrule
+In the context of Strategic Information Management, which one of the following triples belongs together the most?
+& Information Management -- IT Service Management -- Operational Information Management & Information Management -- Tactical Management of Information Systems -- Information System & Information Management -- methodology of Information Management -- Operational Management of Information Systems\\
+\midrule
+In the context of CIOX, which one of the following triples belongs together the most & Chief Information Officer -- Jour Fixe Execution -- Project Monitoring & Chief Information Officer -- Project Risk -- Strategic HIS Planning & Chief Information Officer -- IT General Ledger Account Planning -- Project Portfolio \\
+\midrule
+In the context of CIOX, which one of the following triples belongs together the most & Chief Information Officer -- Project Presentation -- Project Waitlist & Chief Information Officer -- Strategic Information Management Planning -- Project Manager & Chief Information Officer -- Project Priorisation -- Project Web Site \\
+\bottomrule
+\end{tabulary}
+\caption{Examples of generated \emph{intertwined}-questions and answers.}
+\label{tab:intertwined}
+\end{table}
+
 Intertwined focuses on the connectivity of the knowledge graph.
@@
 In this easier variation, two concepts pairs are connected with a shortest path of length 2 instead.
+Examplse are given in \cref{tab:intertwined}.
 
@@
 \end{tabulary}
-\caption{Examples of generated \emph{closeMatch}-questions}
+\caption{Examples of generated \emph{closeMatch}-questions.}
 \label{tab:closematch}
@@
 
-This question type tests the connection between the textbooks \cite{ob} (tactical information management) and \cite{bb} (strategic information management) based on the existing \aurl{skos}{closeMatch} interlinks.
-The aim is to help students transfer knowledge in one textbook to the other.
+This question type tests the connection between a textbook about tactical information management~\cite{ob} and one about strategic information management~\cite{bb} based on the existing \aurl{skos}{closeMatch} interlinks.
+The aim is to help students transfer knowledge about one textbook to the other.
 The questions have the form: \emph{\enquote{In [Tactical / Strategic IM], the [subject] [predicate] the [object], while in [Strategic / Tactical IM], the [subject] [predicate] the [object].}} with two possible answers: true and false.
@@
 \subsection{Evaluation}
+\iffalse
 \begin{table}[h]
@@
 \end{table}
+\fi
+
+% transposed
+\begin{table}[h]
+\begin{tabularx}{\textwidth}{Xcccccc}
+\toprule
+			&label-definition	&definition	&contains	&intertwined	&closeMatch	&occurence\\
+interview 1	&3					&4			&5--6		&7				&8			&7\\
+interview 2	&4					&6			&7			&7--8			&7--8		&5\\
+interview 3	&3					&3			&6			&8				&8			&7\\
+\bottomrule
+\end{tabularx}
+\caption{Rating given by interviewees based on the complexity of the questions on a scale between 1 and 10 with 10 representing maximum complexity.}
+\label{tabelle:eval_komplex}
+\end{table}
 
 \section{Discussion}
+systematic review of multiple choice question generation from ontologies
+complex questions, 
+\cite{ontologybasedmultiplechoice}
+
+performance 
+static html and javascript
+client side
 

### 2022-01-24 quiz.tex
label-definition and definition

@@
 
+\paragraph{label-definition}
+This gives the user a definition and asks for  the correct term.
+The terms to be considered are chosen from terms, which are positioned within a distance of two vertices in the graph.
+So, relative semantic similarity can be assumed.
+This makes the label-definition-approach especially worthwhile when basic knowledge of terms of information management is required.
+
+% set this as second strat to gain constistency to table 5 
 \paragraph{definition}
-describe definition-b...
-
-\paragraph{label-definition}
-describe label-definition...
+The definition approach focuses on the question of which definition is fitting for a given term. 
+To increase the difficulty of the strategy, the distractors are chosen to be at most two vertices away from the edge which represents the correct term, much like in the label-definition-type.
 

### 2022-01-26 quiz.tex
Yet more Quiz.

@@
 %This includes its different application systems, computers, and network components as well as their users.
-Medical informatics students, who are trained for executive positions in information management departments of healthcare institutions, such as hospitals, need a clear terminology of their domain.
+Medical informatics students, who are trained for executive positions in information management (IM) departments of healthcare institutions, such as hospitals, need a clear terminology of their domain.
 This terminology is offered by SNIK~\citep{semantischesnetz,sniktec}, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German), which integrates knowledge extracted from three textbooks~\citep{bb,ob,he} and other sources such as interviews in the form of Linked Open Data.
-SNIK describes information management functions, roles executing these functions and entity types, the information used or updated by these functions.
-Because the textbooks describe those concepts in the abstract sense, and not those of any concrete hospital, those entities are modelled as classes and individuals at the same time, using OWL 2 DL punning.
-Thus, \emph{function}, \emph{role} and \emph{entity type} are metaclasses and the ontology of SNIK is a meta model as depicted in \cref{fig:metamodel}.
+SNIK describes health information management (HIM) functions, roles executing these functions, and entity types, the information used or updated by these functions, see \cref{fig:metamodel}.
+%Because the textbooks describe those concepts in the abstract sense, and not those of any concrete hospital, those entities are modelled as classes and individuals at the same time, using OWL 2 DL punning.
+%Thus, \emph{function}, \emph{role} and \emph{entity type} are metaclasses and the ontology of SNIK is a meta model as depicted in \cref{fig:metamodel}.
 SNIK version \snikversionlink{} contains \sniktriples{} triples describing \snikproperties{} properties and \snikclasses{} classes, of which there are \snikroles{} roles, \snikfunctions{} functions and \snikentitytypes{} entity types.
+%Applications include a graph visualization for exploratiory learning, 
+This paper presents SNIK Quiz, a browser game in which a student answers multiple-choice questions about HIM including an evaluation by domain experts and a student.
+
 %A HIS processes data, information, and knowledge and its management involves planning, monitoring and directing those activities.
@@
 \section{Methods}
-
 \begin{figure*}
@@
 \end{figure*}
+SNIK Quiz is designed as single-player game for English speaking students of Medical Informatics.
+This group is diverse and includes students of Medicine, who may not have a background in Computer Science and Semantic Web technologies.
+Target devices are PCs, tablets and smartphones, so multiple operating systems and input methods need to be supported.
+Thus, SNIK Quiz is designed as an multiple choice quiz with at most four possible answers and published as an open source\footnote{\url{https://github.com/snikproject/snik-quiz}} web application.
 
@@
 \begin{table}[t]
-\caption{Initial Quiz question templates.}
+\caption{Quiz question templates.}
 \label{tab:templates}
@@
 \midrule
-Definition		&Ask for the class that fits the given textbook definition.\\
+\textbf{Definition}		&Ask for the class that fits the given textbook definition.\\
 Example					&What is defined as \enquote{Examination of in and out patients in radiological department}?\\
@@
 \midrule
-Subject			&Ask for the class that is related via a given relation to a given object.\\
-Example			&Who is \emph{involved in} a \emph{healthcare network}?\\
-Distractors		&Labels of other classes (of the same type) that \emph{are not} related via the same relation to the same object.\\
+\textbf{Subject}		&Ask for the class that is related via a given relation to a given object.\\
+						&Key: Who is \emph{involved in} a \emph{healthcare network}?\\
+Example					&Stem: \\
+Distractors				&Labels of other classes (of the same type) that \emph{are not} related via the same relation to the same object, for example .\\
+\midrule
+\textbf{Intertwined}	&\\
+						&Stem: \\
+Example:				&Key: \\
+						&Distractors: \\
+\midrule
+\textbf{CloseMatch}		&\\
+						&Stem: In the context of Strategic Information Management, which one of the following triples belongs together the most?\\
+Example:				&Key: Chief Information Officer – Department of Information management – IM Staff \\
+						&Distractors: Labels of three resources connected in one of the distractor shapes in \cref{fig:intertwined}, such as
+						Chief Information Officer – Strategic Gap – Corporate Strategy\\
+\midrule
+\textbf{}				&\\
+						&Stem: \\
+Example:				&Key: \\
+						&Distractors: \\
 \bottomrule
@@
 
-Clover Quiz~\citep{cloverquiz} shows that ontologies can be used to automatically generate multiple-choice questions over DBpedia~\citep{dbpedia}.
 To provide difficult wrong answers, the so called \emph{distractors}, we use direct neighbours of the classes that represent the correct answers.
@@
 \toprule
-Question			&correct answer	&distractor 1	&distractor 2\\ 
+Question			&Correct answer	&Distractor 1	&Distractor 2\\ 
 \midrule
-In the context of Strategic Information Management, which one of the following triples belongs together the most?
-&Chief Information Officer -- Department of Information management -- Information Management Staff & Chief Information Officer -- Strategic Gap -- Corporate Strategy & Chief Information Officer -- Ticket Evaluation -- Project Monitoring \\
+In the context of Strategic IM, which one of the following triples belongs together the most?
+&Chief Information Officer -- Department of Information management -- IM Staff & Chief Information Officer -- Strategic Gap -- Corporate Strategy & Chief Information Officer -- Ticket Evaluation -- Project Monitoring \\
 \midrule
-In the context of Strategic Information Management, which one of the following triples belongs together the most?
-& Information Management -- IT Service Management -- Operational Information Management & Information Management -- Tactical Management of Information Systems -- Information System & Information Management -- methodology of Information Management -- Operational Management of Information Systems\\
+In the context of Strategic IM, which one of the following triples belongs together the most?
+& IM -- IT Service Management -- Operational IM & IM -- Tactical Management of Information Systems -- Information System & IM -- methodology of IM -- Operational Management of Information Systems\\
 \midrule
@@
 \midrule
-In the context of CIOX, which one of the following triples belongs together the most & Chief Information Officer -- Project Presentation -- Project Waitlist & Chief Information Officer -- Strategic Information Management Planning -- Project Manager & Chief Information Officer -- Project Priorisation -- Project Web Site \\
+In the context of CIOX, which one of the following triples belongs together the most & Chief Information Officer -- Project Presentation -- Project Waitlist & Chief Information Officer -- Strategic IM Planning -- Project Manager & Chief Information Officer -- Project Priorisation -- Project Web Site \\
 \bottomrule
@@
 \section{Discussion}
+Clover Quiz~\citep{cloverquiz} shows that Linked Data knowledge bases can be used to semi-automatically generate cross-domain multiple-choice questions using DBpedia~\citep{dbpedia}.
 systematic review of multiple choice question generation from ontologies

### 2022-01-26 quiz.tex
More question templates.

@@
 \textbf{Definition}		&Ask for the class that fits the given textbook definition.\\
-Example					&What is defined as \enquote{Examination of in and out patients in radiological department}?\\
-Distractors				&Labels of other classes (that have a path of length 2 or less to the correct class.)\\
+						&Stem: \emph{What is defined as \enquote{Examination of in and out patients in radiological department}?}\\
+Example:				&Key: label of the class with that definition, e.g. \emph{CT Scanner}\\
+						&Distractors: Labels of other classes, that have a path of length 2 or less to the correct class., e.g. \emph{Execution of Diagnostic and Therapeutic Procedures}, \emph{Imaging Modality}\\
 \midrule
 \textbf{Subject}		&Ask for the class that is related via a given relation to a given object.\\
-						&Key: Who is \emph{involved in} a \emph{healthcare network}?\\
-Example					&Stem: \\
-Distractors				&Labels of other classes (of the same type) that \emph{are not} related via the same relation to the same object, for example .\\
+						&Stem: Who is \emph{involved in} a \emph{healthcare network}?\\
+Example					&Key: \\
+						&Distractors: Labels of other classes (of the same type) that \emph{are not} related via the same relation to the same object, for example .\\
 \midrule
 \textbf{Intertwined}	&\\
+						&Stem: In the context of Strategic Information Management, which one of the following triples belongs together the most?\\
+Example:				&Key: Chief Information Officer – Department of Information management – IM Staff \\
+						&Distractors: Labels of three resources connected in one of the distractor shapes in \cref{fig:intertwined}, such as
+						Chief Information Officer – Strategic Gap – Corporate Strategy\\
+\midrule
+\textbf{CloseMatch}		&\\
 						&Stem: \\
@@
 \midrule
-\textbf{CloseMatch}		&\\
-						&Stem: In the context of Strategic Information Management, which one of the following triples belongs together the most?\\
-Example:				&Key: Chief Information Officer – Department of Information management – IM Staff \\
-						&Distractors: Labels of three resources connected in one of the distractor shapes in \cref{fig:intertwined}, such as
-						Chief Information Officer – Strategic Gap – Corporate Strategy\\
+\textbf{Contains}		&\\
+						&Stem: \\
+Example:				&Key: \\
+						&Distractors: \\
+\midrule
+\textbf{Label-Definition}	&Present different definitions and ask, which of them fits the given entity.\\
+						&Stem: \emph{How is \enquote{X} defined?}, e.g. \emph{Chief Executive Officer}\\
+Example:				&Key: definition of the entity, e.g. \emph{Responsible for setting and carrying out the strategic plans and policies of an enterprise.}\\
+						&Distractors: labels of entities within a path length of at most 2, e.g. \emph{Anyone who has direct or indirect influence on or interest in a component of an information system.}\\
 \midrule
@@
 
-\paragraph{contains}
-describe contains...
+\paragraph{subject}
+describe subject (contains)...
 
@@
 In this easier variation, two concepts pairs are connected with a shortest path of length 2 instead.
-Examplse are given in \cref{tab:intertwined}.
+%Examplse are given in \cref{tab:intertwined}.
 

### 2022-01-27 quiz.tex
Shorten to 5 pages. Extend template table.

--- 2022-01-26 quiz.tex
@@
 \begin{table}[t]
-\caption{Quiz question templates.}
+\caption{Quiz question templates with examples. When not otherwise noted, distractors are wrong answers that are neighbors of degree at most 2 of the correct answer.}
 \label{tab:templates}
@@
 \midrule
+\textbf{Subject}		&Ask for the subject of a (subject, relation, object) triple given the relation and object.\\
+%						&Distractors are labels of other classes (of the same type) that \emph{are not} related via the same relation to the same object but which are connected to the correct answer with a path of length at most 2.\\
+Stem					&\emph{Who is \emph{responsible for} \emph{Medical Admission}?}\\
+Key						&\emph{Physician}\\
+Distractors				&\emph{Surgeon}, \emph{Health Care Professional}, \emph{Senior Physician}\\
+\midrule
+\textbf{Object}			&Ask for the object of a (subject, relation, object) triple given the subject and relation.\\
+Stem					&\emph{The \emph{Specification Team} is \emph{responsible for}}$\ldots$\\
+Key						&\emph{Functional Specification Document}\\
+Distractors: 			&\emph{Project Team Member}, \emph{Sponsor}, \emph{Defining Project Organization}\\
+\midrule
 \textbf{Definition}		&Ask for the class that fits the given textbook definition.\\
-Example					&What is defined as \enquote{Examination of in and out patients in radiological department}?\\
-Distractors				&Labels of other classes (that have a path of length 2 or less to the correct class.)\\
+Stem					&\emph{What is defined by this?}\\
+						&\enquote{X assures a defined quality of all processes and outcomes of the hospital}\\
+Key						&\emph{Internal Quality Management}\\
+Distractors				&\emph{Activity}, \emph{Complaint}, \emph{Diagnosis}\\
 \midrule
-\textbf{Subject}		&Ask for the class that is related via a given relation to a given object.\\
-						&Key: Who is \emph{involved in} a \emph{healthcare network}?\\
-Example					&Stem: \\
-Distractors				&Labels of other classes (of the same type) that \emph{are not} related via the same relation to the same object, for example .\\
-\midrule
-\textbf{Intertwined}	&\\
-						&Stem: \\
-Example:				&Key: \\
-						&Distractors: \\
+\textbf{Intertwined}	&The correct answer is a triangle in the graph. The distractors instead form different shapes based on the difficulty. See \cref{fig:intertwined}.\\
+Stem					&In the context of Strategic Information Management, which one of the following triples belongs together the most?\\
+Key						&\emph{Chief Information Officer -- Department of Information management -- IM Staff}\\
+Distractor 1			&\emph{Chief Information Officer -- Strategic Gap -- Corporate Strategy}\\
+Distractor 2			&\emph{Chief Information Officer -- Ticket Evaluation -- Project Monitoring}\\
 \midrule
 \textbf{CloseMatch}		&\\
-						&Stem: In the context of Strategic Information Management, which one of the following triples belongs together the most?\\
-Example:				&Key: Chief Information Officer – Department of Information management – IM Staff \\
-						&Distractors: Labels of three resources connected in one of the distractor shapes in \cref{fig:intertwined}, such as
-						Chief Information Officer – Strategic Gap – Corporate Strategy\\
+Stem \\
+Key \\
+Distractors \\
+\midrule
+\textbf{Label-Definition}	&Present different definitions and ask, which of them fits the given entity.\\
+Stem					&\emph{How is \enquote{Chief Executive Officer} defined?}\\
+Key						&\emph{Responsible for setting and carrying out the strategic plans and policies of an enterprise.}\\
+Distractors				&labels of entities within a path length of at most 2, e.g. \emph{Anyone who has direct or indirect influence on or interest in a component of an information system.}\\
 \midrule
@@
 \subsection{Question Types}
-
+\iffalse
 TODO: Add the old question types, but only the b-versions (so no suffix is required in this paper)
@@
 To increase the difficulty of the strategy, the distractors are chosen to be at most two vertices away from the edge which represents the correct term, much like in the label-definition-type.
-
-\paragraph{contains}
-describe contains...
-
-\paragraph{intertwined}
+\fi
 
@@
 \end{table}
+\fi
 
+\iffalse
 Intertwined focuses on the connectivity of the knowledge graph.
@@
 The correct answer is the one where all concepts are directly connected with each other, that is, they form a triangle in the graph, see \cref{fig:intertwined}.
-The distractors should then not form this triangle.
+The distractors should then not form this triangle but a different shape based on the difficulty.
 In the difficult variation of the distractor, one of the three concept pairs is directly connected, one has a shortest path of length 2 and the third one has neither.
 In this easier variation, two concepts pairs are connected with a shortest path of length 2 instead.
-Examplse are given in \cref{tab:intertwined}.
 
@@
 The context of the questions can be augmented with the definitions of the concepts.
+\fi
 
@@
 
+\label{tab:templates} subclasses and superclasses
+because of OWL punning entities are individuals but also classes and can have superclasses and subclasses
+subclasses of the correct answer are semantically correct because subclasses are subsets
+superclasses on the other hand are only partially correct because the relation only holds for the subclass
+example who is responsible for medical admission: Health Care Professional - Physician - Senior Physician
+however SPARQL endpoint doesnt do inferencing so it will not mark that as correct
+we could add the transitive hull for such triples but we think it is good like this, it increases the difficulty and you need to choose the most general correct answer
+
 %We showed that knowledge on the management of information systems in medicine and health care is made publicly available using open standards over several interfaces with different compromises between expressivity and accessibility.
@@
 %\nocite{*} 
-\todo{FJ: Hmmm, ziemlich selbstreferenziell. Brauchen wir da wirklich so viele Publikationen von uns? Welche anderen könnte man noch einbringen? Ich schaue auch mal, was mir so über den Weg läuft.}
 %\bibliographystyle{ios1}
@@
 \bibliography{paper,snik,relatedwork}
-%\section{Notes to Reviewers}
-%This paper includes and extends text and figures from:
-%\citet{snikgraph}
 \end{document}

### 2022-01-27 quiz.tex
Occurence

@@
 \midrule
-\textbf{Intertwined}	&The correct answer is a triangle in the graph. The distractors instead form different shapes based on the difficulty. See \cref{fig:intertwined}.\\
+\textbf{Intertwined}	&The correct answer is a triangle in the graph. The distractors are of different shapes based on the difficulty. See \cref{fig:intertwined}.\\
 Stem					&In the context of Strategic Information Management, which one of the following triples belongs together the most?\\
@@
 \midrule
-\textbf{CloseMatch}		&\\
+\textbf{Close-By}		&\\
 Stem \\
@@
 \midrule
-\textbf{}				&\\
-						&Stem: \\
-Example:				&Key: \\
-						&Distractors: \\
+\textbf{Occurence}				& Asks where a term is set to exist in the used ontologies of SNIK. These ontologies are made up of relations based on the books \cite{ob} and \cite{bb}.\\
+						&Stem:  In which contexts does the term \enquote{Health Insurance Company} occur? \\
+Example:				&Key: Strategic Information Management \\
+						&Distractor 1: Tactical Information Management \\
+						&Distractor 2: Both Contexts \\
+						&Distractor 3: Neither \\
 \bottomrule

### 2022-01-27 quiz.tex
Finish question templates. Add screenshot.

@@
 %SNIK consists of the manually transformed content of three textbooks as Linked Open Data.
-SNIK Quiz demonstrates 
-More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract 
-More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract 
-More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract 
-More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract More abstract 
+We present SNIK Quiz, a browser game in which students answer multiple-choice questions about information management in hospitals based on SNIK.
+The questions are semi-automatically generated using several question templates in order to train basic facts, more complex patterns, and connections between textbooks encoded in SNIK.
 %The data model describes information management functions, roles executing these functions and the information used or updated by these functions.
@@
 %Applications include a graph visualization for exploratiory learning, 
-This paper presents SNIK Quiz, a browser game in which a student answers multiple-choice questions about HIM including an evaluation by domain experts and a student.
-
+%
+\begin{figure}
+\caption{The SNIK Quiz browser game available at \url{https://www.snik.eu/quiz}.}
+\label{fig:snikquiz}
+\centering
+\includegraphics[width=0.6\textwidth]{img/snik-quiz.png}
+\end{figure}
+This paper presents SNIK Quiz, an open source\footnotemark{} browser game, see \cref{fig:snikquiz}, in which students answer multiple-choice questions about HIM, including an evaluation by domain experts and a student.
+\footnotetext{Source code available under the MIT license at \url{https://github.com/snikproject/quiz}. The user interface is a fork of \url{https://github.com/davidrayoussef/react-quiz}.}
 %A HIS processes data, information, and knowledge and its management involves planning, monitoring and directing those activities.
@@
 Target devices are PCs, tablets and smartphones, so multiple operating systems and input methods need to be supported.
-Thus, SNIK Quiz is designed as an multiple choice quiz with at most four possible answers and published as an open source\footnote{\url{https://github.com/snikproject/snik-quiz}} web application.
+Thus, SNIK Quiz is designed as an multiple choice quiz with at most four possible answers and published as an open source web application.
 
@@
 %						&Distractors are labels of other classes (of the same type) that \emph{are not} related via the same relation to the same object but which are connected to the correct answer with a path of length at most 2.\\
-Stem					&\emph{Who is \emph{responsible for} \emph{Medical Admission}?}\\
+Stem					&\emph{Who is responsible for Medical Admission?}\\
 Key						&\emph{Physician}\\
@@
 \textbf{Object}			&Ask for the object of a (subject, relation, object) triple given the subject and relation.\\
-Stem					&\emph{The \emph{Specification Team} is \emph{responsible for}}$\ldots$\\
+Stem					&\emph{The Specification Team is responsible for}$\ldots$\\
 Key						&\emph{Functional Specification Document}\\
@@
 \midrule
-\textbf{Definition}		&Ask for the class that fits the given textbook definition.\\
-Stem					&\emph{What is defined by this?}\\
-						&\enquote{X assures a defined quality of all processes and outcomes of the hospital}\\
+\textbf{Definition}		&Ask for the entity that fits the given textbook definition.\\
+Stem					&\emph{What is defined by this? \enquote{X assures a defined quality of all processes and outcomes of the hospital}}\\
 Key						&\emph{Internal Quality Management}\\
@@
 \midrule
-\textbf{Intertwined}	&The correct answer is a triangle in the graph. The distractors are of different shapes based on the difficulty. See \cref{fig:intertwined}.\\
+\textbf{Definitions}	&Present different definitions and ask, which of them fits the given entity.\\
+Stem					&\emph{What defines the term Data Warehouse System?}\\
+Key						&\emph{Application component that contains data which have been extracted from other application components, in order to support either hospital management or clinical research.}\\
+Distractors				&\emph{Defines the hospital’s long-term strategic goals}\\
+						&\emph{An application component where the controlling rules for data processing are implemented as executable software.}\\
+						&\emph{Summarizes monitored key performance indicators (KPIs) and compares them to the expected future state}\\
+\midrule
+\textbf{Intertwined}	&The correct answer is a triangle in the graph. The distractors instead form different shapes based on the difficulty. See \cref{fig:intertwined}.\\
 Stem					&In the context of Strategic Information Management, which one of the following triples belongs together the most?\\
@@
 \midrule
-\textbf{Close-By}		&\\
-Stem \\
-Key \\
-Distractors \\
-\midrule
-\textbf{Label-Definition}	&Present different definitions and ask, which of them fits the given entity.\\
-Stem					&\emph{How is \enquote{Chief Executive Officer} defined?}\\
-Key						&\emph{Responsible for setting and carrying out the strategic plans and policies of an enterprise.}\\
-Distractors				&labels of entities within a path length of at most 2, e.g. \emph{Anyone who has direct or indirect influence on or interest in a component of an information system.}\\
-\midrule
-\textbf{Occurence}				& Asks where a term is set to exist in the used ontologies of SNIK. These ontologies are made up of relations based on the books \cite{ob} and \cite{bb}.\\
+\textbf{Occurence}		&Asks where a term is set to exist in the used ontologies of SNIK. These ontologies are made up of relations based on the books \cite{ob} and \cite{bb}.\\
 						&Stem:  In which contexts does the term \enquote{Health Insurance Company} occur? \\
@@
 						&Distractor 3: Neither \\
+\textbf{CloseMatch}		&Transfer knowledge about one textbook~\citep{bb} to the other~\citep{ob} by confirming or denying statements about entity pairs that are marked as near equivalent.\\
+Stem					&\emph{In the Strategic Information Management, the Consultant is associated with the Long-Term HIS Planning, while in the Tactical Information Management, the Consultant is responsible for the Functional Specification Document.}\\
+Key						&True\\
+Distractor				&False\\
 \bottomrule
@@
 It is unclear, whether the CIO is also responsible for operational information management.
-
-% TODO: reformulate now that its not future work anymore but part of the paper
-An ongoing bachelor's thesis~\citep{snikquiz} is investigating different strategies of generating complex questions, answers and distrators.
 
@@
 \includegraphics[width=0.7\textwidth]{img/intertwined_cml.png} 
-\caption{Left to right: correct answer, difficult and easy distractor for the \emph{intertwined} question type.}
+\caption{Left to right: correct answer, difficult and easy distractor for the \emph{intertwined} question type~\citep{snikquizba}.}
 \label{fig:intertwined}
@@
 \toprule
-			&label-definition	&definition	&contains	&intertwined	&closeMatch	&occurence\\
+			&definitions		&definition	&subject	&intertwined	&closeMatch	&occurence\\
+\midrule
 interview 1	&3					&4			&5--6		&7				&8			&7\\
@@
 \end{tabularx}
-\caption{Rating given by interviewees based on the complexity of the questions on a scale between 1 and 10 with 10 representing maximum complexity.}
+\caption{Rating given by interviewees based on the complexity of the questions on a scale between 1 and 10 with 10 representing maximum complexity~\citep{snikquizba}.}
 \label{tabelle:eval_komplex}

### 2022-01-27 quiz.tex
Remove duplicate bib entry. Shorten to 5 pages.

@@
 \begin{table}[t]
-\caption{Quiz question templates with examples. When not otherwise noted, distractors are wrong answers that are neighbors of degree at most 2 of the correct answer.}
+\caption{Quiz question templates with examples.
+When not otherwise noted, distractor entites are wrong answers that are neighbors of degree at most 2 of the correct answer.}%, with all being of the same type, see \cref{fig:metamodel}.}
 \label{tab:templates}
@@
 Key						&\emph{Application component that contains data which have been extracted from other application components, in order to support either hospital management or clinical research.}\\
-Distractors				&\emph{Defines the hospital’s long-term strategic goals}\\
-						&\emph{An application component where the controlling rules for data processing are implemented as executable software.}\\
-						&\emph{Summarizes monitored key performance indicators (KPIs) and compares them to the expected future state}\\
+Distractors				&\emph{Defines the hospital’s long-term strategic goals},
+						\emph{An application component where the controlling rules for data processing are implemented as executable software},
+						\emph{Summarizes monitored key performance indicators (KPIs) and compares them to the expected future state}\\
 \midrule
-\textbf{Intertwined}	&The correct answer is a triangle in the graph. The distractors instead form different shapes based on the difficulty. See \cref{fig:intertwined}.\\
-Stem					&In the context of Strategic Information Management, which one of the following triples belongs together the most?\\
+\textbf{Intertwined}	&The correct answer and distractors are connected as shown in \cref{fig:intertwined}.\\
+Stem					&\emph{In the context of Strategic Information Management, which one of the following triples belongs together the most?}\\
 Key						&\emph{Chief Information Officer -- Department of Information management -- IM Staff}\\
@@
 \midrule
-\textbf{Occurence}		&Asks where a term is set to exist in the used ontologies of SNIK. These ontologies are made up of relations based on the books \cite{ob} and \cite{bb}.\\
-						&Stem:  In which contexts does the term \enquote{Health Insurance Company} occur? \\
-Example:				&Key: Strategic Information Management \\
-						&Distractor 1: Tactical Information Management \\
-						&Distractor 2: Both Contexts \\
-						&Distractor 3: Neither \\
+\textbf{Occurence}		&Ask whether a given term is defined in one of the textbooks \cite{bb,ob}, both or neither.\\
+Stem					&\emph{In which contexts does the term \enquote{Health Insurance Company} occur?}\\
+Key						&\emph{Strategic Information Management}\\
+Distractors				&\emph{Tactical Information Management, Both Contexts, Neither}\\
+\midrule
 \textbf{CloseMatch}		&Transfer knowledge about one textbook~\citep{bb} to the other~\citep{ob} by confirming or denying statements about entity pairs that are marked as near equivalent.\\
-Stem					&\emph{In the Strategic Information Management, the Consultant is associated with the Long-Term HIS Planning, while in the Tactical Information Management, the Consultant is responsible for the Functional Specification Document.}\\
-Key						&True\\
-Distractor				&False\\
+Stem					&\emph{In the Strategic Information Management, the Consultant is associated with the Long-Term HIS Planning, while in the Tactical Information Management, the Consultant is responsible for the Functional Specification Document.} (Key: True, Distractor: False)\\
+%Key						&True\\
+%Distractor				&False\\
 \bottomrule

### 2022-01-27 quiz.tex
Add SPARQL query.

@@
 SNIK is a knowledge base about the management of health information systems generated by extracting Linked Data from textbooks and other sources.
+SNIK describes health information management functions, roles executing these functions, and entity types, the information used or updated by these functions.
 %SNIK consists of the manually transformed content of three textbooks as Linked Open Data.
@@
 \end{table}
+%
+\begin{lstlisting}[captionpos=b, caption=SPARQL query generating the key and distractors for the \emph{definition} questions., label=lst:sparql,basicstyle=\ttfamily\footnotesize,frame=single]
+SELECT SAMPLE(replace(str(?def),str(?cl),"X","i") as ?def) SAMPLE(str(?cl) as ?cl) SAMPLE(str(?a1l) as ?a1l) SAMPLE(str(?a2l) as ?a2l) SAMPLE(str(?a3l) as ?a3l) {
+  ?class a owl:Class.
+  ?class rdfs:label ?cl.
+  FILTER(LANGMATCHES(LANG(?cl),"en"))
+  ?class skos:definition ?def.
+  FILTER(STRLEN(?def)>10&&STRLEN(?def)<600).
+  FILTER(LANGMATCHES(LANG(?def),"en"))
+
+  ?class (!(meta:subTopClass|rdf:type)){1,2} ?a1,?a2,?a3.
+  owl:Class ^a ?a1,?a2,?a3.
+  FILTER(?class!=?a1&&?class!=?a2&&?class!=?a3&&?a1<?a2&&?a2<?a3)
+
+  ?a1 rdfs:label ?a1l. FILTER(LANGMATCHES(LANG(?a1l),"en"))
+  ?a2 rdfs:label ?a2l. FILTER(LANGMATCHES(LANG(?a2l),"en"))
+  ?a3 rdfs:label ?a3l. FILTER(LANGMATCHES(LANG(?a3l),"en"))
+} GROUP BY ?class limit 1000
+\end{lstlisting}
+%
+The questions are semi-automatically generated using several question templates as shown in \cref{tab:templates} to train basic facts, more complex patterns, as well as connections between textbooks encoded in SNIK.
 
 To provide difficult wrong answers, the so called \emph{distractors}, we use direct neighbours of the classes that represent the correct answers.
-Due to the limited nesting capabilties of SPARQL, which does not support loops but only subqueries that cannot access variables declared outside their scope, querying for distinct sets of 4 neighbours, where exactly one was connected over a certain property to a certain object, was not possible with our Virtuoso SPARQL endpoint initially.
+%Due to the limited nesting capabilties of SPARQL, which does not support loops but only subqueries that cannot access variables declared outside their scope, querying for distinct sets of 4 neighbours, where exactly one was connected over a certain property to a certain object, was not possible with our Virtuoso SPARQL endpoint initially.
 We circumvented the timeout by calculating the neighbour-relation as a first step and uploading it to the endpoint.

### 2022-01-27 quiz.tex
Finish SNIK Quiz paper.

@@
 \usepackage{graphicx}
+\usepackage{float}
 \usepackage{textcomp}
@@
 This terminology is offered by SNIK~\citep{semantischesnetz,sniktec}, the Semantic Network of Information Management in Hospitals (\enquote{Krankenhaus} in German), which integrates knowledge extracted from three textbooks~\citep{bb,ob,he} and other sources such as interviews in the form of Linked Open Data.
-SNIK describes health information management (HIM) functions, roles executing these functions, and entity types, the information used or updated by these functions, see \cref{fig:metamodel}.
+In order to specify, which information should be extracted from the books and to facilitate comparisons, we use a common data model.
+Because processed textbooks contain abstract knowledge instead of information about any specific hospital, all concepts are modelled as classes and instances simultaneusly, using OWL punning.
+We thus call our data model the \enquote{meta model} in accordance with the term's definition as a shared modelling language~\citep[p.~8]{ob}.
+The central entities of the meta model are enterprise functions, roles executing these functions, and entity types, the information used or updated by these functions, see \cref{fig:metamodel}.
 %Because the textbooks describe those concepts in the abstract sense, and not those of any concrete hospital, those entities are modelled as classes and individuals at the same time, using OWL 2 DL punning.
@@
 Thus, SNIK Quiz is designed as an multiple choice quiz with at most four possible answers and published as an open source web application.
+We employ an approach similar to Clover Quiz~\citep{cloverquiz}, which shows that Linked Data knowledge bases can be used to semi-automatically generate cross-domain multiple-choice questions on DBpedia~\citep{dbpedia}.
+For each question (\emph{stem}), we generate a correct answer (\emph{key}) and one or more incorrect answers (\emph{distractors}).
+The data is semi-automatically generated using SPARQL queries, such as \cref{lst:sparql}, on the public SPARQL endpoint of SNIK\footnote{\url{https://www.snik.eu/sparql}} followed by minimal postprocessing to achieve more natural looking questions.
+The graph structure of SNIK allows the generation of difficult to guess distractors by using entities that are semantically close to the key, using path lengths of at most 2 in the graph.
 
@@
 
-%\section{Data Quality}
-%- extractors not always same people who wrote book/part of book
-%- mistake when extracting
-%- reasoning does not work (test)
-%\begin{itemize}
-%\item Multiple Subtops
-%\item Inconsistent Subtop with Subclass
-%\item SKOS Link to Different Subtop
-%\item Subclass Cycles
-%\item Missing superclass
-%\item Undefined Objects
-%\item Domain Violation
-%\item Range Violation
-%\item Class URL Naming Convention Violations
-%\item Property URL Naming Convention Violations
-%\item Missing Definition
-%\item Literals with Semicolons
-%\item Classes with too many subclasses
-%\item No Restriction
-%\item Non-HTTP URI
-%\item OWL 2 DL: Undefined Properties
-%\item OWL 2 DL: Multiply Defined Properties
-%\item OWL 2 DL: Class Typing Constraints
-%\end{itemize}
-%-> use SHAPE/SHACL ?
+\section{Results}\label{sec:application}
+The resulting question data is generated by several templates, as shown in \cref{tab:templates}, which are designed to either train basic facts, more complex patterns in one of the textbooks or connections between the different textbooks modelled in SNIK.
 
-%\section{Data Model}\label{sec:architecture}
-%The initial structure of the meta model is defined in \citet{domaene}.
-%The \enquote{meta model} defines three basic disjunctive classes and their possible relations: Roles (who), Function (does what) and Entity Types (and which information is therefore needed).
-%A set of modular subontologies define subclasses of those three classes and their relations as described by a certain knowledge source about information management in hospitals:
-%The \textbf{Semantic Network of Information Management in Hospitals} (SNIK\footnote{Hospital means \enquote{Krankenhaus} in German.}) is a modular OWL 2 DL ontology.
-%from different sources:% three textbooks, an interview and a standard.
-
-\iffalse
-\section{Data Set Description}\label{sec:dsd}
-\begin{table*}
-\caption{use of established vocabularies}
-\label{tab:templates}
-\begin{tabulary}{\textwidth}{lL}
-\toprule
-\textbf{Vocabulary}	&\textbf{Description}\\
-\midrule
-&\\
-&\\
-&\\
-\bottomrule
-\end{tabulary}
-\end{table*}
-\fi
-
-%\section{Data Transformation Process}
-%\subsection{Interlinking}
-%As shown in \cref{...}, we aligned our dataset to several others:
-%We manually aligned our dataset to HITO, the Health IT Ontology~\citep{hitometh2}
-%We automatically aligned our dataset to:
-%- DBpedia~\citep{dbpedia}, as  
-%- 
-%\section{Lifecycle}\label{sec:application}
-%Virtuoso SPARQL endpoint is used for querying and as data source for all our applications
-
-%After the initial extraction step, SNIK is published and users suggest problems or enhancements.%Besides issues of collaborative editing, this required a regular export and reupload, which delayed the effect of changes in SNIK to the user facing applications.
-%- git repository (link): collaborative editing but protege can cause large diffs -> people who understand rdf enough to use text editor ttl/rdfxml 
-%- use ontowiki: slow but user friendly, undo function,
-%- SPARUL queries
-%The technical environment of those services is described in~\citet{sniktec}.
-
-%\section{Applications}
-
-\begin{table}[t]
+\begin{table}[h]
 \caption{Quiz question templates with examples.
@@
 \midrule
-\textbf{Occurence}		&Ask whether a given term is defined in one of the textbooks \cite{bb,ob}, both or neither.\\
+\textbf{Occurence}		&Ask whether a given term is defined in one of the textbooks \cite{bb,ob}, both or neither~\citep{snikquizba}.\\
 Stem					&\emph{In which contexts does the term \enquote{Health Insurance Company} occur?}\\
@@
 \midrule
-\textbf{CloseMatch}		&Transfer knowledge about one textbook~\citep{bb} to the other~\citep{ob} by confirming or denying statements about entity pairs that are marked as near equivalent.\\
+\textbf{CloseMatch}		&Transfer knowledge about one textbook~\citep{bb} to the other~\citep{ob} by confirming or denying statements about entity pairs that are marked as near equivalent~\citep{snikquizba}.\\
 Stem					&\emph{In the Strategic Information Management, the Consultant is associated with the Long-Term HIS Planning, while in the Tactical Information Management, the Consultant is responsible for the Functional Specification Document.} (Key: True, Distractor: False)\\
@@
 %
-\begin{lstlisting}[captionpos=b, caption=SPARQL query generating the key and distractors for the \emph{definition} questions., label=lst:sparql,basicstyle=\ttfamily\footnotesize,frame=single]
+\begin{lstlisting}[float, caption=SPARQL query generating the key and distractors for the \emph{definition} questions., label=lst:sparql,basicstyle=\ttfamily\footnotesize,frame=single]
 SELECT SAMPLE(replace(str(?def),str(?cl),"X","i") as ?def) SAMPLE(str(?cl) as ?cl) SAMPLE(str(?a1l) as ?a1l) SAMPLE(str(?a2l) as ?a2l) SAMPLE(str(?a3l) as ?a3l) {
@@
 %
-The questions are semi-automatically generated using several question templates as shown in \cref{tab:templates} to train basic facts, more complex patterns, as well as connections between textbooks encoded in SNIK.
-
-To provide difficult wrong answers, the so called \emph{distractors}, we use direct neighbours of the classes that represent the correct answers.
 %Due to the limited nesting capabilties of SPARQL, which does not support loops but only subqueries that cannot access variables declared outside their scope, querying for distinct sets of 4 neighbours, where exactly one was connected over a certain property to a certain object, was not possible with our Virtuoso SPARQL endpoint initially.
-We circumvented the timeout by calculating the neighbour-relation as a first step and uploading it to the endpoint.
+%We circumvented the timeout by calculating the neighbour-relation as a first step and uploading it to the endpoint.
 %
 %SNIK Quiz is freely available as an open source web application.
-Because SNIK only contains the knowledge that the source textbooks describe, it does not contain the complete domain of Hospital Information Management.
-As such, negative questions are problematic, as the given relationship could hold in the real world but not be described in the textbook source of the class.
-The same problem concerns the distractors of positive questions.
-So this problem cannot be avoided.
-However, this is one of the reasons that we do not ask count questions like \enquote{How many functions is the CIO responsible for?}, which do not help much for learning anyways. % improve writing
-Relationships could also hold implicitly through the subclass hierarchy.
-For example, \aurl{bb}{ChiefInformationOfficer} is a subclass of \aurl{bb}{InformationManagementStaff}, which is responsible for \aurl{bb}{OperationalInformationManagement}.
-It is unclear, whether the CIO is also responsible for operational information management.
 
-\subsection{Question Types}
-\iffalse
-TODO: Add the old question types, but only the b-versions (so no suffix is required in this paper)
-
-\paragraph{label-definition}
-This gives the user a definition and asks for  the correct term.
-The terms to be considered are chosen from terms, which are positioned within a distance of two vertices in the graph.
-So, relative semantic similarity can be assumed.
-This makes the label-definition-approach especially worthwhile when basic knowledge of terms of information management is required.
-
-% set this as second strat to gain constistency to table 5 
-\paragraph{definition}
-The definition approach focuses on the question of which definition is fitting for a given term. 
-To increase the difficulty of the strategy, the distractors are chosen to be at most two vertices away from the edge which represents the correct term, much like in the label-definition-type.
-\fi
+%The definition approach focuses on the question of which definition is fitting for a given term. 
+%To increase the difficulty of the strategy, the distractors are chosen to be at most two vertices away from the edge which represents the correct term, much like in the label-definition-type.
 
@@
 
-\iffalse
-\begin{table}
-\begin{tabulary}{\textwidth}{LLLL}
-\toprule
-Question			&Correct answer	&Distractor 1	&Distractor 2\\ 
-\midrule
-In the context of Strategic IM, which one of the following triples belongs together the most?
-&Chief Information Officer -- Department of Information management -- IM Staff & Chief Information Officer -- Strategic Gap -- Corporate Strategy & Chief Information Officer -- Ticket Evaluation -- Project Monitoring \\
-\midrule
-In the context of Strategic IM, which one of the following triples belongs together the most?
-& IM -- IT Service Management -- Operational IM & IM -- Tactical Management of Information Systems -- Information System & IM -- methodology of IM -- Operational Management of Information Systems\\
-\midrule
-In the context of CIOX, which one of the following triples belongs together the most & Chief Information Officer -- Jour Fixe Execution -- Project Monitoring & Chief Information Officer -- Project Risk -- Strategic HIS Planning & Chief Information Officer -- IT General Ledger Account Planning -- Project Portfolio \\
-\midrule
-In the context of CIOX, which one of the following triples belongs together the most & Chief Information Officer -- Project Presentation -- Project Waitlist & Chief Information Officer -- Strategic IM Planning -- Project Manager & Chief Information Officer -- Project Priorisation -- Project Web Site \\
-\bottomrule
-\end{tabulary}
-\caption{Examples of generated \emph{intertwined}-questions and answers.}
-\label{tab:intertwined}
-\end{table}
-\fi
-
-\iffalse
-Intertwined focuses on the connectivity of the knowledge graph.
-In each answer, three related concepts are given.
-The correct answer is the one where all concepts are directly connected with each other, that is, they form a triangle in the graph, see \cref{fig:intertwined}.
-The distractors should then not form this triangle but a different shape based on the difficulty.
-In the difficult variation of the distractor, one of the three concept pairs is directly connected, one has a shortest path of length 2 and the third one has neither.
-In this easier variation, two concepts pairs are connected with a shortest path of length 2 instead.
-
-\paragraph{closeMatch}
-\begin{table}[h]
-\begin{tabulary}{\textwidth}{Ll}
-\toprule
-question	& correct\\
-\midrule
-In the Tactical Information Management, the Project Execution uses the Resource, while in the Strategic Information Management, the Project Execution uses the Resource.											&True \\
-In the Strategic Information Management, the Project Management Board is responsible for  the Project Monitoring, while in the Tactical Information Management, the Project Steering Committee uses the Project.	&False \\
-\bottomrule
-\end{tabulary}
-\caption{Examples of generated \emph{closeMatch}-questions.}
-\label{tab:closematch}
-\end{table}
-
-This question type tests the connection between a textbook about tactical information management~\cite{ob} and one about strategic information management~\cite{bb} based on the existing \aurl{skos}{closeMatch} interlinks.
-The aim is to help students transfer knowledge about one textbook to the other.
-The questions have the form: \emph{\enquote{In [Tactical / Strategic IM], the [subject] [predicate] the [object], while in [Strategic / Tactical IM], the [subject] [predicate] the [object].}} with two possible answers: true and false.
-Examplarly questions are given in \cref{tab:closematch}.
+%This question type tests the connection between a textbook about tactical information management~\cite{ob} and one about strategic information management~\cite{bb} based on the existing \aurl{skos}{closeMatch} interlinks.
+%The aim is to help students transfer knowledge about one textbook to the other.
+%The questions have the form: \emph{\enquote{In [Tactical / Strategic IM], the [subject] [predicate] the [object], while in [Strategic / Tactical IM], the [subject] [predicate] the [object].}} with two possible answers: true and false.
+%Examplarly questions are given in \cref{tab:closematch}.
 
@@
 
+\iffalse
 \paragraph{occurence}
@@
 
-\section{Results}\label{sec:application}
-The question types definition-a and contains-a were used to generate 1231 English questions, see \cref{tab:templates}.
-It was used by students from the Universities of Amsterdam, Heidelberg and Leipzig during the international Frank van Swieten lectures in 2019.
-
+%The question types definition and subject were used to generate 1231 English questions, see \cref{tab:templates}.
+The question types definition and subject were presented in a first prototype of SNIK Quiz to students of Medical Informatics from the Universities of Amsterdam, Heidelberg and Leipzig during the international Frank van Swieten lectures in 2019 and were positively received.
+An extended version of SNIK Quiz including the query types definition, definitions, subject, intertwined, closeMatch and occurrence was rated by two experts of hospital information management and one student of Medical Informatics, see \cref{tab:eval}.
 %The technical environment of SNIK is characterized in previous work [11].
 
-\subsection{Evaluation}
 \iffalse
@@
 \toprule
-			&definitions		&definition	&subject	&intertwined	&closeMatch	&occurence\\
+			&definition		&definitions	&subject	&intertwined	&closeMatch	&occurence\\
 \midrule
-interview 1	&3					&4			&5--6		&7				&8			&7\\
-interview 2	&4					&6			&7			&7--8			&7--8		&5\\
+interview 1	&4					&3			&5--6		&7				&8			&7\\
+interview 2	&6					&4			&7			&7--8			&7--8		&5\\
 interview 3	&3					&3			&6			&8				&8			&7\\
@@
 \end{tabularx}
-\caption{Rating given by interviewees based on the complexity of the questions on a scale between 1 and 10 with 10 representing maximum complexity~\citep{snikquizba}.}
-\label{tabelle:eval_komplex}
+\caption{Rating given by interviewees based on the subjective complexity of the questions on a scale between 1 and 10 with 10 representing maximum complexity~\citep{snikquizba}.}
+\label{tab:eval}
 \end{table}
@@
 \section{Discussion}
-Clover Quiz~\citep{cloverquiz} shows that Linked Data knowledge bases can be used to semi-automatically generate cross-domain multiple-choice questions using DBpedia~\citep{dbpedia}.
-systematic review of multiple choice question generation from ontologies
-complex questions, 
-\cite{ontologybasedmultiplechoice}
+While the basic question types were received positively during the Frank van Swieten lectures and the evaluation ratings on subjective complexity show promising results, qualitative feedback on the more complex question types was mixed.
+One of the experts questioned the didactic value of the \emph{intertwined} questions, as the users cannot know the exact graph structure of SNIK and thus may difficulties deciding, which triplets of concepts are most strongly connected.
+While the closeMatch were rated as most complex, the complexity was seen mostly seen as grammatical and not enough on the didactically relevant relationships between the entities.
+Contrary to the initial assumption that well explained questions would be beneficial, shorter questions were preferred as they could be read more quickly.
 
-performance 
-static html and javascript
-client side
+%However, only certain question types are feasible with SNIK.
+%Because SNIK only contains the knowledge that the source textbooks describe, it does not contain the complete domain of Hospital Information Management.
+%As such, negative questions are problematic, as the given relationship could hold in the real world but not be described in the textbook source of the class (open world assumption).
+%The same problem concerns the distractors of positive questions.
+%So this problem cannot be avoided.
+%This is one of the reasons that we do not ask count questions like \enquote{How many functions is the CIO responsible for?}.
+%Relationships could also hold implicitly through the subclass hierarchy.
+%For example, \aurl{bb}{ChiefInformationOfficer} is a subclass of \aurl{bb}{InformationManagementStaff}, which is responsible for \aurl{bb}{OperationalInformationManagement}.
+%It is unclear, whether the CIO is also responsible for operational information management.
+%systematic review of multiple choice question generation from ontologies
+%complex questions, 
+%\cite{ontologybasedmultiplechoice}
 
-\label{tab:templates} subclasses and superclasses
-because of OWL punning entities are individuals but also classes and can have superclasses and subclasses
-subclasses of the correct answer are semantically correct because subclasses are subsets
-superclasses on the other hand are only partially correct because the relation only holds for the subclass
-example who is responsible for medical admission: Health Care Professional - Physician - Senior Physician
-however SPARQL endpoint doesnt do inferencing so it will not mark that as correct
-we could add the transitive hull for such triples but we think it is good like this, it increases the difficulty and you need to choose the most general correct answer
+%\cref{tab:templates} subclasses and superclasses
+%because of OWL punning entities are individuals but also classes and can have superclasses and subclasses
+%subclasses of the correct answer are semantically correct because subclasses are subsets
+%superclasses on the other hand are only partially correct because the relation only holds for the subclass
+%example who is responsible for medical admission: Health Care Professional - Physician - Senior Physician
+%however SPARQL endpoint doesnt do inferencing so it will not mark that as correct
+%we could add the transitive hull for such triples but we think it is good like this, it increases the difficulty and you need to choose the most general correct answer
 
-%We showed that knowledge on the management of information systems in medicine and health care is made publicly available using open standards over several interfaces with different compromises between expressivity and accessibility.
 %It can be combined with other knowledge in biomedical and health informatics and in other disciplines.
@@
 \section{Conclusions}
+SNIK Quiz shows that knowledge on the management of information systems in medicine and health care can be used to semi-automatically generate multiple-choice questions.
+While preliminary evaluations show promising results, the complex question types need to be investigated further.
+Future work should also include a quantitative evaluation of learning efficiency when supplementing courses with SNIK Quiz.
+%Use of neural networks and language models in combination with linked data to generate more natural questions and answers is another avenue for further exploration.
 

### 2022-01-28 quiz.tex
Change title.
Move SNIK Quiz paper into its own repository at https://github.com/snikproject/paper-snik-quiz.

@@
+% arara: pdflatex if changed(tex): { options: [-halt-on-error] }
 \documentclass{IOS-Book-Article}     %[seceqn,secfloat,secthm]
@@
 %\pretitle{Pretitle}
-\title{SNIK Quiz: Automatic Generation of Multiple Choice Questionnaires about Information Management in Hospitals}
+\title{SNIK Quiz: A Multiple Choice Game about Information Management in Hospitals}
 %\thanks{This work is supported by the DFG (German Research Foundation) under the Project SNIK, Grant no. 1605/7-1 and 1387/8-1.}%bug: will create an empty first page
@@
 \markboth{}{May 2022\hb}
-%\subtitle{}
+\subtitle{The SNIK Quiz paper has moved to \url{https://github.com/snikproject/paper-snik-quiz}.}
 
@@
 SNIK describes health information management functions, roles executing these functions, and entity types, the information used or updated by these functions.
-%SNIK consists of the manually transformed content of three textbooks as Linked Open Data.
 We present SNIK Quiz, a browser game in which students answer multiple-choice questions about information management in hospitals based on SNIK.
@@
 \end{frontmatter}
-
 \section{Introduction}
@@
 %The data is continously revised: users constantly report wrong or missing data in the visualization, which is then corrected, respectively revised, by the researchers of the project.
-
 \iffalse
